### 2014-09-05 paper.tex
Initial commit.

\documentclass{article}

\usepackage[utf8]{inputenc}

\title{LMS paper}
\author{T. Ben Thompson, Brendan J. Meade, All the people}
\date{November 2013}
\usepackage{graphicx}

\begin{document}

\maketitle

A chunk of text.

An equation:
$\nabla^2 u_t = 0$

\begin{figure}[h!]
  \centering
     \includegraphics[width=5in]{log_energy_density.pdf}
   \caption{Note the audible tearing of a heart-sick planet}
\end{figure}

\end{document}

### 2014-09-10 paper.tex
Some changes.
Fixed image ref.
Hook test 1.
Test again.
latexdiff

@@
 
-\title{LMS paper}
-\author{T. Ben Thompson, Brendan J. Meade, All the people}
-\date{November 2013}
+\title{Rapid shortening at the eastern margin of the Tibetan plateau prior to the 2008 Mw=7.9 Wenchuan earthquake}
+\author{T. Ben Thompson, Brendan J. Meade}
+\date{September 2014}
 \usepackage{graphicx}
@@
 
-A chunk of text.
+\section{Abstract}
+The Longmen Shan is the steepest topographic front of the India-Asia collision and was the site of the Mw=7.9 Wenchuan earthquake.
+Shortening estimates across the Longmen Shan provide strain accumulation rates and clarify the eastward extrusion of the Tibetan plateau.
+Here, to explain the interseismic GPS velocities across the greater Longmen Shan region, we develop a boundary element model including earthquake cycle effects, topography, the westward dipping Beichuan fault, and a ~20 km deep, shallowly dipping, detachment.
+The detachment is inferred from observations of slip during and after the Wenchuan earthquake and from structural considerations.
+Previous analyses which neglected the detachment and earthquake cycle effects have found shortening rates near zero.
+In contrast, we find that interseismic GPS data are consistent with a shortening rate of 6$\pm$2 mm/yr.
+These results suggest that the Longmen Shan is an active fold-and-thrust belt with Wenchuan style earthquake recurrence intervals of <600 years.
 
-An equation:
-$\nabla^2 u_t = 0$
+\section{Introduction}
+The Longmen Shan range is located on the eastern margin of the Tibetan and rises 4000m over 30km (CHECK THESE NUMBERS ON GOOGLE EARTH).
+The 2008 $\mathrm{M_w}$ 7.9 Wenchuan earthquake ruptured 250km along strike on the Beichuan and Pengguan faults, resulting in approximately XXXXX fatalities (CITE STONE)
+Prior geodetic estimates suggest almost zero horizontal shortening (Chen 2000, Shen 2005, etc.), in contrast with the dominant thrust slip-sense during the Wenchuan rupture. 
+The lack of shortening is reconciled with the steep topography by appealing to a theory of lower crustal inflation whereby high topography can be formed without upper crustal shortening (CITE clark, royden, england, molnar, kirby, etc)
+However, the magnitude and location of the Wenchuan earthquake suggest a more typical fold-and-thrust belt.
+Using seismic reflection data, Hubbard and Shaw find almost 100\% shortening in the range-front 
+Following on these works, there is a growing body of literature supporting the interpretation of the Longmen Shan as a fold-and-thrust belt (CITE, CITE, CITE) as opposed to a product of viscous lower crustal inflation.
 
-\begin{figure}[h!]
-  \centering
-     \includegraphics[width=5in]{log_energy_density.pdf}
-   \caption{Note the audible tearing of a heart-sick planet}
-\end{figure}
+Consistent with the fold-and-thrust mechanism
+However, it remains to be explained how slow present-day geodetic shortening estimates can be reconciled with the Wenchuan earthquake and the large structurally-observed shortening.
+Here, we present a model that explains the regional geodetic velocities as the effect of a complex interseismically locked fault geometry, including a 20km deep detachment as well as the range-front thrust Beichuan fault.
+This geometry is consistent with the structural interpretations of Hubbard,Shaw,Klinger, with coseismic and postseismic Wenchuan slip distributions that attribute significant motion to a deep detachment (CITE QI, FIELDING).
+Also, seismic observations of P/sP conversion in the crust are suggestive of a low velocity layer near 20km depth (CITE ZHANG ET. AL. CRUSTAL STRUCTURE). 
+Using this range-front thrust plus detachment geometry, we infer an average horizontal shortening.
+Next, we show how this geometry can drives the steep velocity gradients far to the northwest of the range front, thus explaining the absence of a range-front interseismic velocity gradient.
+Finally, we discuss the effects of Longmen Shan fault geometry on regional seismic hazard.
+
+The Longmen Shan are a mountain range at boundary between the Tibetan Plateau and the Sichuan Basin.
+West of the Longmen Shan, in the eastern Tibetan plateau, surface GPS velocity vectors rotate from northeasterly to easterly or south-easterly.
+This rotation has been attributed to either the extrusion of crustal blocks or topographically driven viscous flow impacting a cold, elastically and viscously rigid, Sichuan basin.
+Because of its uniquely steep topographic gradient, lack of foreland basin, etc..., the Longmen Shan is a natural testing ground for these hypotheses.
+The steep topographic front of the Longmen Shan has been explained by both horizontal shortening (cite meade, hubbard/shaw) and lower crustal inflation (cite clark, royden, england, molnar, others). 
+The contrast between high topography and low shortening rates is a central point in this debate.
+Cite kirby
+The Mw 7.9 2008 Wenchuan earthquake ruptured 250km NE-SW along the Pengguan and Beichuan faults in the Longmenshan (cite xu 2008);e .
+Average slip was 4m 
+80,000 fatalities. How much slip.
+Using InSAR measurements and post-Wenchuan GPS observations, Qi et. al. 2011 found evidence of 2-6m of slip on a 20km deep detachment that extends 60km northwest underneath the Tibetan Plateau. 
+Our goal is to demonstrate that locking on this deep detachment is observable in the interseismic GPS field prior to the Wenchuan earthquake.
+Further, we will show how a deep detachment in the Longmen Shan shifts the main velocity gradient far to the northwest of the Beichuan fault trace.
+Hence, the detachment obscures the near-field geodetic shortening signature. 
+Accounting for the detachment, we present increased estimates of the shortening rate in the Longmen Shan.
+Fielding and .. (CITE) interpret high resolution satellite gravity observations to indicate that the Tibetan Plateau is being thrust over the Sichuan Basin.
+
+\section{Geodetic analysis of Longmen Shan shortening}
+Geodetic shortening across the Longmen Shan
+Show figure with the location of the profile, location of the GPS stations, beichuan fault trace, DEM background.
+GPS observations from Apel, Banerjee, Calais, Gan, Vigny, assembled into a single reference frame by Loveless and Meade.
+All observations were made prior to the Wenchuan earthquake.
+Although sparse, improvements in interseismic GPS resolution will not be available for a long time (many decades?) due to overprinting postseismic signals from Wenchuan.
+
+The debate over shortening in the Longmen Shan has centered around the clear lack of velocity gradient at the range front.
+Chen et. al. 2000 < 3mm/yr of shortening in the Longmen Shan.
+However, all their stations are within 50km of the range front. CHECK THIS
+Reference the plot of GPS velocities in the figure which shows the best fitting solution.
+General theme of previous interpretations: either ignored earthquake cycle effects or only looked at the near-field GPS.
+Which, if there is a deep detachment, are useless because they would indicate almost no shortening.
+Observations may also have been diminished because they were late in the earthquake cycle – just before Wenchuan.
+Interseismic locking leads to a smooth profile of geodetic velocities across a locked fault.
+Interpreting this profile is highly dependent on fault geometry.
+In the Longmen Shan, a deep detachment leads to a surface velocity gradient that is localized very far to the west of the range front.
+However, ignoring the detachment would indicate minimal shortening across the range front (see figures).
+We developed a 2D boundary element model.
+We assume isotropic elasticity.
+How much is worth saying about the BEM code?
+The standard seems to be almost nothing...
+I'm uncomfortable with that.
+Should I link to a copy of the code used?
+
+Our fault geometry is based on a community fault model of the Longmen Shan fault system (in development, shaw, plesch) that is based on seismic reflection imagery.
+Although not directly observed, based on structural considerations, they include the deep detachment.
+The location of this detachment is approximately the same as the Qi et.
+al.
+2011 detachment.
+We use the SRTM30\_Plus global DEM(becker, sandwell et.
+al.)  to provide a representative, smoothed, topographic profile across the range.
+I need to talk to John Shaw to get reasonable elastic moduli for the region.
+For topography, GPS, and fault geometry, we project the data onto the vertical plane of the cross section line (this should just go in a figure).
+We fit the GPS velocities with a constant shortening rate on the whole Beichuan + detachment. The best fit is ~8 mm/yr (+-!!). 
+I need to actually do some fitting to get uncertainties and best fit!
+
+\section{Implications for tectonics and earthquake hazard}
+A detachment underneath the Longmen Shan and Tibetan Plateau indicates that the Tibetan Plateau is thrust over the Sichuan Basin.
+I should read some of the previous literature on this question.
+Is there any evidence for Sichuan-like crust beneath easternmost Tibet?
+Hence, our detachment-based interpretation of the geodetic velocities supports a block-like interpretation for the tectonics of the eastern Tibetan Plateau.
+Lower crustal flow may also be able to explain the steep velocity gradient localized to the west of the range front.
+However, observations of postseismic afterslip on the detachment surface corroborate our
+
+Longriba fault zone?
+
+I need to calculate recurrence intervals and look up previous estimates for comparison. 
+Assuming 5m of slip in the Wenchuan earthquake, a 45 degree dipping fault, and 8mm of shortening: 8*sqrt(2) mm/yr slip deficit -> 5000/8*sqrt(2) = 441 years.
+Assuming that all 8 mm/yr of slip-deficit is concentrated on a single 45 degree dipping fault with the same surface area as the Wenchuan earthquake rupture, we calculate a recurrence interval of 450 years.
+However, this recurrence interval ignores the presence of other Longmen Shan faults and should be considered as an upper bound.
+The Beichuan and Pengguan faults are only two members of an imbricated thrust sequence.
+Interpretations of seismic reflection data (hubbard et.
+al) find multiple other thrusts including a detachment called the Range Front Thrust that extends 50km into the Sichuan Basin.
+This detachment lies directly under Chengdu, a city of 14 million.
+The slip-deficit distribution amongst these faults is unknown.
+Accounting for the subhorizontal Range Front Thrust, which has a large surface area, might seriously reduce our recurrence interval.
+
+150kms west of the Beichuan, there are 3 GPS stations with similar profile location. But, the velocities are significantly different. This indicates some lateral heterogeneity in the orogeny.
+
+The strain accumulating on the detachment may release coseismically or may be released in any number of slower ways. 
+An analysis without the detachment would observe almost no horizontal velocity gradient across the Longmen Shan range front.
+This emphasizes that geometrically accurate fault models can dramatically alter the interpretation of geodetic data.
+Importance of geometry in seismic hazard. Even including earthquake cycle effects, we would estimate almost zero shortening without the presence of the detachment. Raise the impact of detachments on t
+
+\subsection{A hazard lower bound}
+There are reasons to believe that the seismic hazard estimate presented here is a low estimate.
+First, coseismic and long term slip estimates include (CITE VARIOUS COSEISMIC + DENSMORE) a significant component of strike-slip motion. But, we only model horizontal shortening. 
+Second, our two-dimensional model assumes plane strain conditions. 
+The approximation will result in under-estimates of the shortening required to produce far-field velocities.
+This is because the assumption is equivalent to assuming the fault geometry extends infinitely far along strike, resulting in a fault surface much larger than in reality. 
+Third, the modeled range-front thrust is the steeply dipping Beichuan fault.
+The foreland Range Front thrust aor the shallow detachment beneath Chengdu (CITE HUBBARDSHAW) dip more shallowly, producing a greater total moment deficit.
+These first three deficiencies could be remedied through a more geometrically realistic three-dimensional model of the region. However, the data density may not justify such a complex model. 
+(MAYBE REMOVE) Also, kinematically consistent three-dimensional modelling of interseismic deformation is an outstanding issue. 
+Finally, the shortening rate estimate may be low due to being estimated late in the earthquake cycle (GOOD CITATION FOR THE GENERAL CONCEPT?).
+
+On the other hand, there are reasons to believe that the strain energy budget in the Longmen Shan region is not entirely released in major earthquakes.
+Critical-taper wedge theory results in very low friction coefficients for the shallowly dipping detachment in the foreland and the deep detachment (Hubbard,Shaw, Klinger + deep detachment here).
+Hence, these fault segments are more likely to release significant energy aseismically.
+Analogously, it has been suggested that the Himalayan range front consumes most of its energy budget aseismically (CITE!!!).
+
+\section{Conclusions}
+Previous interseismic shortening estimates suggested almost no shortening.
+This is at odds with the Mw 7.9 Wenchuan earthquake.
+We clarify this conflict by including earthquake cycle effects and a 20km deep detachment.
+Our revised interseismic shortening rate is ().
+This interpretation of the interseismic GPS velocities unifies the three phases of the Longmen Shan earthquake cycle.
+The Mw 7.9 Wenchuan rupture and deep postseismic afterslip suggest an active fold-and-thrust belt in the Longmen Shan orogen.
+Unlike previous interseismic coupling estimates (cite), our model is aligned with this view of the Longmen Shan.
+Our model suggest that Wenchuan-like events have much shorter recurrence intervals. Yikes!
 

### 2014-09-11 paper.tex
Test modifications.
Introduction edits.

@@
 \usepackage{graphicx}
+\usepackage[margin=1.0in]{geometry}
 
@@
 In contrast, we find that interseismic GPS data are consistent with a shortening rate of 6$\pm$2 mm/yr.
-These results suggest that the Longmen Shan is an active fold-and-thrust belt with Wenchuan style earthquake recurrence intervals of <600 years.
+These results suggest that the Longmen Shan is an active fold-and-thrust belt with Wenchuan style earthquake recurrence intervals of $<$600 years.
 
@@
 The 2008 $\mathrm{M_w}$ 7.9 Wenchuan earthquake ruptured 250km along strike on the Beichuan and Pengguan faults, resulting in approximately XXXXX fatalities (CITE STONE)
-Prior geodetic estimates suggest almost zero horizontal shortening (Chen 2000, Shen 2005, etc.), in contrast with the dominant thrust slip-sense during the Wenchuan rupture. 
-The lack of shortening is reconciled with the steep topography by appealing to a theory of lower crustal inflation whereby high topography can be formed without upper crustal shortening (CITE clark, royden, england, molnar, kirby, etc)
+Prior geodetic estimates suggest almost zero horizontal shortening (Chen 2000, Shen 2005, Meade 2007, Loveless and Meade 2011), in contrast with the dominant thrust slip-sense during the Wenchuan rupture.
+Others authors reconcile the lack of shortening with the steep topography by appealing a lower crustal inflation tectonic model for the eastern Tibetan Plateau (CITE clark, royden, burchfiel, kirby, etc).
 However, the magnitude and location of the Wenchuan earthquake suggest a more typical fold-and-thrust belt. 
-Using seismic reflection data, Hubbard and Shaw find almost 100\% shortening in the range-front 
-Following on these works, there is a growing body of literature supporting the interpretation of the Longmen Shan as a fold-and-thrust belt (CITE, CITE, CITE) as opposed to a product of viscous lower crustal inflation.
+Using seismic reflection data, Hubbard and Shaw 2009 find a well-developed fold-and-thrust fault system geometry with long-term shortenings up to 100\%. 
+A growing body of literature supports this interpretation, 
+Fielding and McKenzie 2012 interpret satellite gravity-field data to infer that mass of the Longmen Shan is supported by a flexed Sichuan Basin. 
+This fold-and-thrust structure may extend deep into the plateau as evidenced by seismic tomography and coseismic slip inversions (CITE ZHANG, QI, FIELDING). 
+CITE THE PAPERS EXAMINING THE LONGRIBA FAULT THAT PROPOSE A DETACHMENT
 
-Consistent with the fold-and-thrust mechanism
 However, it remains to be explained how slow present-day geodetic shortening estimates can be reconciled with the Wenchuan earthquake and the large structurally-observed shortening.
-Here, we present a model that explains the regional geodetic velocities as the effect of a complex interseismically locked fault geometry, including a 20km deep detachment as well as the range-front thrust Beichuan fault.
-This geometry is consistent with the structural interpretations of Hubbard,Shaw,Klinger, with coseismic and postseismic Wenchuan slip distributions that attribute significant motion to a deep detachment (CITE QI, FIELDING).
-Also, seismic observations of P/sP conversion in the crust are suggestive of a low velocity layer near 20km depth (CITE ZHANG ET. AL. CRUSTAL STRUCTURE). 
-Using this range-front thrust plus detachment geometry, we infer an average horizontal shortening.
+Prior estimates can be categorized into those that neglect earthquake cycle effects, analyzing the fault slip-deficit without no explicit model of the geometry or tectonics, (CHEN, SHEN) and those that use block models, which normally assume a simplified fault geometry.
+Here, we present a model that includes earthquake cycle effects to explain the regional geodetic velocities as the result of a complex interseismically locked fault geometry, including a 20km deep detachment and the range-front thrust Beichuan fault.
+This geometry is consistent with structural interpretations and with the geometry of Qi et. al 2012 and Fielding et. al 2013. 
+Using this range-front thrust plus detachment geometry, we infer an average horizontal shortening between the TIbetan Plateau.
 Next, we show how this geometry can drives the steep velocity gradients far to the northwest of the range front, thus explaining the absence of a range-front interseismic velocity gradient.
-Finally, we discuss the effects of Longmen Shan fault geometry on regional seismic hazard.
+Finally, we discuss the effects of Longmen Shan fault geometry on regional seismic hazard and tectonics.
 
-The Longmen Shan are a mountain range at boundary between the Tibetan Plateau and the Sichuan Basin.
-West of the Longmen Shan, in the eastern Tibetan plateau, surface GPS velocity vectors rotate from northeasterly to easterly or south-easterly.
-This rotation has been attributed to either the extrusion of crustal blocks or topographically driven viscous flow impacting a cold, elastically and viscously rigid, Sichuan basin.
-Because of its uniquely steep topographic gradient, lack of foreland basin, etc..., the Longmen Shan is a natural testing ground for these hypotheses.
-The steep topographic front of the Longmen Shan has been explained by both horizontal shortening (cite meade, hubbard/shaw) and lower crustal inflation (cite clark, royden, england, molnar, others). 
-The contrast between high topography and low shortening rates is a central point in this debate.
-Cite kirby
-The Mw 7.9 2008 Wenchuan earthquake ruptured 250km NE-SW along the Pengguan and Beichuan faults in the Longmenshan (cite xu 2008);e .
-Average slip was 4m 
-80,000 fatalities. How much slip.
-Using InSAR measurements and post-Wenchuan GPS observations, Qi et. al. 2011 found evidence of 2-6m of slip on a 20km deep detachment that extends 60km northwest underneath the Tibetan Plateau. 
-Our goal is to demonstrate that locking on this deep detachment is observable in the interseismic GPS field prior to the Wenchuan earthquake.
-Further, we will show how a deep detachment in the Longmen Shan shifts the main velocity gradient far to the northwest of the Beichuan fault trace.
-Hence, the detachment obscures the near-field geodetic shortening signature. 
-Accounting for the detachment, we present increased estimates of the shortening rate in the Longmen Shan.
-Fielding and .. (CITE) interpret high resolution satellite gravity observations to indicate that the Tibetan Plateau is being thrust over the Sichuan Basin.
+Maybe insert this information?
+--- Using InSAR measurements and post-Wenchuan GPS observations, Qi et.
+al.
+2011 found evidence of 2-6m of slip on a 20km deep detachment that extends 60km northwest underneath the Tibetan Plateau.
 

### 2014-09-11 paper.tex
git config file.

@@
 Prior geodetic estimates suggest almost zero horizontal shortening (Chen 2000, Shen 2005, Meade 2007, Loveless and Meade 2011), in contrast with the dominant thrust slip-sense during the Wenchuan rupture.
-Others authors reconcile the lack of shortening with the steep topography by appealing a lower crustal inflation tectonic model for the eastern Tibetan Plateau (CITE clark, royden, burchfiel, kirby, etc).
+Other authors reconcile the lack of shortening with the steep topography by appealing a lower crustal inflation tectonic model for the eastern Tibetan Plateau (CITE clark, royden, burchfiel, kirby, etc).
 However, the magnitude and location of the Wenchuan earthquake suggest a more typical fold-and-thrust belt. 
@@
 However, it remains to be explained how slow present-day geodetic shortening estimates can be reconciled with the Wenchuan earthquake and the large structurally-observed shortening.
-Prior estimates can be categorized into those that neglect earthquake cycle effects, analyzing the fault slip-deficit without no explicit model of the geometry or tectonics, (CHEN, SHEN) and those that use block models, which normally assume a simplified fault geometry.
+Prior estimates can be categorized into those that neglect earthquake cycle effects, analyzing the fault slip-deficit without no explicit model of the geometry or tectonics, (CHEN, SHEN) and those that use block models, which normally assume a simplified fault geometry (CITE MEADE 2007, BURCHFIEL 2008, LOVELESS AND MEADE 2011).
+Loveless and Meade report the highest shortening rate of 3.2 mm/yr.
 Here, we present a model that includes earthquake cycle effects to explain the regional geodetic velocities as the result of a complex interseismically locked fault geometry, including a 20km deep detachment and the range-front thrust Beichuan fault.
-This geometry is consistent with structural interpretations and with the geometry of Qi et. al 2012 and Fielding et. al 2013. 
-Using this range-front thrust plus detachment geometry, we infer an average horizontal shortening between the TIbetan Plateau.
+This geometry is consistent with structural interpretations and with the geometry of Qi et al 2012 and Fielding et al 2013. 
+Using this range-front thrust plus detachment geometry, we infer an average horizontal shortening between the TIbetan Plateau using interseismic GPS data collected prior to the Wenchuan earthquake.
 Next, we show how this geometry can drives the steep velocity gradients far to the northwest of the range front, thus explaining the absence of a range-front interseismic velocity gradient.
@@
 Maybe insert this information?
---- Using InSAR measurements and post-Wenchuan GPS observations, Qi et.
-al.
+--- Using InSAR measurements and post-Wenchuan GPS observations, Qi et al.
 2011 found evidence of 2-6m of slip on a 20km deep detachment that extends 60km northwest underneath the Tibetan Plateau.
@@
 \section{Geodetic analysis of Longmen Shan shortening}
-Geodetic shortening across the Longmen Shan
-Show figure with the location of the profile, location of the GPS stations, beichuan fault trace, DEM background.
-GPS observations from Apel, Banerjee, Calais, Gan, Vigny, assembled into a single reference frame by Loveless and Meade.
-All observations were made prior to the Wenchuan earthquake.
-Although sparse, improvements in interseismic GPS resolution will not be available for a long time (many decades?) due to overprinting postseismic signals from Wenchuan.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=7in]{figs/lms_map_with_faults.pdf}
+    \caption{A map of the Longmen Shan region with the basemap showing shaded relief.
+The focal mechanism plot is located at the epicenter of the Wenchuan earthquake, while the thicker red line shows the surface trace of rupture.
+The thin red lines show 1km depth contours of our model geometry.
+GPS velocity vectors are colored black if they are not included in our parameter estimation and red if they are included.
+The blue line shows the two-dimensional cross-section we study.
+Thin black lines indicate other significant faults in the region.}
+    \label{fig:regional_map}
+\end{figure}
+
+Our fault geometry is derived from a Community Fault Model for the Longmen Shan and Sichuan Basin region being actively developed (Plesch and Shaw, personal communication).
+This geometry is derived from structural interpretation of both seismic reflection surverys and surface geology (Hubbard \& Shaw).
+We make no modifications to the geometry except to take a fault-perpindicular cross-section.
+The cross-section and geometry with depth contours are shown in Figure \ref{fig:regional_map}.
+
+Because our model includes a deep detachment, we must include interseismic GPS data deep into the interior of the Tibetan Plateau.
+We study GPS observations collected prior to the Wenchuan earthquake by (CITES Apel, Banerjee, Calais, Gan, Vigny), assembled into a single reference frame by Loveless and Meade 2011.
+These velocities are shown in Figure \ref{fig:regional_map} along with the site of the Wenchuan earthquake and other regional faults.
+We include in our estimation velocities that are within the potential far-field influence of a deep detachment under the easternmost Tibetan Plateau and the Longmen Shan.
+We exclude velocities that are near major faults like the Xianshuihe fault or the Kunlun fault. 
+Later, we demonstrate that our results are reasonably robust to the specific data set that is studied.
+Addition of more recent observations could bias our results due to possible postseismic transient deformation.
+
+To analyze these data, we use a new quasistatic elastic boundary element code.
+An important feature of our code is the inclusion of accurate surface topography in steep mountain ranges or Earth curvature in large-scale problems.
+In addition, we can accurately reproduce half-space solutions to very high precision. 
+The steep topography of the Longmen Shan range front is included via the SRTM-30 dataset (CITE SANDWELL).
+
+As demonstrated by the leveling-off of the model predicted velocities, velocities deeper into the plateau are likely controlled by other fault systems.
 
 The debate over shortening in the Longmen Shan has centered around the clear lack of velocity gradient at the range front. 
-Chen et. al. 2000 < 3mm/yr of shortening in the Longmen Shan.
+Chen et al. 2000 < 3mm/yr of shortening in the Longmen Shan.
 However, all their stations are within 50km of the range front. CHECK THIS
@@
 However, ignoring the detachment would indicate minimal shortening across the range front (see figures).
-We developed a 2D boundary element model.
-We assume isotropic elasticity.
-How much is worth saying about the BEM code?
-The standard seems to be almost nothing...
-I'm uncomfortable with that.
-Should I link to a copy of the code used?
 
@@
 Although not directly observed, based on structural considerations, they include the deep detachment.
-The location of this detachment is approximately the same as the Qi et.
-al.
+The location of this detachment is approximately the same as the Qi et al.
 2011 detachment.
-We use the SRTM30\_Plus global DEM(becker, sandwell et.
-al.)  to provide a representative, smoothed, topographic profile across the range.
+We use the SRTM30\_Plus global DEM(becker, sandwell et al.)  to provide a representative, smoothed, topographic profile across the range.
 I need to talk to John Shaw to get reasonable elastic moduli for the region.
@@
 The Beichuan and Pengguan faults are only two members of an imbricated thrust sequence.
-Interpretations of seismic reflection data (hubbard et.
-al) find multiple other thrusts including a detachment called the Range Front Thrust that extends 50km into the Sichuan Basin.
+Interpretations of seismic reflection data (hubbard et al) find multiple other thrusts including a detachment called the Range Front Thrust that extends 50km into the Sichuan Basin.
 This detachment lies directly under Chengdu, a city of 14 million.
@@
 
+It is also important that the recurrence interval calculation assumed that the entire range-front was ruptured by the Wenchuan earthquake.
+However, the southern portion of the Longmen Shan did not rupture during the earthquake.
+
 \section{Conclusions}

### 2014-09-12 paper.tex
Deleted crazy number of line breaks in Abstract.
Going to add the second figure. Working on the results description.

@@
 \usepackage[margin=1.0in]{geometry}
+\usepackage[nomarkers,figuresonly]{endfloat}
 
@@
 The Longmen Shan range is located on the eastern margin of the Tibetan and rises 4000m over 30km (CHECK THESE NUMBERS ON GOOGLE EARTH).
-The 2008 $\mathrm{M_w}$ 7.9 Wenchuan earthquake ruptured 250km along strike on the Beichuan and Pengguan faults, resulting in approximately XXXXX fatalities (CITE STONE)
-Prior geodetic estimates suggest almost zero horizontal shortening (Chen 2000, Shen 2005, Meade 2007, Loveless and Meade 2011), in contrast with the dominant thrust slip-sense during the Wenchuan rupture.
+The 2008 $\mathrm{M_w}$ 7.9 Wenchuan earthquake ruptured 250km along strike on the Beichuan and Pengguan faults, resulting in approximately XXXXX fatalities (CITE STONE) Prior geodetic estimates suggest almost zero horizontal shortening (Chen 2000, Shen 2005, Meade 2007, Loveless and Meade 2011), in contrast with the dominant thrust slip-sense during the Wenchuan rupture.
 Other authors reconcile the lack of shortening with the steep topography by appealing a lower crustal inflation tectonic model for the eastern Tibetan Plateau (CITE clark, royden, burchfiel, kirby, etc).
@@
 Using seismic reflection data, Hubbard and Shaw 2009 find a well-developed fold-and-thrust fault system geometry with long-term shortenings up to 100\%.
-A growing body of literature supports this interpretation, 
-Fielding and McKenzie 2012 interpret satellite gravity-field data to infer that mass of the Longmen Shan is supported by a flexed Sichuan Basin. 
+A growing body of literature supports this interpretation, Fielding and McKenzie 2012 interpret satellite gravity-field data to infer that mass of the Longmen Shan is supported by a flexed Sichuan Basin.
 This fold-and-thrust structure may extend deep into the plateau as evidenced by seismic tomography and coseismic slip inversions (CITE ZHANG, QI, FIELDING).
@@
 Using this range-front thrust plus detachment geometry, we infer an average horizontal shortening between the TIbetan Plateau using interseismic GPS data collected prior to the Wenchuan earthquake.
-Next, we show how this geometry can drives the steep velocity gradients far to the northwest of the range front, thus explaining the absence of a range-front interseismic velocity gradient.
-Finally, we discuss the effects of Longmen Shan fault geometry on regional seismic hazard and tectonics.
+Next, we show how this geometry can drive the steep velocity gradients far to the northwest of the range front, thus explaining the absence of a range-front interseismic velocity gradient.
+Importantly, this model provides increased estimates of seismic hazard.
+This model proposes that strain accumulates deep in the interior of the Tibetan plateau while being released at the Longmen Shan range front, thus providing a unified theory of the Longmen Shan earthquake cycle.
 
@@
 
-Our fault geometry is derived from a Community Fault Model for the Longmen Shan and Sichuan Basin region being actively developed (Plesch and Shaw, personal communication).
+The fault geometry we consider is derived from a Community Fault Model for the Longmen Shan and Sichuan Basin region being actively developed (Plesch and Shaw, personal communication).
 This geometry is derived from structural interpretation of both seismic reflection surverys and surface geology (Hubbard \& Shaw).
@@
 
-Because our model includes a deep detachment, we must include interseismic GPS data deep into the interior of the Tibetan Plateau.
+Because this model includes a deep detachment, we must include interseismic GPS data deep into the interior of the Tibetan Plateau.
 We study GPS observations collected prior to the Wenchuan earthquake by (CITES Apel, Banerjee, Calais, Gan, Vigny), assembled into a single reference frame by Loveless and Meade 2011.
 These velocities are shown in Figure \ref{fig:regional_map} along with the site of the Wenchuan earthquake and other regional faults.
-We include in our estimation velocities that are within the potential far-field influence of a deep detachment under the easternmost Tibetan Plateau and the Longmen Shan.
+We include in the estimation velocities that are within the potential far-field influence of a deep detachment under the easternmost Tibetan Plateau.
 We exclude velocities that are near major faults like the Xianshuihe fault or the Kunlun fault.
@@
 
-To analyze these data, we use a new quasistatic elastic boundary element code.
-An important feature of our code is the inclusion of accurate surface topography in steep mountain ranges or Earth curvature in large-scale problems.
+To analyze these data, we use new quasistatic elastic boundary element software.
+An important feature of our tool is the inclusion of accurate surface topography in steep mountain ranges or Earth curvature in large-scale problems.
 In addition, we can accurately reproduce half-space solutions to very high precision.
 The steep topography of the Longmen Shan range front is included via the SRTM-30 dataset (CITE SANDWELL).
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.8\linewidth]{vels_geom_interior.pdf}
+    \caption{Vels_geom_interior}
+    \label{fig:vels_geom_interior}
+\end{figure}
+In
+
+A Beichuan-only fault geometry would predict very little deformation beyond 100km (CITE savage 1983), leading one to exclude velocities beyond this range.
+Both the observed and predicted profile-parallel velocities shown in Figure \ref{fig:big_stack} make clear that such an estimate would result in a near-zero shortening estimate.
 

### 2014-09-12 paper.tex
Added stacked comparison plot and caption.

@@
 \usepackage[nomarkers,figuresonly]{endfloat}
+\usepackage{setspace}
+\doublespacing
 
@@
 Importantly, this model provides increased estimates of seismic hazard.
-This model proposes that strain accumulates deep in the interior of the Tibetan plateau while being released at the Longmen Shan range front, thus providing a unified theory of the Longmen Shan earthquake cycle.
+This model proposes that strain accumulates up to 300km from the margin of the Tibetan plateau while being released at the Longmen Shan range front, thus providing a unified theory of the Longmen Shan earthquake cycle.
 
@@
     \centering
-    \includegraphics[width=7in]{figs/lms_map_with_faults.pdf}
+    \includegraphics[width=7in]{figs/lms_map_all.pdf}
     \caption{A map of the Longmen Shan region with the basemap showing shaded relief.
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{vels_geom_interior.pdf}
-    \caption{Vels_geom_interior}
-    \label{fig:vels_geom_interior}
+    \includegraphics[width=0.8\linewidth]{figs/stack_figure_all_details.pdf}
+    \caption{a) Observed profile-parallel horizontal observed velocities are shown in black with 68\% confidence errorbars.
+Model-predicted velocities are shown by the red line with the surrounding gray region showing a 68\% confidence interval.
+b) Model geometry with colors and width along the locked fault showing predicted slip-deficit rates.
+Because shortening is the horizontal component of slip-deficit, increasing dips near the surface lead to great slip-deficits.
+c) Predicted horizontal velocity as a function of depth.
+Note the steep velocity gradient and corresponding strain accumulation above the end of the detachment surface.}
+    \label{fig:big_stack}
 \end{figure}

### 2014-09-13 paper.tex
Added remaining figures. Some text.

@@
-\documentclass{article}
+\documentclass[12pt]{article}
 
@@
     \centering
-    \includegraphics[width=7in]{figs/lms_map_all.pdf}
+    \includegraphics{figs/lms_map_all.pdf}
     \caption{A map of the Longmen Shan region with the basemap showing shaded relief.
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{figs/stack_figure_all_details.pdf}
+    \includegraphics{figs/stack_figure_all_details.pdf}
     \caption{a) Observed profile-parallel horizontal observed velocities are shown in black with 68\% confidence errorbars.
@@
 b) Model geometry with colors and width along the locked fault showing predicted slip-deficit rates.
-Because shortening is the horizontal component of slip-deficit, increasing dips near the surface lead to great slip-deficits.
+Because shortening is the horizontal component of slip-deficit, increasing dips near the surface lead to larger slip-deficits.
 c) Predicted horizontal velocity as a function of depth.
@@
 \end{figure}
-In
 
@@
 Both the observed and predicted profile-parallel velocities shown in Figure \ref{fig:big_stack} make clear that such an estimate would result in a near-zero shortening estimate.
-
+A locked detachment, however, drives the predicted location of steep velocity gradient towards the northwest.
 As demonstrated by the leveling-off of the model predicted velocities, velocities deeper into the plateau are likely controlled by other fault systems.
@@
 In the Longmen Shan, a deep detachment leads to a surface velocity gradient that is localized very far to the west of the range front.
-However, ignoring the detachment would indicate minimal shortening across the range front (see figures).
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics{figs/depth_slip_contour.pdf}
+    \caption{}
+    \label{fig:distribution}
+\end{figure}
 
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics{figs/hazard_all_details.pdf}
+    \caption{Recurrence as a function of slip, assuming the best-fit slip rate for the Beichuan fault.
+The discrepancy between the Wenchuan-like recurrence and the paleoseismic estimate may be due to the presence of multiple active faults in the range-front.}
+    \label{fig:hazard}
+\end{figure}
+
 \subsection{A hazard lower bound}

### 2014-09-13 figs/crop_all.py

@@ -1,0 +1,10 @@
+from os import listdir, system
+from os.path import isfile, join
+
+directory = './'
+for f in listdir(directory):
+    if not isfile(join(directory, f)):
+        continue
+    if not f.lower().endswith('.pdf'):
+        continue
+    system('pdfcrop ' + f + ' ' + f)

### 2014-09-16 paper.tex
Substantial writing. Added many citations.

@@
 \usepackage{setspace}
+\usepackage{natbib}
 \doublespacing
@@
 \section{Introduction}
-The Longmen Shan range is located on the eastern margin of the Tibetan and rises 4000m over 30km (CHECK THESE NUMBERS ON GOOGLE EARTH).
-The 2008 $\mathrm{M_w}$ 7.9 Wenchuan earthquake ruptured 250km along strike on the Beichuan and Pengguan faults, resulting in approximately XXXXX fatalities (CITE STONE) Prior geodetic estimates suggest almost zero horizontal shortening (Chen 2000, Shen 2005, Meade 2007, Loveless and Meade 2011), in contrast with the dominant thrust slip-sense during the Wenchuan rupture.
-Other authors reconcile the lack of shortening with the steep topography by appealing a lower crustal inflation tectonic model for the eastern Tibetan Plateau (CITE clark, royden, burchfiel, kirby, etc).
+The Longmen Shan range is located on the eastern margin of the Tibetan and rises 4000m over 30km.
+The 2008 $\mathrm{M_w}$ 7.9 Wenchuan earthquake ruptured 250km along strike on the Beichuan and Pengguan faults, resulting in approximately 70000 fatalities.
+Prior geodetic estimates suggest almost zero horizontal shortening \citep{king97, chen00, shen05, Meade07c, Loveless2011}, in contrast with the dominant thrust slip-sense during the Wenchuan rupture.
+Other authors reconcile the lack of shortening with the steep topography by appealing a lower crustal inflation tectonic model for the eastern Tibetan Plateau \citep{royden97, bird91, Burchfiel2008a}.
 However, the magnitude and location of the Wenchuan earthquake suggest a more typical fold-and-thrust belt.
-Using seismic reflection data, Hubbard and Shaw 2009 find a well-developed fold-and-thrust fault system geometry with long-term shortenings up to 100\%.
-A growing body of literature supports this interpretation, Fielding and McKenzie 2012 interpret satellite gravity-field data to infer that mass of the Longmen Shan is supported by a flexed Sichuan Basin.
-This fold-and-thrust structure may extend deep into the plateau as evidenced by seismic tomography and coseismic slip inversions (CITE ZHANG, QI, FIELDING).
-CITE THE PAPERS EXAMINING THE LONGRIBA FAULT THAT PROPOSE A DETACHMENT
+Using seismic reflection data, \citet{hubbard09} find a well-developed fold-and-thrust fault system geometry with long-term shortenings up to 100\%.
+A growing body of literature supports this interpretation.
+\citet{Fielding2012a} interpret satellite gravity-field data to infer that mass of the Longmen Shan is supported by a flexed Sichuan Basin.
+This fold-and-thrust structure may extend deep into the plateau as suggested by seismic tomography, reflection, magnetotelluric data and studies of the Longriba fault zone 200km northwest of the Longmen Shan \citep{Xu2008, Zhang2009, Zhao2012, Ren2013, Guo2013}.
+(LOTS TO EXTEND HERE, SPACE?)
 
 However, it remains to be explained how slow present-day geodetic shortening estimates can be reconciled with the Wenchuan earthquake and the large structurally-observed shortening.
-Prior estimates can be categorized into those that neglect earthquake cycle effects, analyzing the fault slip-deficit without no explicit model of the geometry or tectonics, (CHEN, SHEN) and those that use block models, which normally assume a simplified fault geometry (CITE MEADE 2007, BURCHFIEL 2008, LOVELESS AND MEADE 2011).
-Loveless and Meade report the highest shortening rate of 3.2 mm/yr.
+Prior estimates can be categorized into those that neglect earthquake cycle effects, analyzing the fault slip-deficit without no explicit model of the geometry or tectonics, \citep{chen00, shen05} and those that use block models, which normally assume a simplified fault geometry \citep{Meade07c, Loveless2011, Burchfiel2008a}.
+\citet{Loveless2011} report the highest shortening rate of 3.2 mm/yr.
 Here, we present a model that includes earthquake cycle effects to explain the regional geodetic velocities as the result of a complex interseismically locked fault geometry, including a 20km deep detachment and the range-front thrust Beichuan fault.
-This geometry is consistent with structural interpretations and with the geometry of Qi et al 2012 and Fielding et al 2013.
+This geometry is consistent with the inference of 2-6m of slip of a 20km deep detachment that extends 90km northwest using InSAR measurements and post-Wenchuan GPS observations \citep{Qi2011}.
+\citet{Fielding2013b} also include a detachment fault plane in their joint geodetic-teleseismic coseismic slip inversion.
+Further, this geometry is consistent with structural interpretations \citep{Hubbard2010, Li2010}.
 Using this range-front thrust plus detachment geometry, we infer an average horizontal shortening between the TIbetan Plateau using interseismic GPS data collected prior to the Wenchuan earthquake.
@@
 This model proposes that strain accumulates up to 300km from the margin of the Tibetan plateau while being released at the Longmen Shan range front, thus providing a unified theory of the Longmen Shan earthquake cycle.
-
-Maybe insert this information?
---- Using InSAR measurements and post-Wenchuan GPS observations, Qi et al.
-2011 found evidence of 2-6m of slip on a 20km deep detachment that extends 60km northwest underneath the Tibetan Plateau.
 
@@
 The fault geometry we consider is derived from a Community Fault Model for the Longmen Shan and Sichuan Basin region being actively developed (Plesch and Shaw, personal communication).
-This geometry is derived from structural interpretation of both seismic reflection surverys and surface geology (Hubbard \& Shaw).
-We make no modifications to the geometry except to take a fault-perpindicular cross-section.
+This geometry is derived from structural interpretation of both seismic reflection surverys and surface geology \citep{Hubbard2010}.
+We make no modifications to this geometry except to take a fault-perpindicular cross-section.
+The depth and dip of the detachment are similar to the best-fitting surface found by \citet{Qi2011}.
 The cross-section and geometry with depth contours are shown in Figure \ref{fig:regional_map}.
@@
 Because this model includes a deep detachment, we must include interseismic GPS data deep into the interior of the Tibetan Plateau.
-We study GPS observations collected prior to the Wenchuan earthquake by (CITES Apel, Banerjee, Calais, Gan, Vigny), assembled into a single reference frame by Loveless and Meade 2011.
-These velocities are shown in Figure \ref{fig:regional_map} along with the site of the Wenchuan earthquake and other regional faults.
+We study GPS observations collected prior to the Wenchuan earthquake \citep{apel06,Banerjee2008,calais06, gan07, vigny03}, assembled into a single reference frame by \citet{Loveless2011}.
+These velocities are shown in Figure \ref{fig:regional_map} along with the site of the Wenchuan earthquake and other regional faults \citep{Taylor09}.
 We include in the estimation velocities that are within the potential far-field influence of a deep detachment under the easternmost Tibetan Plateau.
@@
 In addition, we can accurately reproduce half-space solutions to very high precision.
-The steep topography of the Longmen Shan range front is included via the SRTM-30 dataset (CITE SANDWELL).
+The steep topography of the Longmen Shan range front is included via the SRTM30\_PLUS digital elevation dataset \citep{Becker2009}.
 
@@
 
-A Beichuan-only fault geometry would predict very little deformation beyond 100km (CITE savage 1983), leading one to exclude velocities beyond this range.
-Both the observed and predicted profile-parallel velocities shown in Figure \ref{fig:big_stack} make clear that such an estimate would result in a near-zero shortening estimate.
-A locked detachment, however, drives the predicted location of steep velocity gradient towards the northwest.
-As demonstrated by the leveling-off of the model predicted velocities, velocities deeper into the plateau are likely controlled by other fault systems.
+The debate over shortening in the Longmen Shan has centered around the clear lack of velocity gradient at the range front.
+However, this discussion has neglected the interplay between earthquake cycle effects and fault geometry.
+A Beichuan-only fault geometry would predict very little deformation beyond 100km \citep{savage83}, leading to the exclusion of velocities beyond this range.
+Both the observed and predicted profile-parallel velocities shown in Figure \ref{fig:big_stack}a make clear that such an estimate would result in a near-zero shortening estimate.
 
-The debate over shortening in the Longmen Shan has centered around the clear lack of velocity gradient at the range front. 
-Chen et al. 2000 < 3mm/yr of shortening in the Longmen Shan.
-However, all their stations are within 50km of the range front. CHECK THIS
-Reference the plot of GPS velocities in the figure which shows the best fitting solution.
-General theme of previous interpretations: either ignored earthquake cycle effects or only looked at the near-field GPS.
-Which, if there is a deep detachment, are useless because they would indicate almost no shortening.
-Observations may also have been diminished because they were late in the earthquake cycle – just before Wenchuan.
-Interseismic locking leads to a smooth profile of geodetic velocities across a locked fault.
-Interpreting this profile is highly dependent on fault geometry.
-In the Longmen Shan, a deep detachment leads to a surface velocity gradient that is localized very far to the west of the range front.
+A locked detachment, however, drives the predicted location of steep velocity gradient far to the northwest as shown by Figure \ref{fig:big_stack}a.
+The best-fit shortening rate for the Longmen Shan region is 5.7 $\pm$ 1.5 mm/yr.
+For discussion of seismic hazard, shortening must be translated to fault-parallel slip-deficit.
+Figure \ref{fig:big_stack}b shows the increase in slip-deficit from ~6.0 mm/yr up to 9.5 mm/yr at the surface.
+Figure \ref{fig:big_stack}c shows the horizontal velocities as a function of depth, demonstrating that the velocity gradients derive from the detachment tip.
+The smoothing behavior of an elastic Earth prevents distinguishing between a sharp drop in slip-deficit or a gradual change over many kilometers.
+These results demonstrate the mechanism by which it is possible for interseismic strain accumulation to occur far from the range-front structures on which it is released.
+
+The primary uncertainty in our shortening estimate is the lack of good constraints on the detachment geometry in the hinterland.
+However, assuming the geometry is correct, two further forms of error are evident.
+First, the uncertainty in true interseismic velocities is propagated through to the estimated shortening, resulting in the gray region of uncertainty in Figure \ref{fig:big_stack}a.
+Further, the sparsity of observations suggests that certain observations may undue influence on the shortening estimate.
+Excluding the northwest-most observation decreases the best-fit shortening to 3.9 $\pm$ 1.9 mm/yr.
+The least squares uncertainty is increased (from 1.5 mm/yr to 1.9 mm/yr) because the northwestern portion of velocities are poorly constrained once the observation is excluded.
+
+Figure \ref{fig:distribution} shows a more complete analysis of the effect of case resampling the observations on the slip-deficit rate.
+Because there only $2^{20} = 1.048 $ million different possible subsets of the observations, we perform a complete case resampling bootstrap.
+Subsequently, we sample jointly from this distribution and the distribution of surface-area weighted fault depths.
+This analysis demonstrates that the majority of the slip-deficit-depth probability landscape lies at slip-deficit rate between 4 and 6 mm/yr while certain parts of the geometry (the range-front) have much faster slip-deficit rates.
 
@@
     \includegraphics{figs/depth_slip_contour.pdf}
-    \caption{}
+    \caption{The depth-slip-deficit probability landscape.
+The upper histogram is the projection of the contour plot onto a depth histogram, demonstrating that the majority of the fault surface area lies on the deep detachment.
+The right histogram is the projection of the contour plot onto a slip-rate histogram showing that the majority of fault surface area is slipping at rates near 6 mm/yr.}
     \label{fig:distribution}
@@
 
-Our fault geometry is based on a community fault model of the Longmen Shan fault system (in development, shaw, plesch) that is based on seismic reflection imagery.
-Although not directly observed, based on structural considerations, they include the deep detachment.
-The location of this detachment is approximately the same as the Qi et al.
-2011 detachment.
-We use the SRTM30\_Plus global DEM(becker, sandwell et al.)  to provide a representative, smoothed, topographic profile across the range.
-I need to talk to John Shaw to get reasonable elastic moduli for the region.
-For topography, GPS, and fault geometry, we project the data onto the vertical plane of the cross section line (this should just go in a figure).
-We fit the GPS velocities with a constant shortening rate on the whole Beichuan + detachment. The best fit is ~8 mm/yr (+-!!). 
-I need to actually do some fitting to get uncertainties and best fit!
+\section{Implications for the Longmen Shan earthquake cycle}
+The striking feature of Figure \ref{fig:big_stack}a is the lack of a range-front geodetic signature of shortening.
 
-\section{Implications for tectonics and earthquake hazard}
-A detachment underneath the Longmen Shan and Tibetan Plateau indicates that the Tibetan Plateau is thrust over the Sichuan Basin.
-I should read some of the previous literature on this question.
-Is there any evidence for Sichuan-like crust beneath easternmost Tibet?
-Hence, our detachment-based interpretation of the geodetic velocities supports a block-like interpretation for the tectonics of the eastern Tibetan Plateau.
-Lower crustal flow may also be able to explain the steep velocity gradient localized to the west of the range front.
-However, observations of postseismic afterslip on the detachment surface corroborate our
+Model without topography do not have a local minimum northwest of the Beichuan, indicating the relevance of topography in crustal deformation.
 
-Longriba fault zone?
+This far-field deformation is consistent with vertical leveling data that suggests a vertical velocity profile that increases from the Longmen Shan range front until peaking 50km southeast of the Longriba fault zone \citep{Hao2014}.
+The Longriba fault zone has the wrong slip-sense (up to the south-east, \citep{Ren2013}) to accomodate such vertical deformation.
+No other structure is nearby.
 
-I need to calculate recurrence intervals and look up previous estimates for comparison. 
-Assuming 5m of slip in the Wenchuan earthquake, a 45 degree dipping fault, and 8mm of shortening: 8*sqrt(2) mm/yr slip deficit -> 5000/8*sqrt(2) = 441 years.
-Assuming that all 8 mm/yr of slip-deficit is concentrated on a single 45 degree dipping fault with the same surface area as the Wenchuan earthquake rupture, we calculate a recurrence interval of 450 years.
-However, this recurrence interval ignores the presence of other Longmen Shan faults and should be considered as an upper bound.
-The Beichuan and Pengguan faults are only two members of an imbricated thrust sequence.
-Interpretations of seismic reflection data (hubbard et al) find multiple other thrusts including a detachment called the Range Front Thrust that extends 50km into the Sichuan Basin.
-This detachment lies directly under Chengdu, a city of 14 million.
-The slip-deficit distribution amongst these faults is unknown.
-Accounting for the subhorizontal Range Front Thrust, which has a large surface area, might seriously reduce our recurrence interval.
+Figure \ref{fig:hazard} shows a summary of recurrence intervals as a function of total event slip.
+Our shortening estimate suggests a recurrence interval for Wenchuan-like ruptures of approximately 600 years.
+Recurrence for smaller events is much more frequent.
+This contrasts with the average recurrence of ~2000 years determined by \citep{Ran2010} for the Beichuan fault.
+Our estimate would predict $M_w = $9.0 ruptures on that recurrence, seemingly unlikely given the total slip and fault length required.
+However, our estimate is an system-wide recurrence while that paleoseismic estimate is limited by only studying one of multiple imbricated thrusts.
 
-150kms west of the Beichuan, there are 3 GPS stations with similar profile location. But, the velocities are significantly different. This indicates some lateral heterogeneity in the orogeny.
+There are reasons to believe that the seismic hazard estimate presented here is a low estimate.
+First, coseismic and long term slip estimates include (CITE VARIOUS COSEISMIC + DENSMORE) a significant component of strike-slip motion.
+But, we only model horizontal shortening.
+Second, our two-dimensional model assumes plane strain conditions.
+The approximation will result in under-estimates of the shortening required to produce far-field velocities.
+This is because the assumption is equivalent to assuming the fault geometry extends infinitely far along strike, resulting in a fault surface much larger than in reality.
+Third, the modeled thrust is the steeply dipping Beichuan fault.
+The foreland Range Front thrust or the shallow detachment beneath Chengdu dip more shallowly \citep{Hubbard2010}, producing a greater total moment deficit.
+Finally, the shortening rate estimate may be low due to being estimated late in the earthquake cycle \citep{savage00}.
+It is also important that the recurrence interval calculation assumed that the entire range-front was ruptured by the Wenchuan earthquake whereas the southern portion of the Longmen Shan did not rupture during the earthquake and is thought to similarly active to the portions that did rupture \citep{Li2010}.
 
-The strain accumulating on the detachment may release coseismically or may be released in any number of slower ways. 
-An analysis without the detachment would observe almost no horizontal velocity gradient across the Longmen Shan range front.
-This emphasizes that geometrically accurate fault models can dramatically alter the interpretation of geodetic data.
-Importance of geometry in seismic hazard. Even including earthquake cycle effects, we would estimate almost zero shortening without the presence of the detachment. Raise the impact of detachments on t
+Approximately 150 kilometers west of the Beichuan fault, there are 3 GPS stations with similar profile location.
+But, the velocities are significantly different.
+This indicates some lateral heterogeneity not represented by our two-dimensional modelling.
 
@@
     \caption{Recurrence as a function of slip, assuming the best-fit slip rate for the Beichuan fault.
+The gray region indicates the uncertainty implied by the error in our shortening estimate.
 The discrepancy between the Wenchuan-like recurrence and the paleoseismic estimate may be due to the presence of multiple active faults in the range-front.}
@@
 
-\subsection{A hazard lower bound}
-There are reasons to believe that the seismic hazard estimate presented here is a low estimate.
-First, coseismic and long term slip estimates include (CITE VARIOUS COSEISMIC + DENSMORE) a significant component of strike-slip motion. But, we only model horizontal shortening. 
-Second, our two-dimensional model assumes plane strain conditions. 
-The approximation will result in under-estimates of the shortening required to produce far-field velocities.
-This is because the assumption is equivalent to assuming the fault geometry extends infinitely far along strike, resulting in a fault surface much larger than in reality. 
-Third, the modeled range-front thrust is the steeply dipping Beichuan fault.
-The foreland Range Front thrust aor the shallow detachment beneath Chengdu (CITE HUBBARDSHAW) dip more shallowly, producing a greater total moment deficit.
-These first three deficiencies could be remedied through a more geometrically realistic three-dimensional model of the region. However, the data density may not justify such a complex model. 
-(MAYBE REMOVE) Also, kinematically consistent three-dimensional modelling of interseismic deformation is an outstanding issue. 
-Finally, the shortening rate estimate may be low due to being estimated late in the earthquake cycle (GOOD CITATION FOR THE GENERAL CONCEPT?).
+Previous near-zero shortening estimates in the Longmen Shan are at odds with the large-moment of the Wenchuan earthquake, very steep topography, structural interpretations indicate fold-and-thrust geometry, fast erosion rates.
+We clarify this conflict by including earthquake cycle effects and a 20km deep detachment.
+Our revised interseismic shortening rate is 5.7 $\pm$ 1.5 mm/yr.
+This interpretation unifies models of coseismic (Wenchuan), postseismic and interseismic  based on geodetic data.
+Our model suggest that Wenchuan-like events have much shorter recurrence intervals.
 
-On the other hand, there are reasons to believe that the strain energy budget in the Longmen Shan region is not entirely released in major earthquakes.
-Critical-taper wedge theory results in very low friction coefficients for the shallowly dipping detachment in the foreland and the deep detachment (Hubbard,Shaw, Klinger + deep detachment here).
-Hence, these fault segments are more likely to release significant energy aseismically.
-Analogously, it has been suggested that the Himalayan range front consumes most of its energy budget aseismically (CITE!!!).
-
-It is also important that the recurrence interval calculation assumed that the entire range-front was ruptured by the Wenchuan earthquake.
-However, the southern portion of the Longmen Shan did not rupture during the earthquake.
-
-\section{Conclusions}
-Previous interseismic shortening estimates suggested almost no shortening.
-This is at odds with the Mw 7.9 Wenchuan earthquake.
-We clarify this conflict by including earthquake cycle effects and a 20km deep detachment.
-Our revised interseismic shortening rate is ().
-This interpretation of the interseismic GPS velocities unifies the three phases of the Longmen Shan earthquake cycle.
-The Mw 7.9 Wenchuan rupture and deep postseismic afterslip suggest an active fold-and-thrust belt in the Longmen Shan orogen.
-Unlike previous interseismic coupling estimates (cite), our model is aligned with this view of the Longmen Shan.
-Our model suggest that Wenchuan-like events have much shorter recurrence intervals. Yikes!
-
+\bibliographystyle{plainnat}
+\bibliography{biblio,library}
 \end{document}

### 2014-09-17 paper.tex
Added to discussion section.

@@
 Other authors reconcile the lack of shortening with the steep topography by appealing a lower crustal inflation tectonic model for the eastern Tibetan Plateau \citep{royden97, bird91, Burchfiel2008a}.
-However, the magnitude and location of the Wenchuan earthquake suggest a more typical fold-and-thrust belt.
-Using seismic reflection data, \citet{hubbard09} find a well-developed fold-and-thrust fault system geometry with long-term shortenings up to 100\%.
-A growing body of literature supports this interpretation.
+However, the magnitude and location of the Wenchuan earthquake suggest a rapidly shortening fold-and-thrust belt.
+From seismic reflection data, \citet{hubbard09} find a well-developed fold-and-thrust fault system geometry with long-term shortenings up to 100\%.
 \citet{Fielding2012a} interpret satellite gravity-field data to infer that mass of the Longmen Shan is supported by a flexed Sichuan Basin.
-This fold-and-thrust structure may extend deep into the plateau as suggested by seismic tomography, reflection, magnetotelluric data and studies of the Longriba fault zone 200km northwest of the Longmen Shan \citep{Xu2008, Zhang2009, Zhao2012, Ren2013, Guo2013}.
-(LOTS TO EXTEND HERE, SPACE?)
+Seismic tomography, seismic reflection and magnetotelluric data have been used to suggest that this fold-and-thrust structure extends deep into the plateau as a deep detachment \citep{Zhang2009, Zhao2012, Guo2013}.
+Erosion rates are The Longriba fault zone, 200km northwest of the Longmen Shan, is suggested to also sole into this detachment \citep{Xu2008, Ren2013}.
 
@@
 \section{Implications for the Longmen Shan earthquake cycle}
-The striking feature of Figure \ref{fig:big_stack}a is the lack of a range-front geodetic signature of shortening.
+The striking feature of Figure \ref{fig:big_stack}a is the lack of a range-front geodetic signature despite the significant shortening rate.
+Locking on the deep detachment obscures the signal of deformation on the range-front faults.
+According to this model, interseismically, the eastern Tibetan Plateau moves cohesively with the Sichuan Basin, whereas coseismic slip is concentrated at the boundary between these two tectonic blocks.
+Traditionally, the velocity gradient we attribute to a deep detachment has been attributed to broad viscous strain across the eastern Tibetan Plateau (CITE).
 
-Model without topography do not have a local minimum northwest of the Beichuan, indicating the relevance of topography in crustal deformation.
+Models without surface topography do not have a local minimum northwest of the Beichuan, indicating the relevance of topography in crustal deformation.
+This effect may be more pronounced with a shallower fault dip like in many subduction zones.
 
-This far-field deformation is consistent with vertical leveling data that suggests a vertical velocity profile that increases from the Longmen Shan range front until peaking 50km southeast of the Longriba fault zone \citep{Hao2014}.
+Our model of far-field deformation is consistent with a similar detachment model of leveling and GPS-based uplift rates \citep{Hao2014}.
+The vertical velocities increase from the Longmen Shan range front until peaking 50km southeast of the Longriba fault zone at approximately 3.5 mm/yr of uplift.
 The Longriba fault zone has the wrong slip-sense (up to the south-east, \citep{Ren2013}) to accomodate such vertical deformation.
-No other structure is nearby.
+No other identified structures are nearby.
+
+\citet{kirby07} study an eastward decrease in interseismic velocities across the Kunlun fault.
+They conclude that the decline in slip-rate must be attributed to internal deformation.
+However, the same feature could be explained with slip being accomodated by a wide eastern Tibetan fold-and-thrust system, a suggestion supported by the continuation of shortening north from the Longmen Shan onto the Huya \citep{kirby00} and Min Jiang faults \citep{Chen1994}.
+
+We have not addressed when the accumulated slip-deficit on the deep detachment is released.
+Postseismic slip incorporated in the model of \cite{Qi2011} shows up to 5 meters of slip on a deep detachment.
+Our data analysis depends on the detachment being locked during the period of data collection (WHEN?).
+Aseismic slip may nonetheless be the predominant mode of strain release given the short observed portion of the earthquake cycle.
+Seismicity is approximately limited to upper 20km of crust in the range-front \citep{Li2010}.
+On the other hand, the surface area and shortening rate of the detachment could sustain very large earthquakes if slip is released coseismically.
+
+Using steady-state critical taper wedge theory, we estimate a basal coefficient of friction between 0.02 and 0.30 for the detachment.
+The mean parameter estimate of 0.05 is similar to estimates of friction on the low-angle detachment system beneath the Longmen Shan foreland basin \citep{Hubbard2010}.
 
@@
 This contrasts with the average recurrence of ~2000 years determined by \citep{Ran2010} for the Beichuan fault.
-Our estimate would predict $M_w = $9.0 ruptures on that recurrence, seemingly unlikely given the total slip and fault length required.
-However, our estimate is an system-wide recurrence while that paleoseismic estimate is limited by only studying one of multiple imbricated thrusts.
+Our estimate would predict $M_w = $9.0 ruptures for that recurrence, seemingly unlikely given the total slip and fault length required.
+However, our estimate is an system-wide recurrence while paleoseismic estimates have been limited by only studying one of multiple imbricated thrusts.
 
 There are reasons to believe that the seismic hazard estimate presented here is a low estimate.
-First, coseismic and long term slip estimates include (CITE VARIOUS COSEISMIC + DENSMORE) a significant component of strike-slip motion.
+First, coseismic and long term slip estimates include a significant component of strike-slip motion\citep{Shen2009, Qi2011, Densmore2007} .
 But, we only model horizontal shortening.
@@
 Our revised interseismic shortening rate is 5.7 $\pm$ 1.5 mm/yr.
-This interpretation unifies models of coseismic (Wenchuan), postseismic and interseismic  based on geodetic data.
-Our model suggest that Wenchuan-like events have much shorter recurrence intervals.
+The model shows that interseismic strain accumulation may be spatially disjoint with coseismic strain release, emphasizing that accurate fault geometries are critical in the analysis of geodetic data.
+This interpretation unifies geodetic models of coseismic (Wenchuan), postseismic and interseismic crustal deformation in the Longmen Shan.
+Finally, our model suggest that Wenchuan-like events have much shorter recurrence intervals.
 

### 2014-09-17 paper.tex
Significant additions to all sections.

@@
 Previous analyses which neglected the detachment and earthquake cycle effects have found shortening rates near zero.
-In contrast, we find that interseismic GPS data are consistent with a shortening rate of 6$\pm$2 mm/yr.
+In contrast, we find that interseismic GPS data are consistent with a shortening rate of 5.7$\pm$1.5 mm/yr.
 These results suggest that the Longmen Shan is an active fold-and-thrust belt with Wenchuan style earthquake recurrence intervals of $<$600 years.
@@
 The Longmen Shan range is located on the eastern margin of the Tibetan and rises 4000m over 30km.
-The 2008 $\mathrm{M_w}$ 7.9 Wenchuan earthquake ruptured 250km along strike on the Beichuan and Pengguan faults, resulting in approximately 70000 fatalities.
+The 2008 $M_{\textrm{w}}$ 7.9 Wenchuan earthquake ruptured 250km along strike on the Beichuan and Pengguan faults, resulting in approximately 70000 fatalities.
 Prior geodetic estimates suggest almost zero horizontal shortening \citep{king97, chen00, shen05, Meade07c, Loveless2011}, in contrast with the dominant thrust slip-sense during the Wenchuan rupture.
@@
 Seismic tomography, seismic reflection and magnetotelluric data have been used to suggest that this fold-and-thrust structure extends deep into the plateau as a deep detachment \citep{Zhang2009, Zhao2012, Guo2013}.
-Erosion rates are The Longriba fault zone, 200km northwest of the Longmen Shan, is suggested to also sole into this detachment \citep{Xu2008, Ren2013}.
+The Longriba fault zone, 200km northwest of the Longmen Shan, is suggested to also sole into this detachment \citep{Xu2008, Ren2013}.
 
@@
 The cross-section and geometry with depth contours are shown in Figure \ref{fig:regional_map}.
+The fault geometry consists of two main segments.
+The range-front Beichuan fault dips steeply at ~60$^{\circ}$ at its surface trace shallowing to almost horizontal where it joins the detachment surface at depth.
+The detachment dips at ~1.5$^{\circ}$ from an initial depth of 20km for 110km to the northwest at its deepest point with a depth of 23km.
 
@@
 
-To analyze these data, we use new quasistatic elastic boundary element software.
-An important feature of our tool is the inclusion of accurate surface topography in steep mountain ranges or Earth curvature in large-scale problems.
+To analyze these data, we use a quasistatic elastic boundary element software under active development.
+An important feature of our tool is the inclusion of accurate surface topography in steep mountainous regions or Earth curvature in large-scale problems.
 In addition, we can accurately reproduce half-space solutions to very high precision.
 The steep topography of the Longmen Shan range front is included via the SRTM30\_PLUS digital elevation dataset \citep{Becker2009}.
+To maintain interseismic kinematic consistency between the Sichuan Basin and Tibetan Plateau, we use a 2D block formulation.
+The average interseismic slip-deficit on the fault geometry is subtracted from block motion to calculate the interseismic surface velocity profile.
 
@@
     \includegraphics{figs/stack_figure_all_details.pdf}
-    \caption{a) Observed profile-parallel horizontal observed velocities are shown in black with 68\% confidence errorbars.
+    \caption{Axes are not one to one.
+a) Observed profile-parallel horizontal observed velocities are shown in black with 68\% confidence errorbars.
 Model-predicted velocities are shown by the red line with the surrounding gray region showing a 68\% confidence interval.
@@
 
-The debate over shortening in the Longmen Shan has centered around the clear lack of velocity gradient at the range front.
+The debate over shortening in the Longmen Shan has centered around the clear lack of an interseismic velocity gradient at the range front.
 However, this discussion has neglected the interplay between earthquake cycle effects and fault geometry.
@@
 A locked detachment, however, drives the predicted location of steep velocity gradient far to the northwest as shown by Figure \ref{fig:big_stack}a.
-The best-fit shortening rate for the Longmen Shan region is 5.7 $\pm$ 1.5 mm/yr.
+The best-fit shortening rate for the Longmen Shan region is 5.7 $\pm$ 1.5 mm/yr, a substantial increase over previous estimates.
 For discussion of seismic hazard, shortening must be translated to fault-parallel slip-deficit.
@@
 The smoothing behavior of an elastic Earth prevents distinguishing between a sharp drop in slip-deficit or a gradual change over many kilometers.
-These results demonstrate the mechanism by which it is possible for interseismic strain accumulation to occur far from the range-front structures on which it is released.
+These results demonstrate a mechanism by which it is possible for interseismic strain accumulation to occur far from the range-front structures on which it is released.
 
@@
 The least squares uncertainty is increased (from 1.5 mm/yr to 1.9 mm/yr) because the northwestern portion of velocities are poorly constrained once the observation is excluded.
+This northwestern-most observation is adjacent to the Longriba fault, suggesting that it should not be included.
+However, the Longriba fault is primarily a dextral strike-slip fault \citep{Ren2013}, so it's influence should be predominantly orthogonal to the profile parallel velocities we examine here.
 
@@
 This analysis demonstrates that the majority of the slip-deficit-depth probability landscape lies at slip-deficit rate between 4 and 6 mm/yr while certain parts of the geometry (the range-front) have much faster slip-deficit rates.
+In Figure \ref{fig:distribution}, there is a non-negligible probability assigned to slip-rates above 10 mm/yr in the steepest, near-surface portion of the fault system.
+This is consistent with large near-surface coseismic Wenchuan slip \citep{Shen2009}.
 
@@
 Locking on the deep detachment obscures the signal of deformation on the range-front faults.
-According to this model, interseismically, the eastern Tibetan Plateau moves cohesively with the Sichuan Basin, whereas coseismic slip is concentrated at the boundary between these two tectonic blocks.
-Traditionally, the velocity gradient we attribute to a deep detachment has been attributed to broad viscous strain across the eastern Tibetan Plateau (CITE).
-
-Models without surface topography do not have a local minimum northwest of the Beichuan, indicating the relevance of topography in crustal deformation.
-This effect may be more pronounced with a shallower fault dip like in many subduction zones.
-
+The interseismic deformation is much more broadly distributed across over 300km.
+According to this model, interseismically, the eastern Tibetan Plateau moves cohesively with the Sichuan Basin, whereas coseismic slip is concentrated at the boundary between these two blocks.
 Our model of far-field deformation is consistent with a similar detachment model of leveling and GPS-based uplift rates \citep{Hao2014}.
-The vertical velocities increase from the Longmen Shan range front until peaking 50km southeast of the Longriba fault zone at approximately 3.5 mm/yr of uplift.
-The Longriba fault zone has the wrong slip-sense (up to the south-east, \citep{Ren2013}) to accomodate such vertical deformation.
+The vertical velocities they measure increase from the Longmen Shan range front until peaking 50km southeast of the Longriba fault zone at approximately 3.5 mm/yr of uplift.
+The Longriba fault zone has the wrong slip-sense (thrusting up to the south-east, \citep{Ren2013}) to accomodate such vertical deformation.
 No other identified structures are nearby.
 
-\citet{kirby07} study an eastward decrease in interseismic velocities across the Kunlun fault.
+The velocity gradient we attribute to a deep detachment has been attributed to broad viscous strain across the eastern Tibetan Plateau \citep{Royden2008}.
+Relatedly, \citet{kirby07} study an eastward decrease in interseismic velocities across the Kunlun fault.
 They conclude that the decline in slip-rate must be attributed to internal deformation.
-However, the same feature could be explained with slip being accomodated by a wide eastern Tibetan fold-and-thrust system, a suggestion supported by the continuation of shortening north from the Longmen Shan onto the Huya \citep{kirby00} and Min Jiang faults \citep{Chen1994}.
+Further, fast erosion and inferred uplift rates have been used to support lower crustal inflation models in the absence of shortening \citep{Kirby2003}.
+All of these features can be explained with slip accomodated by a wide eastern Tibetan fold-and-thrust system, a suggestion supported by the continuation of shortening north from the Longmen Shan onto the Huya \citep{kirby00} and Min Jiang faults \citep{Chen1994}.
+In Figure \ref{fig:big_stack}c, broad strain is explained with long-term slip on a localized structure.
+Eastward declines in strike-slip motion on the major block-bounding faults can be explained by uplift and shortening accomodated across the eastern Tibetan Plateau.
+With growing evidence for long-term shortening, significant inferred erosion rates are no longer anomalous, being expected in a convergent setting.
+It is unknown how the spatial decoupling of long-term fault motion from short-term interseismic motion influences erosion.
+By the correspondence principle (cite), deep elastic fault motions and viscoelastic relaxation are difficult to distinguish without time-dependent data.
+The Wenchuan earthquake and significant observed afterslip on a deep detachment \citep{Qi2011, Huang2014} in the Longmen Shan support the detachment-based interpretation of eastern Tibetan Plateau tectonics.
+Similar to the Longmen Shan, low shortening rates have been observed on the Huya and Min Jiang fault \citep{Kirby2000}, in contrast to $M_{\textrm{w}}$ 7.5 and 7.2 earthquakes on in that region over the past century \citep{Chen1994}.
+It would be worthwhile to re-examine these shortening rates in the context of a detachment-based interpretation of the eastern Tibetan Plateau.
 
-We have not addressed when the accumulated slip-deficit on the deep detachment is released.
+We have not addressed when accumulated slip-deficit on the deep detachment is released.
 Postseismic slip incorporated in the model of \cite{Qi2011} shows up to 5 meters of slip on a deep detachment.
-Our data analysis depends on the detachment being locked during the period of data collection (WHEN?).
+Our data analysis depends on the detachment being locked during the period of data collection (when?).
 Aseismic slip may nonetheless be the predominant mode of strain release given the short observed portion of the earthquake cycle.
 Seismicity is approximately limited to upper 20km of crust in the range-front \citep{Li2010}.
+Using steady-state critical taper wedge theory, we estimate a basal coefficient of friction between 0.02 and 0.30 for the detachment.
+The mean parameter estimate of 0.05 is similar to estimates of friction on the low-angle detachment system beneath the Longmen Shan foreland \citep{Hubbard2010}, indicating a very weak fault over long time-periods.
 On the other hand, the surface area and shortening rate of the detachment could sustain very large earthquakes if slip is released coseismically.
 
-Using steady-state critical taper wedge theory, we estimate a basal coefficient of friction between 0.02 and 0.30 for the detachment.
-The mean parameter estimate of 0.05 is similar to estimates of friction on the low-angle detachment system beneath the Longmen Shan foreland basin \citep{Hubbard2010}.
+As one of the first boundary element models to explicitly treat topography in crustal deformation analyses, it is interesting to ask which features of the predicted horizontal velocities are a result of topography.
+By comparing to a model without surface topography, the local minimum in velocity (Figure \ref{fig:big_stack}a) northwest of the Beichuan is identified as the main topographically-influenced feature.
+The effects may be more pronounced with a shallower fault dip like in a subduction zones.
 
@@
 This contrasts with the average recurrence of ~2000 years determined by \citep{Ran2010} for the Beichuan fault.
-Our estimate would predict $M_w = $9.0 ruptures for that recurrence, seemingly unlikely given the total slip and fault length required.
+Our estimate would predict $M_{\textrm{w}} = $9.0 ruptures for that recurrence, seemingly unlikely given the total slip and fault length required.
 However, our estimate is an system-wide recurrence while paleoseismic estimates have been limited by only studying one of multiple imbricated thrusts.
@@
 This is because the assumption is equivalent to assuming the fault geometry extends infinitely far along strike, resulting in a fault surface much larger than in reality.
+Approximately 150 kilometers west of the Beichuan fault, there are 3 GPS stations with similar profile location.
+But, the velocities are significantly different.
+Hence, some lateral heterogeneity may not be represented by our two-dimensional modelling.
+This can also be seen in the shift of high topography from being SW-NE striking in the Longmen Shan to S-N striking in the adjacent Min Shan to north.
 Third, the modeled thrust is the steeply dipping Beichuan fault.
@@
 It is also important that the recurrence interval calculation assumed that the entire range-front was ruptured by the Wenchuan earthquake whereas the southern portion of the Longmen Shan did not rupture during the earthquake and is thought to similarly active to the portions that did rupture \citep{Li2010}.
-
-Approximately 150 kilometers west of the Beichuan fault, there are 3 GPS stations with similar profile location.
-But, the velocities are significantly different.
-This indicates some lateral heterogeneity not represented by our two-dimensional modelling.
 
@@
 
+\section{Conclusion}
 Previous near-zero shortening estimates in the Longmen Shan are at odds with the large-moment of the Wenchuan earthquake, very steep topography, structural interpretations indicate fold-and-thrust geometry, fast erosion rates.
 We clarify this conflict by including earthquake cycle effects and a 20km deep detachment.
-Our revised interseismic shortening rate is 5.7 $\pm$ 1.5 mm/yr.
+Our revised interseismic shortening rate is 5.7 $\pm$ 1.5 mm/yr with fault slip-deficit rates up to 9.5 mm/yr near the surface, where the Beichuan fault is steepest.
 The model shows that interseismic strain accumulation may be spatially disjoint with coseismic strain release, emphasizing that accurate fault geometries are critical in the analysis of geodetic data.

### 2014-09-17 paper.tex
Ready for Brendan comments.

@@
 \section{Introduction}
-The Longmen Shan range is located on the eastern margin of the Tibetan and rises 4000m over 30km.
+The Longmen Shan range is located on the eastern margin of the Tibetan and rises 5000m over 30km.
 The 2008 $M_{\textrm{w}}$ 7.9 Wenchuan earthquake ruptured 250km along strike on the Beichuan and Pengguan faults, resulting in approximately 70000 fatalities.
@@
 \citet{Fielding2012a} interpret satellite gravity-field data to infer that mass of the Longmen Shan is supported by a flexed Sichuan Basin.
-Seismic tomography, seismic reflection and magnetotelluric data have been used to suggest that this fold-and-thrust structure extends deep into the plateau as a deep detachment \citep{Zhang2009, Zhao2012, Guo2013}.
+Seismic tomography, seismic reflection and magnetotelluric data have been used to suggest that this fold-and-thrust structure extends far into the plateau as a deep detachment \citep{Zhang2009, Zhao2012, Guo2013}.
 The Longriba fault zone, 200km northwest of the Longmen Shan, is suggested to also sole into this detachment \citep{Xu2008, Ren2013}.
@@
 Prior estimates can be categorized into those that neglect earthquake cycle effects, analyzing the fault slip-deficit without no explicit model of the geometry or tectonics, \citep{chen00, shen05} and those that use block models, which normally assume a simplified fault geometry \citep{Meade07c, Loveless2011, Burchfiel2008a}.
-\citet{Loveless2011} report the highest shortening rate of 3.2 mm/yr.
-Here, we present a model that includes earthquake cycle effects to explain the regional geodetic velocities as the result of a complex interseismically locked fault geometry, including a 20km deep detachment and the range-front thrust Beichuan fault.
+\citet{Loveless2011} report the highest shortening rate of 3.2 mm/yr, as well as significant strain interior to the Longmen Shan block.
+Here, we present a model that includes earthquake cycle effects to explain the regional geodetic velocities as the result of a complex interseismically locked fault geometry, including a 20km deep detachment and the range-front Beichuan fault.
 This geometry is consistent with the inference of 2-6m of slip of a 20km deep detachment that extends 90km northwest using InSAR measurements and post-Wenchuan GPS observations \citep{Qi2011}.
 \citet{Fielding2013b} also include a detachment fault plane in their joint geodetic-teleseismic coseismic slip inversion.
-Further, this geometry is consistent with structural interpretations \citep{Hubbard2010, Li2010}.
+Further, this geometry is suggested by structural interpretations \citep{Hubbard2010, Li2010}.
 Using this range-front thrust plus detachment geometry, we infer an average horizontal shortening between the TIbetan Plateau using interseismic GPS data collected prior to the Wenchuan earthquake.
@@
 Importantly, this model provides increased estimates of seismic hazard.
-This model proposes that strain accumulates up to 300km from the margin of the Tibetan plateau while being released at the Longmen Shan range front, thus providing a unified theory of the Longmen Shan earthquake cycle.
+This model proposes that strain accumulates up to 300km from the margin of the Tibetan plateau while being released at the Longmen Shan range front, providing a unified theory of the Longmen Shan earthquake cycle.
 
@@
 
-The fault geometry we consider is derived from a Community Fault Model for the Longmen Shan and Sichuan Basin region being actively developed (Plesch and Shaw, personal communication).
+The fault geometry we consider is derived from a community fault model for the Longmen Shan and Sichuan Basin region being actively developed (Plesch and Shaw, personal communication).
 This geometry is derived from structural interpretation of both seismic reflection surverys and surface geology \citep{Hubbard2010}.
@@
 The fault geometry consists of two main segments.
-The range-front Beichuan fault dips steeply at ~60$^{\circ}$ at its surface trace shallowing to almost horizontal where it joins the detachment surface at depth.
+The range-front Beichuan fault dips steeply at ~60$^{\circ}$ at its surface trace shallowing to sub-horizontal near 20km depth where it joins the detachment surface at depth.
 The detachment dips at ~1.5$^{\circ}$ from an initial depth of 20km for 110km to the northwest at its deepest point with a depth of 23km.
 
-Because this model includes a deep detachment, we must include interseismic GPS data deep into the interior of the Tibetan Plateau.
+Because this model includes a deep detachment, we must include interseismic GPS data far into the interior of the Tibetan Plateau.
 We study GPS observations collected prior to the Wenchuan earthquake \citep{apel06,Banerjee2008,calais06, gan07, vigny03}, assembled into a single reference frame by \citet{Loveless2011}.
@@
 We include in the estimation velocities that are within the potential far-field influence of a deep detachment under the easternmost Tibetan Plateau.
-We exclude velocities that are near major faults like the Xianshuihe fault or the Kunlun fault.
+We exclude velocities that are near other major faults like the Xianshuihe fault or the Kunlun fault.
 Later, we demonstrate that our results are reasonably robust to the specific data set that is studied.
@@
 To maintain interseismic kinematic consistency between the Sichuan Basin and Tibetan Plateau, we use a 2D block formulation.
-The average interseismic slip-deficit on the fault geometry is subtracted from block motion to calculate the interseismic surface velocity profile.
+The average interseismic slip-deficit on the fault geometry is subtracted from block motion to calculate the interseismic velocity profile (CITE?).
 
@@
 A locked detachment, however, drives the predicted location of steep velocity gradient far to the northwest as shown by Figure \ref{fig:big_stack}a.
-The best-fit shortening rate for the Longmen Shan region is 5.7 $\pm$ 1.5 mm/yr, a substantial increase over previous estimates.
+When including a locked detachment, the least squares best-fit shortening rate for the Longmen Shan region is 5.7 $\pm$ 1.5 mm/yr, a substantial increase over previous estimates.
 For discussion of seismic hazard, shortening must be translated to fault-parallel slip-deficit.
@@
 First, the uncertainty in true interseismic velocities is propagated through to the estimated shortening, resulting in the gray region of uncertainty in Figure \ref{fig:big_stack}a.
-Further, the sparsity of observations suggests that certain observations may undue influence on the shortening estimate.
+Further, the sparsity of observations suggests that certain observations may have undue influence on the shortening estimate.
 Excluding the northwest-most observation decreases the best-fit shortening to 3.9 $\pm$ 1.9 mm/yr.
@@
 Subsequently, we sample jointly from this distribution and the distribution of surface-area weighted fault depths.
-This analysis demonstrates that the majority of the slip-deficit-depth probability landscape lies at slip-deficit rate between 4 and 6 mm/yr while certain parts of the geometry (the range-front) have much faster slip-deficit rates.
+This analysis demonstrates that the majority of the slip-deficit-depth probability landscape lies at slip-deficit rates between 4 and 6 mm/yr while certain parts of the geometry (the range-front) have much faster slip-deficit rates.
 In Figure \ref{fig:distribution}, there is a non-negligible probability assigned to slip-rates above 10 mm/yr in the steepest, near-surface portion of the fault system.
-This is consistent with large near-surface coseismic Wenchuan slip \citep{Shen2009}.
+This is consistent with large near-surface coseismic slip during the Wenchuan earthquake \citep{Shen2009}.
 
@@
 Locking on the deep detachment obscures the signal of deformation on the range-front faults.
-The interseismic deformation is much more broadly distributed across over 300km.
+The interseismic deformation is much more broadly distributed across more than 300 kilometers.
 According to this model, interseismically, the eastern Tibetan Plateau moves cohesively with the Sichuan Basin, whereas coseismic slip is concentrated at the boundary between these two blocks.
-Our model of far-field deformation is consistent with a similar detachment model of leveling and GPS-based uplift rates \citep{Hao2014}.
+
+\citet{Loveless2011} is the only prior geodetic model to include both a dipping Longmen Shan thrust and earthquake cycle effects.
+Compared to their estimated shortening rate of 3.2 mm/yr, our estimate is almost twice as large.
+Their calculations show a comparatively high likelihood of internal block strain in the Longmen Shan region \citep{Loveless2011}.
+Attributing that block-internal strain to a deep detachment instead can explain the discrepancy between the two models.
+Our model of far-field deformation is also consistent with a similar detachment model of leveling and GPS-based uplift rates \citep{Hao2014}.
 The vertical velocities they measure increase from the Longmen Shan range front until peaking 50km southeast of the Longriba fault zone at approximately 3.5 mm/yr of uplift.
@@
 With growing evidence for long-term shortening, significant inferred erosion rates are no longer anomalous, being expected in a convergent setting.
-It is unknown how the spatial decoupling of long-term fault motion from short-term interseismic motion influences erosion.
+It is unknown how the demonstrated spatial decoupling of long-term fault slip from short-term interseismic uplift influences erosion.
+
 By the correspondence principle (cite), deep elastic fault motions and viscoelastic relaxation are difficult to distinguish without time-dependent data.
 The Wenchuan earthquake and significant observed afterslip on a deep detachment \citep{Qi2011, Huang2014} in the Longmen Shan support the detachment-based interpretation of eastern Tibetan Plateau tectonics.
-Similar to the Longmen Shan, low shortening rates have been observed on the Huya and Min Jiang fault \citep{Kirby2000}, in contrast to $M_{\textrm{w}}$ 7.5 and 7.2 earthquakes on in that region over the past century \citep{Chen1994}.
+Similar to the Longmen Shan, low shortening rates have been observed on the Huya and Min Jiang fault \citep{Kirby2000}, in contrast to the $M_{\textrm{w}}$ 7.5 and 7.2 earthquakes in that region over the past century \citep{Chen1994}.
 It would be worthwhile to re-examine these shortening rates in the context of a detachment-based interpretation of the eastern Tibetan Plateau.
@@
 Aseismic slip may nonetheless be the predominant mode of strain release given the short observed portion of the earthquake cycle.
-Seismicity is approximately limited to upper 20km of crust in the range-front \citep{Li2010}.
+Seismicity is approximately limited to the upper 20km of crust in the range-front \citep{Li2010}.
 Using steady-state critical taper wedge theory, we estimate a basal coefficient of friction between 0.02 and 0.30 for the detachment.
-The mean parameter estimate of 0.05 is similar to estimates of friction on the low-angle detachment system beneath the Longmen Shan foreland \citep{Hubbard2010}, indicating a very weak fault over long time-periods.
+The mean parameter estimate of 0.05 is comparable to estimates of friction on the low-angle detachment system beneath the Longmen Shan foreland \citep{Hubbard2010}, indicating a very weak fault over long time-periods.
 On the other hand, the surface area and shortening rate of the detachment could sustain very large earthquakes if slip is released coseismically.
+Specifically, if 100km down-dip and 300km along-strike were to slip 4 meters, similar to average slip during the Wenchuan earthquake, a $M_{\textrm{w}}$ 8.4 rupture would be produced.
 
@@
 The effects may be more pronounced with a shallower fault dip like in a subduction zones.
+Despite the mediocre effect (~15\%), geodetic analyses in steep regions should consider topographic effects to avoid systematic bias.
 
@@
 The approximation will result in under-estimates of the shortening required to produce far-field velocities.
-This is because the assumption is equivalent to assuming the fault geometry extends infinitely far along strike, resulting in a fault surface much larger than in reality.
+This is because the assumption is equivalent to assuming the fault geometry extends infinitely far along strike, a fault surface much larger than reality.
 Approximately 150 kilometers west of the Beichuan fault, there are 3 GPS stations with similar profile location.
@@
 Finally, the shortening rate estimate may be low due to being estimated late in the earthquake cycle \citep{savage00}.
-It is also important that the recurrence interval calculation assumed that the entire range-front was ruptured by the Wenchuan earthquake whereas the southern portion of the Longmen Shan did not rupture during the earthquake and is thought to similarly active to the portions that did rupture \citep{Li2010}.
+It is also important that our recurrence interval calculation assumed the entire range-front was ruptured by the Wenchuan earthquake whereas the southern portion of the Longmen Shan did not rupture during the earthquake and is thought to similarly active to the portions that did rupture \citep{Li2010}.
 
@@
 The model shows that interseismic strain accumulation may be spatially disjoint with coseismic strain release, emphasizing that accurate fault geometries are critical in the analysis of geodetic data.
-This interpretation unifies geodetic models of coseismic (Wenchuan), postseismic and interseismic crustal deformation in the Longmen Shan.
+This interpretation unifies geodetic models of coseismic, postseismic and interseismic crustal deformation in the Longmen Shan.
 Finally, our model suggest that Wenchuan-like events have much shorter recurrence intervals.

### 2014-10-03 paper.tex
Set of BJM edits on 10/03/14

@@
 
-\title{Rapid shortening at the eastern margin of the Tibetan plateau prior to the 2008 Mw=7.9 Wenchuan earthquake}
+\title{Rapid shortening at the eastern margin of the Tibetan plateau prior to the 2008 M_W=7.9 Wenchuan earthquake}
 \author{T. Ben Thompson, Brendan J. Meade}
@@
 \section{Abstract}
-The Longmen Shan is the steepest topographic front of the India-Asia collision and was the site of the Mw=7.9 Wenchuan earthquake.
+The Longmen Shan is the steepest topographic front of the India-Asia collision and was the site of the $\textrm{M}_{\textrm{w}}$ 7.9  Wenchuan earthquake.
 Shortening estimates across the Longmen Shan provide strain accumulation rates and clarify the eastward extrusion of the Tibetan plateau.
-Here, to explain the interseismic GPS velocities across the greater Longmen Shan region, we develop a boundary element model including earthquake cycle effects, topography, the westward dipping Beichuan fault, and a ~20 km deep, shallowly dipping, detachment.
+Here, to explain the interseismic GPS velocities across the greater Longmen Shan region, we develop a boundary element model including earthquake cycle effects, topography, the westward dipping Beichuan fault, and a $\sim 20$ km deep, shallowly dipping, detachment.
 The detachment is inferred from observations of slip during and after the Wenchuan earthquake and from structural considerations.
 Previous analyses which neglected the detachment and earthquake cycle effects have found shortening rates near zero.
-In contrast, we find that interseismic GPS data are consistent with a shortening rate of 5.7$\pm$1.5 mm/yr.
-These results suggest that the Longmen Shan is an active fold-and-thrust belt with Wenchuan style earthquake recurrence intervals of $<$600 years.
+In contrast, we find that interseismic GPS data are consistent with a shortening rate of 5.7$\pm$1.5 mm/yr and maximum surface slip deficit rate of $NNN\pm NNN$ mm/yr.
+This model provides a unified interpretation of geodetic deformation throughout the earthquake cycle and is consistent and suggests that the Longmen Shan is an active fold-and-thrust belt with Wenchuan style earthquake recurrence intervals of $<$600 years.
 
 \section{Introduction}
-The Longmen Shan range is located on the eastern margin of the Tibetan and rises 5000m over 30km.
-The 2008 $M_{\textrm{w}}$ 7.9 Wenchuan earthquake ruptured 250km along strike on the Beichuan and Pengguan faults, resulting in approximately 70000 fatalities.
-Prior geodetic estimates suggest almost zero horizontal shortening \citep{king97, chen00, shen05, Meade07c, Loveless2011}, in contrast with the dominant thrust slip-sense during the Wenchuan rupture.
-Other authors reconcile the lack of shortening with the steep topography by appealing a lower crustal inflation tectonic model for the eastern Tibetan Plateau \citep{royden97, bird91, Burchfiel2008a}.
-However, the magnitude and location of the Wenchuan earthquake suggest a rapidly shortening fold-and-thrust belt.
-From seismic reflection data, \citet{hubbard09} find a well-developed fold-and-thrust fault system geometry with long-term shortenings up to 100\%.
-\citet{Fielding2012a} interpret satellite gravity-field data to infer that mass of the Longmen Shan is supported by a flexed Sichuan Basin.
-Seismic tomography, seismic reflection and magnetotelluric data have been used to suggest that this fold-and-thrust structure extends far into the plateau as a deep detachment \citep{Zhang2009, Zhao2012, Guo2013}.
-The Longriba fault zone, 200km northwest of the Longmen Shan, is suggested to also sole into this detachment \citep{Xu2008, Ren2013}.
+The Longmen Shan range is the steepest margin of the Tibetan Plateau rising 5 km over a distance of only 30 km at the western edge of the Sichuan Basin.
+In 2008 $\textrm{M}_{\textrm{w}}$ 7.9 Wenchuan earthquake ruptured the Beichan and Penguan rangefront thrust faults (over a distance $\sim 250$ km along strike) leading to approximately 70,000 earthquake related fatalities.
+GPS oberservations of co- and post-seismic deformation indicate slip not only on the steeply dipping Beichaun fault but also along a $\sim 20$ km deep sub-horizontal detachment surface extending an additional $\sim 100$ km west of the surface rupture.
+The inference of strain release associated with slip on an active detachment requires a complementary period of strain accumulation prior to the coseismic rupture.
+Interseismic GPS velocities measured prior to the occurence of the Wenchuan earthquake have been interpreted as indicating neglible \citep{king97, chen00, shen05, Meade07c, Loveless2011}.
+This apparent lack of present-day shortening across the steepest topography of the has been hypothesized to result from lower crustal inflation tectonic model for the eastern Tibetan Plateau (cite Clark and Royden papers)\citep{royden97, bird91, Burchfiel2008a}.
+However, these models do not provide a mechanism for accumulating the elastic strain released co- and post-seismically on the deep detachment to the west of the Beichuan fault \citep{Qi2011, Fielding2013b}.
 
-However, it remains to be explained how slow present-day geodetic shortening estimates can be reconciled with the Wenchuan earthquake and the large structurally-observed shortening.
-Prior estimates can be categorized into those that neglect earthquake cycle effects, analyzing the fault slip-deficit without no explicit model of the geometry or tectonics, \citep{chen00, shen05} and those that use block models, which normally assume a simplified fault geometry \citep{Meade07c, Loveless2011, Burchfiel2008a}.
-\citet{Loveless2011} report the highest shortening rate of 3.2 mm/yr, as well as significant strain interior to the Longmen Shan block.
-Here, we present a model that includes earthquake cycle effects to explain the regional geodetic velocities as the result of a complex interseismically locked fault geometry, including a 20km deep detachment and the range-front Beichuan fault.
-This geometry is consistent with the inference of 2-6m of slip of a 20km deep detachment that extends 90km northwest using InSAR measurements and post-Wenchuan GPS observations \citep{Qi2011}.
-\citet{Fielding2013b} also include a detachment fault plane in their joint geodetic-teleseismic coseismic slip inversion.
-Further, this geometry is suggested by structural interpretations \citep{Hubbard2010, Li2010}.
-Using this range-front thrust plus detachment geometry, we infer an average horizontal shortening between the TIbetan Plateau using interseismic GPS data collected prior to the Wenchuan earthquake.
-Next, we show how this geometry can drive the steep velocity gradients far to the northwest of the range front, thus explaining the absence of a range-front interseismic velocity gradient.
-Importantly, this model provides increased estimates of seismic hazard.
-This model proposes that strain accumulates up to 300km from the margin of the Tibetan plateau while being released at the Longmen Shan range front, providing a unified theory of the Longmen Shan earthquake cycle.
-
-\section{Geodetic analysis of Longmen Shan shortening}
+Previous estimates of geodetically constrained shortening and slip deficit rates can be categorized into those that either: 1) do not explicitly treat earthquake cycle effects, \citep{chen00, shen05} (Thatcher too) or 2) those that use block models, with a first order approximation of earthquake cycle effects but coarse representations of fault system geometry \citep{Meade07c, Loveless2011, Burchfiel2008a}.
+Of these, the highest Longmen Shan rangefront slip deficit rates range up to 3.2 mm/yr \citep{Loveless2011}, as well as significant strain interior to the Longmen Shan block.
+Here, we present results from a two-dimensional boundary element models that includes earthquake cycle and topographic effects to explain the regional geodetic velocities that result from the combined effects of interseismic locking on both the range-front Beichuan fault and a 20 km deep detachment beneath the hinterland.
+This fault system geometry is consistent with the inference of 2-6 m of slip of a 20 km deep detachment that extends 90 km northwest using InSAR measurements and post-Wenchuan GPS observations \citep{Qi2011, Fielding2013b} and interpretation of structural observations \citep{Hubbard2010, Li2010}.
+Integrating rangefront and detachment fault system geometry, we develop a model of interseismic strain accumulation that is consistent with both interseismic GPS velocities measured prior to the Wenchuan earthquake and previous observations of co- and post-seismic strain release at depth, and rapid slip deficit rates in the abscence of localized velocity gradients at the Longmen Shan rangefront.
 
@@
     \includegraphics{figs/lms_map_all.pdf}
-    \caption{A map of the Longmen Shan region with the basemap showing shaded relief.
-The focal mechanism plot is located at the epicenter of the Wenchuan earthquake, while the thicker red line shows the surface trace of rupture.
-The thin red lines show 1km depth contours of our model geometry.
-GPS velocity vectors are colored black if they are not included in our parameter estimation and red if they are included.
+    \caption{Tectonic setting of the $\textrm{M}_{\textrm{w}}$ 7.9 Wenchuan earthquake and Longmen Shan rangefront.
+The focal mechanism is located at the epicenter of the Wenchuan earthquake and the thick red line shows the surface trace of rupture.
+The thin red lines show 1 km depth contours of the fault model.
+GPS velocity vectors (reference) are colored red if they are included in the two-dimensional models considered here.
 The blue line shows the two-dimensional cross-section we study.
-Thin black lines indicate other significant faults in the region.}
+Thin black lines indicate other significant faults in the region \citep{Taylor09}.}
     \label{fig:regional_map}
@@
 
-The fault geometry we consider is derived from a community fault model for the Longmen Shan and Sichuan Basin region being actively developed (Plesch and Shaw, personal communication).
-This geometry is derived from structural interpretation of both seismic reflection surverys and surface geology \citep{Hubbard2010}.
-We make no modifications to this geometry except to take a fault-perpindicular cross-section.
-The depth and dip of the detachment are similar to the best-fitting surface found by \citet{Qi2011}.
+\section{Geodetic analysis of Longmen Shan shortening}
+The geometry of the Beichaun fault and the detachment to the west have been inferred from 1) structural interpretations  \citet{Hubbard2010} (Plesch and Shaw, personal communication), 2) seismic relection observations (Ben, you found a paper for this where I misinterpreted the travel times as distances), and 3) geodetic modeling of co- and post-seismic deformation associated with the Wenchuan earthquake \citet{Qi2011, Fielding2013b}.
+The depth and dip of the detachment are similar to the best-fitting surface found by.
 The cross-section and geometry with depth contours are shown in Figure \ref{fig:regional_map}.
-The fault geometry consists of two main segments.
-The range-front Beichuan fault dips steeply at ~60$^{\circ}$ at its surface trace shallowing to sub-horizontal near 20km depth where it joins the detachment surface at depth.
-The detachment dips at ~1.5$^{\circ}$ from an initial depth of 20km for 110km to the northwest at its deepest point with a depth of 23km.
+The fault system geometry consists of two main segments.
+The range-front Beichuan fault dips steeply at $\sim 60^{\circ}$ at its surface trace shallowing to sub-horizontal near 20 km depth where it joins the detachment surface at depth.
+The detachment dips at $\sim 1.5^{\circ}$ from an initial depth of 20 km for 110 km to the northwest at its deepest point with a depth of 23 km.
+For the interseismic shortenging and slip deficit models considered here we consider an idealized two-dimensional representation of these fault surfaces (\ref{fig:regional_map}).
 
-Because this model includes a deep detachment, we must include interseismic GPS data far into the interior of the Tibetan Plateau.
-We study GPS observations collected prior to the Wenchuan earthquake \citep{apel06,Banerjee2008,calais06, gan07, vigny03}, assembled into a single reference frame by \citet{Loveless2011}.
-These velocities are shown in Figure \ref{fig:regional_map} along with the site of the Wenchuan earthquake and other regional faults \citep{Taylor09}.
-We include in the estimation velocities that are within the potential far-field influence of a deep detachment under the easternmost Tibetan Plateau.
-We exclude velocities that are near other major faults like the Xianshuihe fault or the Kunlun fault.
-Later, we demonstrate that our results are reasonably robust to the specific data set that is studied.
-Addition of more recent observations could bias our results due to possible postseismic transient deformation.
+Because of the inference that the ~20 km deep detachment slipped, releasing elastic strain, either co- or post-seismically \citet{Qi2011} as distance >100 km to the west of the Beichaun fault, we include interseismic GPS velocities at similar distance into the interior of the Tibetan Plateau (Figure ref).
+We analyzed interseismic GPS velocities measured during the decade prior to the Wenchuan earthquake \citep{gan07}, in a nominally Eurasian reference frame \citet{apel06}.
+We exclude velocities that are near other major active structures including the Xianshuihe and Kunlun faults (Figure \ref{fig:regional_map}).
 
-To analyze these data, we use a quasistatic elastic boundary element software under active development.
+To model both the tectonic and earthquake cycle processes contributing to these interseismic GPS data, we use a quasistatic elastic boundary element software under active development.
 An important feature of our tool is the inclusion of accurate surface topography in steep mountainous regions or Earth curvature in large-scale problems.
-In addition, we can accurately reproduce half-space solutions to very high precision.
+Previous boundary element methods in the earth sciences (cite thomas 93, maerten, etc) have been based on analytic integrations of rectangular or triangular slip surfaces (cite okada 92, meade 07).
+Developing these analytic formulae is time consuming.
+Analytic derivations normally assume an infinite half-space geometry and piecewise constant fault slip.
+To avoid these limitations, the basis for our boundary element formulation is numerical quadrature of the integrals in the elastic Somigliana identity (cite cruse 1969).
+The difficulty in this approach is that simple quadrature methods (for example, Gaussian quadrature) assume a smooth integrand.
+For observation points on or very close to a surface, including the free surface, the integrand is singular.
+We resolve this problem by evaluating singular integrals as the limit of a sequence of non-singular integrals.
+This approach is similar to analytical methods (Sutradhar et al.
+2008) and the series-expansion quadrature by expansion (Klockner et al.
+2013).
+Our tests of relative error show that this quadrature-based method can match, to arbitrary precision, analytically derived solutions and is free of the unpleasant singularities present in those solutions.
+This entirely numerical boundary element method allows for complete flexibility in surface geometry, including topography.
+In addition, we can accurately reproduce half-space solutions to very high precision (Quantify).
 The steep topography of the Longmen Shan range front is included via the SRTM30\_PLUS digital elevation dataset \citep{Becker2009}.
-To maintain interseismic kinematic consistency between the Sichuan Basin and Tibetan Plateau, we use a 2D block formulation.
-The average interseismic slip-deficit on the fault geometry is subtracted from block motion to calculate the interseismic velocity profile (CITE?).
+
+   Say what we actually solve for?  A single shortening rate which leads to the distribution of slip deficit rates.
+
+To maintain kinematic consistency between the Sichuan Basin and Tibetan Plateau, we use a two-dimensional block formulation where the slip deficit rate on the fault, $s_\mathrm{sd}$, is related to the local fault dip, $\delta$, and total convergence rate, $v_0$, by $s_\mathrm{sd}=v_0 / cos\delta$ (Mccafrey 2002; Meade and Hager, 2005).
+The average interseismic slip deficit on the fault geometry is subtracted from block motion to calculate the interseismic velocity profile (Savage, 1983; Meade and Hager, 2005).
 
@@
     \includegraphics{figs/stack_figure_all_details.pdf}
-    \caption{Axes are not one to one.
+    \caption{GPS velocities, model geometry, and inferend slip deficit rates.
 a) Observed profile-parallel horizontal observed velocities are shown in black with 68\% confidence errorbars.
@@
 c) Predicted horizontal velocity as a function of depth.
-Note the steep velocity gradient and corresponding strain accumulation above the end of the detachment surface.}
+Note the steep velocity gradient and corresponding strain accumulation above the end of the detachment surface.
+Vertical axes are exagerated in all panels.}
     \label{fig:big_stack}
@@
 
-The debate over shortening in the Longmen Shan has centered around the clear lack of an interseismic velocity gradient at the range front.
-However, this discussion has neglected the interplay between earthquake cycle effects and fault geometry.
-A Beichuan-only fault geometry would predict very little deformation beyond 100km \citep{savage83}, leading to the exclusion of velocities beyond this range.
-Both the observed and predicted profile-parallel velocities shown in Figure \ref{fig:big_stack}a make clear that such an estimate would result in a near-zero shortening estimate.
+A Beichuan-only fault geometry (give approximate dip and locking depth) would predict <0.1 mm/yr of surface deformation at distances >100 km away from the fault trace \citep{savage83}.
+Both the observed and predicted profile-parallel velocities (Figure \ref{fig:big_stack}a) indicate that such an estimate would appear consistent with a near-zero slip deficit rate estimate and also fail to explain the co-/post-seismic slip on the detachment \citet{Qi2011, Fielding2013b}.
 
-A locked detachment, however, drives the predicted location of steep velocity gradient far to the northwest as shown by Figure \ref{fig:big_stack}a.
-When including a locked detachment, the least squares best-fit shortening rate for the Longmen Shan region is 5.7 $\pm$ 1.5 mm/yr, a substantial increase over previous estimates.
-For discussion of seismic hazard, shortening must be translated to fault-parallel slip-deficit.
-Figure \ref{fig:big_stack}b shows the increase in slip-deficit from ~6.0 mm/yr up to 9.5 mm/yr at the surface.
-Figure \ref{fig:big_stack}c shows the horizontal velocities as a function of depth, demonstrating that the velocity gradients derive from the detachment tip.
+The inclusion of a locked detachment, however, shifts the predicted location of the steep interseismic velocity gradient far to the northwest (Figure \ref{fig:big_stack}a).
+When including a locked detachment, the least squares best-fit shortening rate for the Longmen Shan region is 5.7 $\pm$ 1.5 mm/yr.
+The spatial distribution of interseismic deformation is affected by the fault geometry and kinematic consistency constraints and represented in terms of the slip deficit rate which ranges shows the increase in slip-deficit from ~6.0 mm/yr on the deep detachment to 9.5 mm/yr at the surface trace of the Beichuan fault (Figure \ref{fig:big_stack}b).
+Figure \ref{fig:big_stack}c shows the horizontal velocities as a function of depth, demonstrating that the velocity gradients derive a locked to creeping transition at the detachment tip.
 The smoothing behavior of an elastic Earth prevents distinguishing between a sharp drop in slip-deficit or a gradual change over many kilometers.
@@
 Excluding the northwest-most observation decreases the best-fit shortening to 3.9 $\pm$ 1.9 mm/yr.
+(Add the range of slip deficit rates because the near surface Beichaun rate will still probably be close to 6 mm/yr).
 The least squares uncertainty is increased (from 1.5 mm/yr to 1.9 mm/yr) because the northwestern portion of velocities are poorly constrained once the observation is excluded.
@@
 
-Figure \ref{fig:distribution} shows a more complete analysis of the effect of case resampling the observations on the slip-deficit rate.
-Because there only $2^{20} = 1.048 $ million different possible subsets of the observations, we perform a complete case resampling bootstrap.
+To quantify the sensitivity of slip deficit rate estimates to variability in data selection  we perform a complete case resampling bootstrap shortening and slip deficit rates are calculated for all possible subsets of the observed GPS velocities.
+There only $2^{20} = 1,048,576$ different possible subsets of the observations (Figure \ref{fig:distribution}).
 Subsequently, we sample jointly from this distribution and the distribution of surface-area weighted fault depths.
-This analysis demonstrates that the majority of the slip-deficit-depth probability landscape lies at slip-deficit rates between 4 and 6 mm/yr while certain parts of the geometry (the range-front) have much faster slip-deficit rates.
-In Figure \ref{fig:distribution}, there is a non-negligible probability assigned to slip-rates above 10 mm/yr in the steepest, near-surface portion of the fault system.
-This is consistent with large near-surface coseismic slip during the Wenchuan earthquake \citep{Shen2009}.
+This analysis demonstrates that the majority of the slip-deficit-depth probability landscape lies at slip-deficit rates between 4 and 6 mm/yr while certain parts of the geometry (the range-front) have much faster slip-deficit rates (give range of near surface rates).
+There is a non-negligible probability (give probability) assigned to slip-rates above 10 mm/yr in the steepest, near-surface portion of the fault system (Figure \ref{fig:distribution}).
+This is consistent with large near-surface coseismic slip (Quantify the max slip in any of the coseismic papers) during the Wenchuan earthquake \citep{Shen2009}.
 
@@
 \section{Implications for the Longmen Shan earthquake cycle}
-The striking feature of Figure \ref{fig:big_stack}a is the lack of a range-front geodetic signature despite the significant shortening rate.
+The fundamental effect of allowing a 20 km detachment to accumulated and release elastic strain through the earthquake cycle is the lack of a range-front geodetic signature despite the significant ($5-10$ mm/yr) slip deficit rate (\ref{fig:big_stack}a).
 Locking on the deep detachment obscures the signal of deformation on the range-front faults.
@@
 
-\citet{Loveless2011} is the only prior geodetic model to include both a dipping Longmen Shan thrust and earthquake cycle effects.
-Compared to their estimated shortening rate of 3.2 mm/yr, our estimate is almost twice as large.
-Their calculations show a comparatively high likelihood of internal block strain in the Longmen Shan region \citep{Loveless2011}.
+Prior analyses of interseismic deformation at the Longmen Shan including the effects of dipping fault system geometry and earthquake cycle effects have been limited (\citep{Loveless2011, Qi2011, Hao2014}).
+Of these only the three-dimensional block model of \citet{Loveless2011} is kinematically consistent (Jordan and Minster, McCaffrey, 2002; Meade and Hager, 2005).
+Compared to their estimated slip rate of 3.2 mm/yr, our estimate $2.5 \times$ as large.
+Their calculations indicate a maximum 37\% likelihood of internal block strain west of the Longmen Shan region \citep{Loveless2011}.
 Attributing that block-internal strain to a deep detachment instead can explain the discrepancy between the two models.
 Our model of far-field deformation is also consistent with a similar detachment model of leveling and GPS-based uplift rates \citep{Hao2014}.
-The vertical velocities they measure increase from the Longmen Shan range front until peaking 50km southeast of the Longriba fault zone at approximately 3.5 mm/yr of uplift.
+The vertical velocities they measure increase from the Longmen Shan range front until peaking 50 km southeast of the Longriba fault zone at approximately 3.5 mm/yr of uplift.
 The Longriba fault zone has the wrong slip-sense (thrusting up to the south-east, \citep{Ren2013}) to accomodate such vertical deformation.
@@
 
-The velocity gradient we attribute to a deep detachment has been attributed to broad viscous strain across the eastern Tibetan Plateau \citep{Royden2008}.
+The velocity gradient which we link to the physics of the earthquake cycle and mapped fault system geometry has been alternatively asserted to represent distributed strain across the eastern Tibetan Plateau \citep{Royden2008}.
 Relatedly, \citet{kirby07} study an eastward decrease in interseismic velocities across the Kunlun fault.
@@
 All of these features can be explained with slip accomodated by a wide eastern Tibetan fold-and-thrust system, a suggestion supported by the continuation of shortening north from the Longmen Shan onto the Huya \citep{kirby00} and Min Jiang faults \citep{Chen1994}.
-In Figure \ref{fig:big_stack}c, broad strain is explained with long-term slip on a localized structure.
-Eastward declines in strike-slip motion on the major block-bounding faults can be explained by uplift and shortening accomodated across the eastern Tibetan Plateau.
-With growing evidence for long-term shortening, significant inferred erosion rates are no longer anomalous, being expected in a convergent setting.
-It is unknown how the demonstrated spatial decoupling of long-term fault slip from short-term interseismic uplift influences erosion.
+In Figure \ref{fig:big_stack}c.
 
-By the correspondence principle (cite), deep elastic fault motions and viscoelastic relaxation are difficult to distinguish without time-dependent data.
-The Wenchuan earthquake and significant observed afterslip on a deep detachment \citep{Qi2011, Huang2014} in the Longmen Shan support the detachment-based interpretation of eastern Tibetan Plateau tectonics.
-Similar to the Longmen Shan, low shortening rates have been observed on the Huya and Min Jiang fault \citep{Kirby2000}, in contrast to the $M_{\textrm{w}}$ 7.5 and 7.2 earthquakes in that region over the past century \citep{Chen1994}.
-It would be worthwhile to re-examine these shortening rates in the context of a detachment-based interpretation of the eastern Tibetan Plateau.
-
-We have not addressed when accumulated slip-deficit on the deep detachment is released.
-Postseismic slip incorporated in the model of \cite{Qi2011} shows up to 5 meters of slip on a deep detachment.
+While we have explained the observed the broad GPS interseismic velocity gradient as a result of a locked detachment and the Beichaun fault we have not addressed when accumulated slip-deficit on the deep detachment is released.
+Geodetically constrained estimates of co- and post-seismic slip shows up to 5 m of slip on a deep detachment \citep{Qi2011}.
 Our data analysis depends on the detachment being locked during the period of data collection (when?).
-Aseismic slip may nonetheless be the predominant mode of strain release given the short observed portion of the earthquake cycle.
+It is possible that in addition to coseismic and short-term postseismic activity on the detachment intermittent aseismic throughout the earthquake cycle could also be mode of strain release given that intersimsic GPS observations cover at best 6\% of a complete Longmen Shan earthquake cycle.
 Seismicity is approximately limited to the upper 20km of crust in the range-front \citep{Li2010}.
-Using steady-state critical taper wedge theory, we estimate a basal coefficient of friction between 0.02 and 0.30 for the detachment.
-The mean parameter estimate of 0.05 is comparable to estimates of friction on the low-angle detachment system beneath the Longmen Shan foreland \citep{Hubbard2010}, indicating a very weak fault over long time-periods.
+Using steady-state critical taper wedge theory, we estimate a basal coefficient of friction between 0.02 and 0.30 (why does this vary so much?) for the detachment(cite Dahlen and give assumed pore fluid pressure and whether or not cohesion is assumed).
+The mean value of the coefficient of friction 0.05 is comparable to estimates of friction on the low-angle detachment system beneath the Longmen Shan foreland \citep{Hubbard2010}, indicating a very weak fault over long time-periods.
 On the other hand, the surface area and shortening rate of the detachment could sustain very large earthquakes if slip is released coseismically.
@@
 
-As one of the first boundary element models to explicitly treat topography in crustal deformation analyses, it is interesting to ask which features of the predicted horizontal velocities are a result of topography.
-By comparing to a model without surface topography, the local minimum in velocity (Figure \ref{fig:big_stack}a) northwest of the Beichuan is identified as the main topographically-influenced feature.
+We also quantify the effect of explicitly treating surface topography by comparing models with and without surface topography, the local minimum in velocity (Figure \ref{fig:big_stack}a) northwest of the Beichuan is identified as the main topographically-influenced feature.
 The effects may be more pronounced with a shallower fault dip like in a subduction zones.
-Despite the mediocre effect (~15\%), geodetic analyses in steep regions should consider topographic effects to avoid systematic bias.
+Despite the mediocre effect (~15\%), geodetic analyses in steep regions should consider topographic effects to avoid systematic bias and these effects may be even more signficant for coseismic faulting cases where deformation is localized in regions of steep topography.
 
-Figure \ref{fig:hazard} shows a summary of recurrence intervals as a function of total event slip.
-Our shortening estimate suggests a recurrence interval for Wenchuan-like ruptures of approximately 600 years.
-Recurrence for smaller events is much more frequent.
-This contrasts with the average recurrence of ~2000 years determined by \citep{Ran2010} for the Beichuan fault.
-Our estimate would predict $M_{\textrm{w}} = $9.0 ruptures for that recurrence, seemingly unlikely given the total slip and fault length required.
-However, our estimate is an system-wide recurrence while paleoseismic estimates have been limited by only studying one of multiple imbricated thrusts.
+The 6-9 mm/yr slip deficit rates estimated here can be used to provide new constraints on earthquake recurrence intervals at the Longmen Shan (Figure \ref{fig:hazard}).
+Our shortening estimate suggests a recurrence interval for Wenchuan-like ruptures of approximately 600 years while recurrence intervals for smaller events would be more frequent.
+This contrasts with the average recurrence of ~2000 years inferred from paleoseismic field investigations for the Beichuan fault \citep{Ran2010}.
+The 6-9 mm/ry slip deficit rate estimated here would imply $M_{\textrm{w}} = $9.0 ruptures for that recurrence interval, which would seemingly require a along struke rupture equal to or greater than the $\sim 300$ km along strike length of the Longmen Shan, our slip deficit estimate could include the effects of other imbricated structures and thus serving as a system-wide estimate while paleoseismic estimates have been focused on the analysis of individual thrust faults.
 
-There are reasons to believe that the seismic hazard estimate presented here is a low estimate.
-First, coseismic and long term slip estimates include a significant component of strike-slip motion\citep{Shen2009, Qi2011, Densmore2007} .
-But, we only model horizontal shortening.
-Second, our two-dimensional model assumes plane strain conditions.
-The approximation will result in under-estimates of the shortening required to produce far-field velocities.
-This is because the assumption is equivalent to assuming the fault geometry extends infinitely far along strike, a fault surface much larger than reality.
-Approximately 150 kilometers west of the Beichuan fault, there are 3 GPS stations with similar profile location.
-But, the velocities are significantly different.
-Hence, some lateral heterogeneity may not be represented by our two-dimensional modelling.
-This can also be seen in the shift of high topography from being SW-NE striking in the Longmen Shan to S-N striking in the adjacent Min Shan to north.
+The earthquake recurrence intervals calculated above may be considered a lower bound for three reasons.
+First, coseismic, interseismic, and long-term slip estimates include a significant component of strike-slip motion \citep{Shen2009, Qi2011, Densmore2007} (Meade2007, Loveless and Meade, 2011).
+while we consider an idealized case of fault normal motion only.
+Second, our two-dimensional model assumes plane strain conditions which will result in under-estimates of the shortening required to produce far-field velocities.
+This is because the assumption is equivalent to assuming the fault geometry extends infinitely far along strike.
 Third, the modeled thrust is the steeply dipping Beichuan fault.
 The foreland Range Front thrust or the shallow detachment beneath Chengdu dip more shallowly \citep{Hubbard2010}, producing a greater total moment deficit.
-Finally, the shortening rate estimate may be low due to being estimated late in the earthquake cycle \citep{savage00}.
-It is also important that our recurrence interval calculation assumed the entire range-front was ruptured by the Wenchuan earthquake whereas the southern portion of the Longmen Shan did not rupture during the earthquake and is thought to similarly active to the portions that did rupture \citep{Li2010}.
+Finally, the slip deficit rate estimates may be low due to being estimated late in the earthquake cycle (i.e., in the decade prior to the Wenchuan earthquake) when viscoalstic relaxation effects may further contribute to a lowe near fault trace velocity gradient e.g., \citep{savage00}.
 
@@
     \includegraphics{figs/hazard_all_details.pdf}
-    \caption{Recurrence as a function of slip, assuming the best-fit slip rate for the Beichuan fault.
+    \caption{Earthquake recurrence intervals as a function of coseismic slip, assuming the best-fit slip rate for the Beichuan fault.
 The gray region indicates the uncertainty implied by the error in our shortening estimate.
@@
 \section{Conclusion}
-Previous near-zero shortening estimates in the Longmen Shan are at odds with the large-moment of the Wenchuan earthquake, very steep topography, structural interpretations indicate fold-and-thrust geometry, fast erosion rates.
-We clarify this conflict by including earthquake cycle effects and a 20km deep detachment.
-Our revised interseismic shortening rate is 5.7 $\pm$ 1.5 mm/yr with fault slip-deficit rates up to 9.5 mm/yr near the surface, where the Beichuan fault is steepest.
-The model shows that interseismic strain accumulation may be spatially disjoint with coseismic strain release, emphasizing that accurate fault geometries are critical in the analysis of geodetic data.
-This interpretation unifies geodetic models of coseismic, postseismic and interseismic crustal deformation in the Longmen Shan.
-Finally, our model suggest that Wenchuan-like events have much shorter recurrence intervals.
+Previous inferences of near-zero interseismic shortening estimates at the Longmen Shan have been difficult to reconcile the Longmen Shan are at odds with the large-moment of the Wenchuan earthquake, very steep topography, structural interpretations indicate fold-and-thrust geometry, fast erosion rates.
+We offer a resolution to this conflict by including earthquake cycle effects associated with interseismic locking on the Beichuan fault and a 20 km deep detachment to the west.
+Our revised interseismic shortening rate is 5.7 $\pm$ 1.5 mm/yr with on fault slip-deficit rates up to 9.5 mm/yr near the surface, where the Beichuan fault is most steeply dipping.
+The model demonstrates that interseismic strain accumulation may be spatially disjoint from coseismic strain release, emphasizing that accurate fault system geometries are critical in the analysis of geodetic data.
+This boundary element model presented here unifies geodetic models of coseismic, postseismic and interseismic crustal deformation across the greater Longmen Shan region and suggests that, with current loading rates, Wenchuan-like earthquakes may recurrence as frequently as every 600 years.
 

### 2014-10-12 paper.tex
Ben edits. Cleaned up some of Brendan's edits and added a couple thoughts.

@@
 
-\title{Rapid shortening at the eastern margin of the Tibetan plateau prior to the 2008 M_W=7.9 Wenchuan earthquake}
+\title{Rapid shortening at the eastern margin of the Tibetan plateau prior to the 2008 $M_\textrm{w}$ 7.9 Wenchuan earthquake}
 \author{T. Ben Thompson, Brendan J. Meade}
@@
 Shortening estimates across the Longmen Shan provide strain accumulation rates and clarify the eastward extrusion of the Tibetan plateau.
-Here, to explain the interseismic GPS velocities across the greater Longmen Shan region, we develop a boundary element model including earthquake cycle effects, topography, the westward dipping Beichuan fault, and a $\sim 20$ km deep, shallowly dipping, detachment.
+Here, to explain the interseismic GPS velocities across the greater Longmen Shan region, we develop a boundary element model including earthquake cycle effects, topography, the westward dipping Beichuan fault, and a ${\sim}20$ km deep, shallowly dipping, detachment.
 The detachment is inferred from observations of slip during and after the Wenchuan earthquake and from structural considerations.
 Previous analyses which neglected the detachment and earthquake cycle effects have found shortening rates near zero.
-In contrast, we find that interseismic GPS data are consistent with a shortening rate of 5.7$\pm$1.5 mm/yr and maximum surface slip deficit rate of $NNN\pm NNN$ mm/yr.
-This model provides a unified interpretation of geodetic deformation throughout the earthquake cycle and is consistent and suggests that the Longmen Shan is an active fold-and-thrust belt with Wenchuan style earthquake recurrence intervals of $<$600 years.
+In contrast, we find that interseismic GPS data are consistent with a shortening rate of 5.7$\pm$1.5 mm/yr and maximum surface slip-deficit rate of 9.5 $\pm$ 2.5 mm/yr.
+This model provides a unified interpretation of geodetic deformation throughout the earthquake cycle and suggests that the Longmen Shan is an active fold-and-thrust belt with Wenchuan style earthquake recurrence intervals of $<$600 years.
 
@@
 The Longmen Shan range is the steepest margin of the Tibetan Plateau rising 5 km over a distance of only 30 km at the western edge of the Sichuan Basin.
-In 2008 $\textrm{M}_{\textrm{w}}$ 7.9 Wenchuan earthquake ruptured the Beichan and Penguan rangefront thrust faults (over a distance $\sim 250$ km along strike) leading to approximately 70,000 earthquake related fatalities.
-GPS oberservations of co- and post-seismic deformation indicate slip not only on the steeply dipping Beichaun fault but also along a $\sim 20$ km deep sub-horizontal detachment surface extending an additional $\sim 100$ km west of the surface rupture.
-The inference of strain release associated with slip on an active detachment requires a complementary period of strain accumulation prior to the coseismic rupture.
-Interseismic GPS velocities measured prior to the occurence of the Wenchuan earthquake have been interpreted as indicating neglible \citep{king97, chen00, shen05, Meade07c, Loveless2011}.
-This apparent lack of present-day shortening across the steepest topography of the has been hypothesized to result from lower crustal inflation tectonic model for the eastern Tibetan Plateau (cite Clark and Royden papers)\citep{royden97, bird91, Burchfiel2008a}.
-However, these models do not provide a mechanism for accumulating the elastic strain released co- and post-seismically on the deep detachment to the west of the Beichuan fault \citep{Qi2011, Fielding2013b}.
+In 2008, the $\textrm{M}_{\textrm{w}}$ 7.9 Wenchuan earthquake ruptured the Beichan and Penguan rangefront thrusts ${\sim}250$ km along strike, leading to approximately 70,000 fatalities.
+The energy release associated with the Wenchuan earthquake requires a complimentary period of strain accumulation.
+However, interseismic GPS velocities measured prior to the occurence of the Wenchuan earthquake have been interpreted as indicating neglible shortening \citep{king97, chen00, shen05, Meade07c, Loveless2011}.
+This apparent lack of present-day shortening across the steepest topography of the India-Asia collision has been hypothesized to result from lower crustal inflation tectonic model for the eastern Tibetan Plateau \citep{royden97, bird91, Burchfiel2008, Clark2000}.
+But, these models do not provide a mechanism for accumulating the elastic strain released coseismically during the Wenchuan earthquake.
+Recent observations of co- and post-seismic slip on a deep detachment to the west of the Beichuan fault\citep{Qi2011, Fielding2013b} motivate a new model of Longmen Shan interseismic shortening that includes a locked detachment.
 
-Previous estimates of geodetically constrained shortening and slip deficit rates can be categorized into those that either: 1) do not explicitly treat earthquake cycle effects, \citep{chen00, shen05} (Thatcher too) or 2) those that use block models, with a first order approximation of earthquake cycle effects but coarse representations of fault system geometry \citep{Meade07c, Loveless2011, Burchfiel2008a}.
-Of these, the highest Longmen Shan rangefront slip deficit rates range up to 3.2 mm/yr \citep{Loveless2011}, as well as significant strain interior to the Longmen Shan block.
-Here, we present results from a two-dimensional boundary element models that includes earthquake cycle and topographic effects to explain the regional geodetic velocities that result from the combined effects of interseismic locking on both the range-front Beichuan fault and a 20 km deep detachment beneath the hinterland.
-This fault system geometry is consistent with the inference of 2-6 m of slip of a 20 km deep detachment that extends 90 km northwest using InSAR measurements and post-Wenchuan GPS observations \citep{Qi2011, Fielding2013b} and interpretation of structural observations \citep{Hubbard2010, Li2010}.
-Integrating rangefront and detachment fault system geometry, we develop a model of interseismic strain accumulation that is consistent with both interseismic GPS velocities measured prior to the Wenchuan earthquake and previous observations of co- and post-seismic strain release at depth, and rapid slip deficit rates in the abscence of localized velocity gradients at the Longmen Shan rangefront.
+Previous estimates of geodetically constrained shortening and slip-deficit rates in the Longmen Shan region can be categorized into those that either: 1) do not explicitly treat earthquake cycle effects, \citep{chen00, shen05, Thatcher2007} or 2) those that use block models, with a first order approximation of earthquake cycle effects but coarse representations of fault system geometry \citep{Meade07c, Loveless2011, Burchfiel2008}.
+The inferred slip-deficit rates range from 0.0 mm/yr \citep{Thatcher2007} to 3.2 mm/yr \citep{Loveless2011}, with all analyses reporting significant strain 100-200 km northwest of the Longmen Shan in a region termed the Songpan-Xihe deformation zone \citep{shen05}.
+Here, we present results from a two-dimensional boundary element model that includes earthquake cycle and topographic effects to explain the regional geodetic velocities as the result of interseismic locking on both the range-front Beichuan fault and a 20 km deep detachment beneath the hinterland.
+This fault system geometry is consistent with the inference of 2-6 m of slip of a 20 km deep detachment that extends 90 km northwest using InSAR measurements and post-Wenchuan GPS observations \citep{Qi2011, Fielding2013b} and interpretation of structural observations \citep{Hubbard2010, Li2010a}.
+Integrating rangefront and detachment fault system geometry, we develop a model of interseismic strain accumulation that is consistent with both interseismic GPS velocities measured prior to the Wenchuan earthquake and previous observations of co- and post-seismic strain release at depth.
+Importantly, the model provides a mechanism for large slip-deficit rates in the abscence of a localized velocity gradient at the Longmen Shan rangefront.
 
@@
 The thin red lines show 1 km depth contours of the fault model.
-GPS velocity vectors (reference) are colored red if they are included in the two-dimensional models considered here.
+GPS velocity vectors (reference) are colored red if they are included in the two-dimensional models we consider.
 The blue line shows the two-dimensional cross-section we study.
@@
 \section{Geodetic analysis of Longmen Shan shortening}
-The geometry of the Beichaun fault and the detachment to the west have been inferred from 1) structural interpretations  \citet{Hubbard2010} (Plesch and Shaw, personal communication), 2) seismic relection observations (Ben, you found a paper for this where I misinterpreted the travel times as distances), and 3) geodetic modeling of co- and post-seismic deformation associated with the Wenchuan earthquake \citet{Qi2011, Fielding2013b}.
-The depth and dip of the detachment are similar to the best-fitting surface found by.
-The cross-section and geometry with depth contours are shown in Figure \ref{fig:regional_map}.
-The fault system geometry consists of two main segments.
-The range-front Beichuan fault dips steeply at $\sim 60^{\circ}$ at its surface trace shallowing to sub-horizontal near 20 km depth where it joins the detachment surface at depth.
-The detachment dips at $\sim 1.5^{\circ}$ from an initial depth of 20 km for 110 km to the northwest at its deepest point with a depth of 23 km.
-For the interseismic shortenging and slip deficit models considered here we consider an idealized two-dimensional representation of these fault surfaces (\ref{fig:regional_map}).
+The geometry of the Beichaun fault and the detachment to the west have been inferred from 1) structural interpretations  \citep{Hubbard2010} (Plesch and Shaw, personal communication), 2) seismic relection observations \citep{Zhang2009, Guo2013}, and 3) geodetic modeling of co- and post-seismic deformation associated with the Wenchuan earthquake \citep{Qi2011, Fielding2013b}.
+The depth and dip of the detachment are similar to the best-fitting surface found by \citet{Qi2011}.
+The fault system geometry consists of two main segments (Figure \ref{fig:regional_map}).
+The range-front Beichuan fault dips steeply at ${\sim}60^{\circ}$ at its surface trace shallowing to sub-horizontal near 20 km depth where it joins the detachment surface at depth.
+The detachment dips at ${\sim}1.5^{\circ}$ from an initial depth of 20 km for 110 km to the northwest at its deepest point with a depth of 23 km.
+For our interseismic shortening and slip-deficit models we use an idealized two-dimensional representation of these fault surfaces (Figure \ref{fig:regional_map}).
 
-Because of the inference that the ~20 km deep detachment slipped, releasing elastic strain, either co- or post-seismically \citet{Qi2011} as distance >100 km to the west of the Beichaun fault, we include interseismic GPS velocities at similar distance into the interior of the Tibetan Plateau (Figure ref).
-We analyzed interseismic GPS velocities measured during the decade prior to the Wenchuan earthquake \citep{gan07}, in a nominally Eurasian reference frame \citet{apel06}.
+The inference that a 20km deep detachment slipped, either co- or post-seismically, at distances up to 100km from the range-front \citep{Qi2011}, necessitates including interseismic GPS velocities at a similar distance into the interior of the Tibetan Plateau (Figure \ref{fig:regional_map}).
+We analyze interseismic GPS velocities measured during the decade prior to the Wenchuan earthquake \citep{gan07}, in a nominally Eurasian reference frame \citet{apel06}.
 We exclude velocities that are near other major active structures including the Xianshuihe and Kunlun faults (Figure \ref{fig:regional_map}).
@@
 An important feature of our tool is the inclusion of accurate surface topography in steep mountainous regions or Earth curvature in large-scale problems.
-Previous boundary element methods in the earth sciences (cite thomas 93, maerten, etc) have been based on analytic integrations of rectangular or triangular slip surfaces (cite okada 92, meade 07).
-Developing these analytic formulae is time consuming.
-Analytic derivations normally assume an infinite half-space geometry and piecewise constant fault slip.
-To avoid these limitations, the basis for our boundary element formulation is numerical quadrature of the integrals in the elastic Somigliana identity (cite cruse 1969).
-The difficulty in this approach is that simple quadrature methods (for example, Gaussian quadrature) assume a smooth integrand.
+Previous boundary element methods in the earth sciences \citep{Thomas1993, Maerten2014a} have been based on analytic integrations of rectangular or triangular slip surfaces \citep{Okada1992, Meade2007}.
+These analytic formulae normally assume an infinite half-space geometry and piecewise constant slip.
+Even so, they are time consuming to derive.
+To avoid these limitations, the basis for our boundary element formulation is numerical quadrature of the integrals in the full-space elastic Somigliana identity \citep{Cruse1969}.
+The difficulty in this approach is that straightforward quadrature methods (for example, Gaussian quadrature) assume a smooth integrand.
 For observation points on or very close to a surface, including the free surface, the integrand is singular.
 We resolve this problem by evaluating singular integrals as the limit of a sequence of non-singular integrals.
-This approach is similar to analytical methods (Sutradhar et al.
-2008) and the series-expansion quadrature by expansion (Klockner et al.
-2013).
-Our tests of relative error show that this quadrature-based method can match, to arbitrary precision, analytically derived solutions and is free of the unpleasant singularities present in those solutions.
+This approach is similar to analytical limit-to-the-boundary methods \citep{Sutradhar2008} and series-expansion methods \citep{Klockner2013}.
+Our tests of relative error show that this quadrature-based method can match, to arbitrary precision, analytically derived half-space solutions and is free of the unpleasant singularities present in those solutions.
 This entirely numerical boundary element method allows for complete flexibility in surface geometry, including topography.
-In addition, we can accurately reproduce half-space solutions to very high precision (Quantify).
-The steep topography of the Longmen Shan range front is included via the SRTM30\_PLUS digital elevation dataset \citep{Becker2009}.
+The steep topography of the Longmen Shan rangefront is included via the SRTM30\_PLUS digital elevation dataset \citep{Becker2009}.
 
-   Say what we actually solve for?  A single shortening rate which leads to the distribution of slip deficit rates.
-
-To maintain kinematic consistency between the Sichuan Basin and Tibetan Plateau, we use a two-dimensional block formulation where the slip deficit rate on the fault, $s_\mathrm{sd}$, is related to the local fault dip, $\delta$, and total convergence rate, $v_0$, by $s_\mathrm{sd}=v_0 / cos\delta$ (Mccafrey 2002; Meade and Hager, 2005).
-The average interseismic slip deficit on the fault geometry is subtracted from block motion to calculate the interseismic velocity profile (Savage, 1983; Meade and Hager, 2005).
+Using this boundary element model, we solve for the single best-fitting shortening rate between the Tibetan Plateau and the Sichuan Basin while maintaining kinematic consistency using a two-dimensional block formulation.
+In the block formulation, the slip-deficit rate on the fault, $s_\mathrm{sd}$, is related to the local fault dip, $\delta$, and total convergence rate, $v_0$, by $s_\mathrm{sd}=v_0 / cos\delta$ \citep{mccaffrey02, Meade2005}.
+As a result, slip-deficit rates are strictly larger than shortening rates.
+The average interseismic slip-deficit on the fault geometry is subtracted from block motion to calculate the interseismic velocity profile \citep{savage83, Meade2005}.
 
@@
     \includegraphics{figs/stack_figure_all_details.pdf}
-    \caption{GPS velocities, model geometry, and inferend slip deficit rates.
+    \caption{GPS velocities, model geometry, and inferend slip-deficit rates.
 a) Observed profile-parallel horizontal observed velocities are shown in black with 68\% confidence errorbars.
@@
 
-A Beichuan-only fault geometry (give approximate dip and locking depth) would predict <0.1 mm/yr of surface deformation at distances >100 km away from the fault trace \citep{savage83}.
-Both the observed and predicted profile-parallel velocities (Figure \ref{fig:big_stack}a) indicate that such an estimate would appear consistent with a near-zero slip deficit rate estimate and also fail to explain the co-/post-seismic slip on the detachment \citet{Qi2011, Fielding2013b}.
+The velocity gradient across the Longmen Shan rangefront is negligible (Figure \ref{fig:big_stack}a).
+To fit these observed near-field data, a Beichuan-only fault geometry, dipping ${\sim}60^{\circ}$ at the surface and shallowing to ${\sim}10^{\circ}$ at a maximum locking depth of 20 km would require near-zero slip-deficit rates.
+Such a model is inconsistent with slip during the Wenchuan earthquake on the Beichuan and Penguan faults.
+It is also inconsistent with co- and post-seismic slip on the detachment \citep{Qi2011, Fielding2013b}.
 
@@
 When including a locked detachment, the least squares best-fit shortening rate for the Longmen Shan region is 5.7 $\pm$ 1.5 mm/yr.
-The spatial distribution of interseismic deformation is affected by the fault geometry and kinematic consistency constraints and represented in terms of the slip deficit rate which ranges shows the increase in slip-deficit from ~6.0 mm/yr on the deep detachment to 9.5 mm/yr at the surface trace of the Beichuan fault (Figure \ref{fig:big_stack}b).
-Figure \ref{fig:big_stack}c shows the horizontal velocities as a function of depth, demonstrating that the velocity gradients derive a locked to creeping transition at the detachment tip.
-The smoothing behavior of an elastic Earth prevents distinguishing between a sharp drop in slip-deficit or a gradual change over many kilometers.
-These results demonstrate a mechanism by which it is possible for interseismic strain accumulation to occur far from the range-front structures on which it is released.
+The spatial distribution of interseismic deformation is affected by the fault geometry and kinematic consistency constraints and represented in terms of the slip-deficit rate which ranges from ~6.0 mm/yr on the deep detachment to 9.5 mm/yr at the surface trace of the Beichuan fault (Figure \ref{fig:big_stack}b).
+Figure \ref{fig:big_stack}c shows the horizontal velocities as a function of depth, demonstrating that the velocity gradients derive from a locked to creeping transition at the detachment tip.
+The smoothing behavior of an elastic Earth prevents distinguishing between a sharp drop in slip-deficit and a gradual change over many kilometers.
+These results demonstrate a mechanism for interseismic strain accumulation to occur far from the range-front structures on which it is released.
 
@@
 Further, the sparsity of observations suggests that certain observations may have undue influence on the shortening estimate.
-Excluding the northwest-most observation decreases the best-fit shortening to 3.9 $\pm$ 1.9 mm/yr.
-(Add the range of slip deficit rates because the near surface Beichaun rate will still probably be close to 6 mm/yr).
+Excluding the northwest-most observation decreases the best-fit shortening to 3.9 $\pm$ 1.9 mm/yr, with a near-surface slip-deficit rate of 6.5 mm/yr on the Beichuan fault.
 The least squares uncertainty is increased (from 1.5 mm/yr to 1.9 mm/yr) because the northwestern portion of velocities are poorly constrained once the observation is excluded.
@@
 
-To quantify the sensitivity of slip deficit rate estimates to variability in data selection  we perform a complete case resampling bootstrap shortening and slip deficit rates are calculated for all possible subsets of the observed GPS velocities.
-There only $2^{20} = 1,048,576$ different possible subsets of the observations (Figure \ref{fig:distribution}).
-Subsequently, we sample jointly from this distribution and the distribution of surface-area weighted fault depths.
-This analysis demonstrates that the majority of the slip-deficit-depth probability landscape lies at slip-deficit rates between 4 and 6 mm/yr while certain parts of the geometry (the range-front) have much faster slip-deficit rates (give range of near surface rates).
-There is a non-negligible probability (give probability) assigned to slip-rates above 10 mm/yr in the steepest, near-surface portion of the fault system (Figure \ref{fig:distribution}).
-This is consistent with large near-surface coseismic slip (Quantify the max slip in any of the coseismic papers) during the Wenchuan earthquake \citep{Shen2009}.
+To further quantify the sensitivity of slip-deficit rate estimates to variability in data selection we perform a complete case resampling bootstrap shortening and slip-deficit rates are calculated for all possible subsets of the observed GPS velocities.
+There only $2^{20} = 1,048,576$ possible subsets of the observations (Figure \ref{fig:distribution}).
+We sample jointly from this distribution and the distribution of surface-area weighted fault depths.
+This analysis demonstrates that the majority of the slip-deficit-depth probability landscape lies at slip-deficit rates between 4 and 6 mm/yr.
+In the steepest, near-surface portion of the rangefront has minimum slip-deficit rates of ${\sim}$6 mm/yr, with an ${\sim}$25\% chance of slip-deficit rates above 10 mm/yr (Figure \ref{fig:distribution}).
+This is consistent with large near-surface coseismic slips of 5-7 m during the Wenchuan earthquake \citep{Xu2009, Shen2009a}.
 
@@
 \section{Implications for the Longmen Shan earthquake cycle}
-The fundamental effect of allowing a 20 km detachment to accumulated and release elastic strain through the earthquake cycle is the lack of a range-front geodetic signature despite the significant ($5-10$ mm/yr) slip deficit rate (\ref{fig:big_stack}a).
+The fundamental effect of allowing a 20 km detachment to accumulated and release elastic strain through the earthquake cycle is the lack of a range-front geodetic signature despite the significant ($5-10$ mm/yr) slip-deficit rate (\ref{fig:big_stack}a).
 Locking on the deep detachment obscures the signal of deformation on the range-front faults.
@@
 
-Prior analyses of interseismic deformation at the Longmen Shan including the effects of dipping fault system geometry and earthquake cycle effects have been limited (\citep{Loveless2011, Qi2011, Hao2014}).
-Of these only the three-dimensional block model of \citet{Loveless2011} is kinematically consistent (Jordan and Minster, McCaffrey, 2002; Meade and Hager, 2005).
-Compared to their estimated slip rate of 3.2 mm/yr, our estimate $2.5 \times$ as large.
+Prior analyses of interseismic deformation at the Longmen Shan including the effects of dipping fault system geometry and earthquake cycle effects have been limited \citep{Loveless2011, Qi2011, Hao2014}.
+Of these only the three-dimensional block model of \citet{Loveless2011} is kinematically consistent \citep{minster87,mccaffrey02,Meade2005}.
+Compared to their estimated slip rate of 3.2 mm/yr, our estimate is $2.5 \times$ as large.
 Their calculations indicate a maximum 37\% likelihood of internal block strain west of the Longmen Shan region \citep{Loveless2011}.
@@
 Our model of far-field deformation is also consistent with a similar detachment model of leveling and GPS-based uplift rates \citep{Hao2014}.
-The vertical velocities they measure increase from the Longmen Shan range front until peaking 50 km southeast of the Longriba fault zone at approximately 3.5 mm/yr of uplift.
+The vertical velocities they measure increase from the Longmen Shan rangefront until peaking 50 km southeast of the Longriba fault zone at approximately 3.5 mm/yr of uplift.
 The Longriba fault zone has the wrong slip-sense (thrusting up to the south-east, \citep{Ren2013}) to accomodate such vertical deformation.
@@
 All of these features can be explained with slip accomodated by a wide eastern Tibetan fold-and-thrust system, a suggestion supported by the continuation of shortening north from the Longmen Shan onto the Huya \citep{kirby00} and Min Jiang faults \citep{Chen1994}.
-In Figure \ref{fig:big_stack}c.
 
-While we have explained the observed the broad GPS interseismic velocity gradient as a result of a locked detachment and the Beichaun fault we have not addressed when accumulated slip-deficit on the deep detachment is released.
+While we have explained the observed the broad GPS interseismic velocity gradient as a result of a locked fault geometry including a deep detachment, we have not addressed when accumulated slip-deficit on the deep detachment is released.
 Geodetically constrained estimates of co- and post-seismic slip shows up to 5 m of slip on a deep detachment \citep{Qi2011}.
-Our data analysis depends on the detachment being locked during the period of data collection (when?).
-It is possible that in addition to coseismic and short-term postseismic activity on the detachment intermittent aseismic throughout the earthquake cycle could also be mode of strain release given that intersimsic GPS observations cover at best 6\% of a complete Longmen Shan earthquake cycle.
-Seismicity is approximately limited to the upper 20km of crust in the range-front \citep{Li2010}.
-Using steady-state critical taper wedge theory, we estimate a basal coefficient of friction between 0.02 and 0.30 (why does this vary so much?) for the detachment(cite Dahlen and give assumed pore fluid pressure and whether or not cohesion is assumed).
-The mean value of the coefficient of friction 0.05 is comparable to estimates of friction on the low-angle detachment system beneath the Longmen Shan foreland \citep{Hubbard2010}, indicating a very weak fault over long time-periods.
+Our data analysis depends on the detachment being locked during the period of data collection from 1998 to 2004 \citep{gan07}.
+In addition to coseismic and short-term postseismic activity on the detachment, it is possible that intermittent aseismic slip could be an important mode of strain release.
+This is difficult to assess given that intersimsic GPS observations cover, at best, 6\% of a complete Longmen Shan earthquake cycle.
+Seismicity is approximately limited to the upper 20 km of crust in the range-front \citep{Li2010a}.
+Using steady-state critical taper wedge theory \citep{dahlen90}, with a internal coefficient of friction of 0.7, a cohesion of 5 MPa, and a hydrostatic fluid pressure, we find a basal friction coefficient of 0.08.
+This is comparable to estimates of friction on the low-angle detachment system beneath the Longmen Shan foreland \citep{Hubbard2010}, indicating a very weak fault over long time-periods.
 On the other hand, the surface area and shortening rate of the detachment could sustain very large earthquakes if slip is released coseismically.
-Specifically, if 100km down-dip and 300km along-strike were to slip 4 meters, similar to average slip during the Wenchuan earthquake, a $M_{\textrm{w}}$ 8.4 rupture would be produced.
+Specifically, if 100 km down-dip and 300 km along-strike were to slip 4 meters, similar to average slip during the Wenchuan earthquake, a $M_{\textrm{w}}$ 8.4 rupture would be produced.
 
 We also quantify the effect of explicitly treating surface topography by comparing models with and without surface topography, the local minimum in velocity (Figure \ref{fig:big_stack}a) northwest of the Beichuan is identified as the main topographically-influenced feature.
-The effects may be more pronounced with a shallower fault dip like in a subduction zones.
-Despite the mediocre effect (~15\%), geodetic analyses in steep regions should consider topographic effects to avoid systematic bias and these effects may be even more signficant for coseismic faulting cases where deformation is localized in regions of steep topography.
+The effects may be more pronounced with a shallower fault dip (e.g.
+many subduction zones).
+Despite the mediocre effect (~15\%), geodetic analyses in steep regions should consider topographic effects to avoid systematic bias.
+The effect may be even more signficant for coseismic faulting where deformation is localized in regions of steep topography.
 
-The 6-9 mm/yr slip deficit rates estimated here can be used to provide new constraints on earthquake recurrence intervals at the Longmen Shan (Figure \ref{fig:hazard}).
-Our shortening estimate suggests a recurrence interval for Wenchuan-like ruptures of approximately 600 years while recurrence intervals for smaller events would be more frequent.
+The 6-9 mm/yr slip-deficit rates estimated here can be used to provide new constraints on earthquake recurrence intervals at the Longmen Shan (Figure \ref{fig:hazard}).
+Our shortening estimate suggests a recurrence interval for Wenchuan-like ruptures of approximately 600 years, with much shorter recurrence intervals for smaller events.
 This contrasts with the average recurrence of ~2000 years inferred from paleoseismic field investigations for the Beichuan fault \citep{Ran2010}.
-The 6-9 mm/ry slip deficit rate estimated here would imply $M_{\textrm{w}} = $9.0 ruptures for that recurrence interval, which would seemingly require a along struke rupture equal to or greater than the $\sim 300$ km along strike length of the Longmen Shan, our slip deficit estimate could include the effects of other imbricated structures and thus serving as a system-wide estimate while paleoseismic estimates have been focused on the analysis of individual thrust faults.
+The 6-9 mm/yr slip-deficit rate estimated here would imply $M_{\textrm{w}} = $9.0 ruptures for that recurrence interval, which would seemingly require an along strike rupture equal to or greater than the ${\sim}300$ km along strike length of the Longmen Shan.
+However, our slip-deficit estimate could include the effects of other imbricated structures.
+As a result our estimate is a system-wide estimate while paleoseismic estimates have focused on the analysis of individual thrust faults.
 
-The earthquake recurrence intervals calculated above may be considered a lower bound for three reasons.
-First, coseismic, interseismic, and long-term slip estimates include a significant component of strike-slip motion \citep{Shen2009, Qi2011, Densmore2007} (Meade2007, Loveless and Meade, 2011).
-while we consider an idealized case of fault normal motion only.
+The earthquake recurrence intervals calculated above may be considered a lower bound.
+First, coseismic, interseismic, and geologic slip estimates include a significant component of strike-slip motion \citep{Shen2009a, Qi2011, Densmore2007, Meade2007, Loveless2011}, while we consider an idealized case of fault normal motion only.
 Second, our two-dimensional model assumes plane strain conditions which will result in under-estimates of the shortening required to produce far-field velocities.
-This is because the assumption is equivalent to assuming the fault geometry extends infinitely far along strike.
+This is because the plane strain simplification is equivalent to assuming the fault geometry extends infinitely far along strike.
 Third, the modeled thrust is the steeply dipping Beichuan fault.
 The foreland Range Front thrust or the shallow detachment beneath Chengdu dip more shallowly \citep{Hubbard2010}, producing a greater total moment deficit.
-Finally, the slip deficit rate estimates may be low due to being estimated late in the earthquake cycle (i.e., in the decade prior to the Wenchuan earthquake) when viscoalstic relaxation effects may further contribute to a lowe near fault trace velocity gradient e.g., \citep{savage00}.
+Finally, the slip-deficit rate estimates may be low due to being estimated late in the earthquake cycle (i.e., in the decade prior to the Wenchuan earthquake) when viscoelastic effects may further contribute to a low near fault trace velocity gradient \citep{savage00}.
 
@@
 \section{Conclusion}
-Previous inferences of near-zero interseismic shortening estimates at the Longmen Shan have been difficult to reconcile the Longmen Shan are at odds with the large-moment of the Wenchuan earthquake, very steep topography, structural interpretations indicate fold-and-thrust geometry, fast erosion rates.
+Previous inferences of near-zero interseismic shortening estimates at the Longmen Shan have been difficult to reconcile with the large-moment of the Wenchuan earthquake, very steep topography, structural interpretations indicating fold-and-thrust geometry, and fast erosion rates.
 We offer a resolution to this conflict by including earthquake cycle effects associated with interseismic locking on the Beichuan fault and a 20 km deep detachment to the west.
-Our revised interseismic shortening rate is 5.7 $\pm$ 1.5 mm/yr with on fault slip-deficit rates up to 9.5 mm/yr near the surface, where the Beichuan fault is most steeply dipping.
+Our revised interseismic shortening rate is 5.7 $\pm$ 1.5 mm/yr with on-fault slip-deficit rates up to 9.5 mm/yr near the surface, where the Beichuan fault is most steeply dipping.
 The model demonstrates that interseismic strain accumulation may be spatially disjoint from coseismic strain release, emphasizing that accurate fault system geometries are critical in the analysis of geodetic data.
-This boundary element model presented here unifies geodetic models of coseismic, postseismic and interseismic crustal deformation across the greater Longmen Shan region and suggests that, with current loading rates, Wenchuan-like earthquakes may recurrence as frequently as every 600 years.
+The boundary element model presented here unifies geodetic models of coseismic, postseismic and interseismic crustal deformation across the greater Longmen Shan region and suggests that, with current loading rates, Wenchuan-like earthquakes may occur as frequently as every 600 years.
 

### 2014-10-13 paper.tex
BJM edits on 10/13/14

@@
 In contrast, we find that interseismic GPS data are consistent with a shortening rate of 5.7$\pm$1.5 mm/yr and maximum surface slip-deficit rate of 9.5 $\pm$ 2.5 mm/yr.
-This model provides a unified interpretation of geodetic deformation throughout the earthquake cycle and suggests that the Longmen Shan is an active fold-and-thrust belt with Wenchuan style earthquake recurrence intervals of $<$600 years.
+This model provides a unified interpretation of geodetic deformation throughout the earthquake cycle and suggests that the Longmen Shan is an active fold-and-thrust belt with Wenchuan style earthquake recurrence intervals of as short as 600 years.
 
@@
 
-Previous estimates of geodetically constrained shortening and slip-deficit rates in the Longmen Shan region can be categorized into those that either: 1) do not explicitly treat earthquake cycle effects, \citep{chen00, shen05, Thatcher2007} or 2) those that use block models, with a first order approximation of earthquake cycle effects but coarse representations of fault system geometry \citep{Meade07c, Loveless2011, Burchfiel2008}.
+Previous estimates of geodetically constrained shortening and slip-deficit rates in the Longmen Shan region can be categorized into those that either: 1) do not explicitly treat earthquake cycle effects, \citep{chen00, shen05, Thatcher2007} or 2) those that use block models, with a first order quasi-static approximation of earthquake cycle effects but coarse representations of fault system geometry \citep{Meade07c, Loveless2011, Burchfiel2008}.
 The inferred slip-deficit rates range from 0.0 mm/yr \citep{Thatcher2007} to 3.2 mm/yr \citep{Loveless2011}, with all analyses reporting significant strain 100-200 km northwest of the Longmen Shan in a region termed the Songpan-Xihe deformation zone \citep{shen05}.
@@
 
-\section{Geodetic analysis of Longmen Shan shortening}
+\section{Geodetic analysis of Longmen Shan shortening and slip deficit rates}
 The geometry of the Beichaun fault and the detachment to the west have been inferred from 1) structural interpretations  \citep{Hubbard2010} (Plesch and Shaw, personal communication), 2) seismic relection observations \citep{Zhang2009, Guo2013}, and 3) geodetic modeling of co- and post-seismic deformation associated with the Wenchuan earthquake \citep{Qi2011, Fielding2013b}.
@@
 The fault system geometry consists of two main segments (Figure \ref{fig:regional_map}).
-The range-front Beichuan fault dips steeply at ${\sim}60^{\circ}$ at its surface trace shallowing to sub-horizontal near 20 km depth where it joins the detachment surface at depth.
-The detachment dips at ${\sim}1.5^{\circ}$ from an initial depth of 20 km for 110 km to the northwest at its deepest point with a depth of 23 km.
+The range-front Beichuan fault dips steeply at ${\sim}60^{\circ}$ at its surface trace shallowing to sub-horizontal ${\sim}20$ km depth where it merges with the detachment surface.
+The detachment dips at ${\sim}1.5^{\circ}$ from an initial depth of 20 km and extends 110 km to the northwest to a depth of 23 km.
 For our interseismic shortening and slip-deficit models we use an idealized two-dimensional representation of these fault surfaces (Figure \ref{fig:regional_map}).
 
-The inference that a 20km deep detachment slipped, either co- or post-seismically, at distances up to 100km from the range-front \citep{Qi2011}, necessitates including interseismic GPS velocities at a similar distance into the interior of the Tibetan Plateau (Figure \ref{fig:regional_map}).
-We analyze interseismic GPS velocities measured during the decade prior to the Wenchuan earthquake \citep{gan07}, in a nominally Eurasian reference frame \citet{apel06}.
+The inference that a 20 km deep detachment slipped, either co- or post-seismically, at distances up to 100 km from the range-front \citep{Qi2011}, necessitates including interseismic GPS velocities at a similar distance into the interior of the Tibetan Plateau (Figure \ref{fig:regional_map}).
+We analyze interseismic GPS velocities measured during the decade prior to the Wenchuan earthquake \citep{gan07}, in a nominally Eurasian reference frame \citep{apel06}.
 We exclude velocities that are near other major active structures including the Xianshuihe and Kunlun faults (Figure \ref{fig:regional_map}).
@@
 This approach is similar to analytical limit-to-the-boundary methods \citep{Sutradhar2008} and series-expansion methods \citep{Klockner2013}.
-Our tests of relative error show that this quadrature-based method can match, to arbitrary precision, analytically derived half-space solutions and is free of the unpleasant singularities present in those solutions.
-This entirely numerical boundary element method allows for complete flexibility in surface geometry, including topography.
+Our tests of relative error show that this quadrature-based method can match, to arbitrary precision, analytically derived half-space solutions and is free of the singularities present in those solutions.
+This entirely numerical boundary element method allows for complete flexibility in representing both variable surface topography and complex fault system geometeries.
 The steep topography of the Longmen Shan rangefront is included via the SRTM30\_PLUS digital elevation dataset \citep{Becker2009}.
@@
 Using this boundary element model, we solve for the single best-fitting shortening rate between the Tibetan Plateau and the Sichuan Basin while maintaining kinematic consistency using a two-dimensional block formulation.
-In the block formulation, the slip-deficit rate on the fault, $s_\mathrm{sd}$, is related to the local fault dip, $\delta$, and total convergence rate, $v_0$, by $s_\mathrm{sd}=v_0 / cos\delta$ \citep{mccaffrey02, Meade2005}.
-As a result, slip-deficit rates are strictly larger than shortening rates.
+In the block formulation, the slip-deficit rate on the fault, $s_\mathrm{sd}$, is related to the local fault dip, $\delta$, and total convergence rate, $v_0$, by $s_\mathrm{sd}=v_0 / \cos\delta$ \citep{mccaffrey02, Meade2005}.
+As a result, slip-deficit rates are strictly larger than shortening rates to maintain smoothly varying horizontal surface velocities.
 The average interseismic slip-deficit on the fault geometry is subtracted from block motion to calculate the interseismic velocity profile \citep{savage83, Meade2005}.
@@
 The smoothing behavior of an elastic Earth prevents distinguishing between a sharp drop in slip-deficit and a gradual change over many kilometers.
-These results demonstrate a mechanism for interseismic strain accumulation to occur far from the range-front structures on which it is released.
+These results demonstrate a mechanism for a distinct interseismic strain accumulation to be most clearly manifest far from the range-front structures on which coseismic slip is greatest.
 
@@
 Further, the sparsity of observations suggests that certain observations may have undue influence on the shortening estimate.
-Excluding the northwest-most observation decreases the best-fit shortening to 3.9 $\pm$ 1.9 mm/yr, with a near-surface slip-deficit rate of 6.5 mm/yr on the Beichuan fault.
+Excluding the northwest-most observation decreases the best-fit shortening to 3.9 $\pm$ 1.9 mm/yr, with a near-surface slip-deficit rate of 6.5 (give uncertainty) mm/yr on the Beichuan fault.
 The least squares uncertainty is increased (from 1.5 mm/yr to 1.9 mm/yr) because the northwestern portion of velocities are poorly constrained once the observation is excluded.
@@
 In the steepest, near-surface portion of the rangefront has minimum slip-deficit rates of ${\sim}$6 mm/yr, with an ${\sim}$25\% chance of slip-deficit rates above 10 mm/yr (Figure \ref{fig:distribution}).
-This is consistent with large near-surface coseismic slips of 5-7 m during the Wenchuan earthquake \citep{Xu2009, Shen2009a}.
+This geodetically constrained slip deficit rate can provide a mechanism for the rapid loading of the faults which exhibited 5-7 m of near-surface coseismic slip during the Wenchuan earthquake \citep{Xu2009, Shen2009a}.
 
@@
 In addition to coseismic and short-term postseismic activity on the detachment, it is possible that intermittent aseismic slip could be an important mode of strain release.
-This is difficult to assess given that intersimsic GPS observations cover, at best, 6\% of a complete Longmen Shan earthquake cycle.
+This is difficult to observationally constrain given that intersimsic GPS observations cover, at best, 6\% of a complete Longmen Shan earthquake cycle.
 Seismicity is approximately limited to the upper 20 km of crust in the range-front \citep{Li2010a}.
@@
 On the other hand, the surface area and shortening rate of the detachment could sustain very large earthquakes if slip is released coseismically.
-Specifically, if 100 km down-dip and 300 km along-strike were to slip 4 meters, similar to average slip during the Wenchuan earthquake, a $M_{\textrm{w}}$ 8.4 rupture would be produced.
+Specifically, if 100 km down-dip and 300 km along-strike were to slip 4 meters, similar to average slip during the Wenchuan earthquake, a $\textrm{M}_{\textrm{w}}$ 8.4 rupture would be produced.
 
 We also quantify the effect of explicitly treating surface topography by comparing models with and without surface topography, the local minimum in velocity (Figure \ref{fig:big_stack}a) northwest of the Beichuan is identified as the main topographically-influenced feature.
-The effects may be more pronounced with a shallower fault dip (e.g.
-many subduction zones).
-Despite the mediocre effect (~15\%), geodetic analyses in steep regions should consider topographic effects to avoid systematic bias.
-The effect may be even more signficant for coseismic faulting where deformation is localized in regions of steep topography.
+The effects may be more pronounced with a shallower fault dip (e.g., many subduction zones).
+Despite the mediocre effect (${\sim}$15\%), geodetic analyses in steep regions should consider topographic effects to avoid systematic bias.
+This effect may be even more pronounced for coseismic faulting where deformation is localized in regions of steep topography.
 
-The 6-9 mm/yr slip-deficit rates estimated here can be used to provide new constraints on earthquake recurrence intervals at the Longmen Shan (Figure \ref{fig:hazard}).
+The 6-9 mm/yr slip-deficit rates estimated here can be used to provide a new kinematic constraint on earthquake recurrence intervals at the Longmen Shan (Figure \ref{fig:hazard}).
 Our shortening estimate suggests a recurrence interval for Wenchuan-like ruptures of approximately 600 years, with much shorter recurrence intervals for smaller events.
 This contrasts with the average recurrence of ~2000 years inferred from paleoseismic field investigations for the Beichuan fault \citep{Ran2010}.
-The 6-9 mm/yr slip-deficit rate estimated here would imply $M_{\textrm{w}} = $9.0 ruptures for that recurrence interval, which would seemingly require an along strike rupture equal to or greater than the ${\sim}300$ km along strike length of the Longmen Shan.
+The 6-9 mm/yr slip-deficit rate estimated here would imply $\textrm{M}_{\textrm{w}}$ 9.0 ruptures for that recurrence interval, which would seemingly require an along strike rupture equal to or greater than the ${\sim}300$ km along strike length of the Longmen Shan.
 However, our slip-deficit estimate could include the effects of other imbricated structures.
@@
 The gray region indicates the uncertainty implied by the error in our shortening estimate.
-The discrepancy between the Wenchuan-like recurrence and the paleoseismic estimate may be due to the presence of multiple active faults in the range-front.}
+The discrepancy between the Wenchuan-like recurrence and the paleoseismic estimate may be due to the presence of multiple active faults in the range-front.
+(BJM note: Fixe the Ms in the figure so that they are not italicized and consistent with other in the text)}
     \label{fig:hazard}
@@
 Previous inferences of near-zero interseismic shortening estimates at the Longmen Shan have been difficult to reconcile with the large-moment of the Wenchuan earthquake, very steep topography, structural interpretations indicating fold-and-thrust geometry, and fast erosion rates.
-We offer a resolution to this conflict by including earthquake cycle effects associated with interseismic locking on the Beichuan fault and a 20 km deep detachment to the west.
-Our revised interseismic shortening rate is 5.7 $\pm$ 1.5 mm/yr with on-fault slip-deficit rates up to 9.5 mm/yr near the surface, where the Beichuan fault is most steeply dipping.
+We offer a possible resolution to this conflict by including earthquake cycle effects associated with interseismic locking on the Beichuan fault and a 20 km deep detachment to the west.
+Our revised interseismic shortening rate is 5.7 $\pm$ 1.5 mm/yr with on-fault slip-deficit rates up to 9.5 mm/yr near the surface where the Beichuan fault is most steeply dipping.
 The model demonstrates that interseismic strain accumulation may be spatially disjoint from coseismic strain release, emphasizing that accurate fault system geometries are critical in the analysis of geodetic data.

### 2014-10-14 paper.tex
Quick local.

--- 2014-10-12 paper.tex
@@
 \section{Implications for the Longmen Shan earthquake cycle}
-The fundamental effect of allowing a 20 km detachment to accumulated and release elastic strain through the earthquake cycle is the lack of a range-front geodetic signature despite the significant ($5-10$ mm/yr) slip-deficit rate (\ref{fig:big_stack}a).
+The fundamental effect of allowing a 20 km detachment to accumulate and release elastic strain through the earthquake cycle is the lack of a range-front geodetic signature despite the significant ($5-10$ mm/yr) slip-deficit rate (\ref{fig:big_stack}a).
 Locking on the deep detachment obscures the signal of deformation on the range-front faults.
@@
 Seismicity is approximately limited to the upper 20 km of crust in the range-front \citep{Li2010a}.
-Using steady-state critical taper wedge theory \citep{dahlen90}, with a internal coefficient of friction of 0.7, a cohesion of 5 MPa, and a hydrostatic fluid pressure, we find a basal friction coefficient of 0.08.
+Using steady-state critical taper wedge theory \citep{dahlen90}, with an internal coefficient of friction of 0.7, a cohesion of 5 MPa, and a hydrostatic fluid pressure, we find a basal friction coefficient of 0.08.
 This is comparable to estimates of friction on the low-angle detachment system beneath the Longmen Shan foreland \citep{Hubbard2010}, indicating a very weak fault over long time-periods.
@@
 However, our slip-deficit estimate could include the effects of other imbricated structures.
-As a result our estimate is a system-wide estimate while paleoseismic estimates have focused on the analysis of individual thrust faults.
+As a result our estimate is system-wide while paleoseismic estimates have focused on the analysis of individual thrust faults.
 

### 2014-10-19 paper.tex
Cover letter. Cleaned up bibliography.

@@
 \maketitle
+
+\section{Cover Letter}
+Dear Geology editors,
+
+Please find attached ``Rapid shortening at the eastern margin of the Tibetan plateau prior to the 2008 $M_\textrm{w}$ 7.9 Wenchuan earthquake'' by T.
+Ben Thompson and Brendan J.
+Meade.
+We are submitting this article for consideration as an article in Geology.
+The manuscript contains 3249 words and 4 figures.
+(TBT Note: According to the Geology website, we are supposed to indicate whether we are willing to pay for color figures.)
+
+The 2008 Wenchuan earthquake emphasized that the dicussion over tectonics in eastern Tibet has crucial implications for seismic hazard.
+Geodetic slip-deficit rates are a primary way of translating tectonic models into concrete hazard estimates.
+But, estimates are dependent on assumptions of fault geometry and consideration of earthquake cycle effects.
+Recent observations of co- and post-seismic slip on a deep detachment suggest that Longmen Shan regional fault geometry may be more complex than previously thought.
+Using a boundary element model, we analyze the implications of a interseismically locked detachment beneath eastern Tibet and show that geometry can produce a velocity gradient that is steepest 100-200 km northwest of the Longmen Shan rangefront, exactly as observed, with almost no velocity gradient at the rangefront.
+Inferring from GPS velocities measured before the earthquake, the resulting slip-deficit estimate is as high as 10 mm/yr at the steepest parts of the rangefront Beichuan fault.
+These results show a simple analysis of the GPS velocities across the Longmen Shan will produce dramatically different results than an analysis including more detailed fault geometry.
+
+These results are of wider significance in that they demonstrate the impact of fault geometry on the analysis of fault systems, geodetic observation of a detachment participating in the earthquake cycle, and the effects of surface topography on geodesy.
+We are optimistic that these results will contribute to the understanding of the eastern Tibetan Plateau.
+The active tectonics and seismic hazard of eastern Tibet and the Longmen Shan are addressed by other publications in Geology (Clark and Royden 2000, Meade 2007, OTHERS?), so the journal is a natural arena to continue the discussion.
+
+Thank you for considering this paper for publication. We look forward to hearing from you.
+
+T. Ben Thompson, Graduate Student
+
+Department of Earth \& Planetary Sciences, Harvard University
+
+Cambridge, MA 02138
+
+(617)-NNN-NNNN, tthompson@fas.harvard.edu
 
@@
 In the steepest, near-surface portion of the rangefront has minimum slip-deficit rates of ${\sim}$6 mm/yr, with an ${\sim}$25\% chance of slip-deficit rates above 10 mm/yr (Figure \ref{fig:distribution}).
-This geodetically constrained slip deficit rate can provide a mechanism for the rapid loading of the faults which exhibited 5-7 m of near-surface coseismic slip during the Wenchuan earthquake \citep{Xu2009, Shen2009a}.
+This geodetically constrained slip deficit rate can provide a mechanism for the rapid loading of the faults which exhibited 5-7 m of near-surface coseismic slip during the Wenchuan earthquake \citep{Xu2009, Shen2009b}.
 
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{Jcks}
 \bibliography{biblio,library}

### 2014-10-27 paper.tex
BJM edits to cover letter and a bit more

--- 2014-10-14 paper.tex
@@
 
+\section{Cover Letter}
+Dear Geology editors,
+
+Please find attached a manuscript entitled ``Rapid shortening at the eastern margin of the Tibetan plateau prior to the 2008 $M_\textrm{w}$ 7.9 Wenchuan earthquake'' by T.
+Ben Thompson and Brendan J.
+Meade.
+We are submitting this manuscript for consideration as an article in Geology.
+The manuscript contains 3249 words and 4 figures and we are happy to pay for color figures.
+
+The 2008 $M_\textrm{w}$ 7.9 Wenchuan earthquake led to the loss of more than 60,000 lives and prompted a fundamental reconsideration of how we understand the active tectonics of the remarkably steep Longmen Shan rangefront at the eastern margin of the Tibet Plateau.
+In particular, this seismic event has brought into sharp relief the debate an understanding of the rangefront as a active fold-and-thrust belt or as the passive response to the deeper flow of lower crustal material.
+
+Geodetic slip and slip-deficit rates are a primary way of quantifying fault system kinematics and providing constraints on earthquake recurrence intervals.
+Previous estimates of slip rates that the Longmen Shan have been exceptionally crude (and serve only as a lower bound) due to the neglect of earthquake cycle effects and .
+Recent observations of co- and post-seismic slip on a deep detachment suggest that Longmen Shan regional fault geometry may be more complex than previously thought.
+Using a new boundary element model including the effects of both topography and non-planar fault system geometry, we analyze the implications of a interseismically locked detachment beneath eastern Tibet and show that geometry can produce a velocity gradient that is steepest 100-200 km northwest of the Longmen Shan rangefront, exactly as geodetically observed, with a negligible velocity gradient at the steepest part of the rangefront.
+Interpreting GPS velocities measured prior to the Wenchuan earthquake, inferred slip-deficit rates on the Beichuan fault may be as high as 10 mm/yr at the surface.
+These results show a simple analysis of the GPS velocities across the Longmen Shan will produce dramatically different results than an analysis including more detailed fault geometry.
+
+These results are of wide significance for three reasons: 1) Providing a new interpretation of tectonics and earthquake cycle physics at the Longmen Shan, 2) Demonstrating how topography can be explicitly incorporated into earthquake cycle problems, and 3) Illustrating the complex role the deep detachment faults can play in controlling earthquake cycle behaviors.
+We are optimistic that these results will not only contribute significantly to the debate about the tectonics of the Tibetan Plateau and, more generally, about the nature of the development of fold-and-thrust belts.
+Geology has served as the journal for the greatest insights on the  active tectonics and seismic hazard of eastern Tibet (King et al,.
+1997; Clark and Royden, 2000; Zhang et al., 2004; Clark et al., 2005; Meade, 2007; Xu et al., 2009; Zhao et al., 2012; Wang et al., 2014) and our work offers a substantial conceptual and quantitative revisions of both previous tectonic concepts and earthquake recurrence intervals at this most active of margins.
+
+Thanks for your thoughtful consideration of this paper. We're looking forward to hearing from you soon!
+
+T. Ben Thompson
+
+Department of Earth \& Planetary Sciences, Harvard University
+
+Cambridge, MA 02138
+
+(617)-NNN-NNNN, tthompson@fas.harvard.edu
+
 \section{Abstract}
-The Longmen Shan is the steepest topographic front of the India-Asia collision and was the site of the $\textrm{M}_{\textrm{w}}$ 7.9  Wenchuan earthquake.
+The Longmen Shan is the steepest topographic front assocated with the India-Asia collision and was the site of the $\textrm{M}_{\textrm{w}}$ 7.9  Wenchuan earthquake.
 Shortening estimates across the Longmen Shan provide strain accumulation rates and clarify the eastward extrusion of the Tibetan plateau.

### 2014-10-27 paper.tex
Cite Cooke.

--- 2014-10-19 paper.tex
@@
 An important feature of our tool is the inclusion of accurate surface topography in steep mountainous regions or Earth curvature in large-scale problems.
-Previous boundary element methods in the earth sciences \citep{Thomas1993, Maerten2014a} have been based on analytic integrations of rectangular or triangular slip surfaces \citep{Okada1992, Meade2007}.
+Previous boundary element methods in the earth sciences \citep{Thomas1993, Cooke2011, Maerten2014a} have been based on analytic integrations of rectangular or triangular slip surfaces \citep{Okada1992, Meade2007}.
 These analytic formulae normally assume an infinite half-space geometry and piecewise constant slip.

### 2014-10-27 paper.tex
Moved cover letter out of paper.tex. Edits. Changed title.

--- 2014-10-27 paper.tex
@@
 
-\title{Rapid shortening at the eastern margin of the Tibetan plateau prior to the 2008 $M_\textrm{w}$ 7.9 Wenchuan earthquake}
+\title{Rapid slip-deficit rates at the eastern margin of the Tibetan plateau prior to the 2008 $\textrm{M}_\textrm{w}$ 7.9 Wenchuan earthquake}
 \author{T. Ben Thompson, Brendan J. Meade}
@@
 
-\section{Cover Letter}
-Dear Geology editors,
-
-Please find attached ``Rapid shortening at the eastern margin of the Tibetan plateau prior to the 2008 $M_\textrm{w}$ 7.9 Wenchuan earthquake'' by T.
-Ben Thompson and Brendan J.
-Meade.
-We are submitting this article for consideration as an article in Geology.
-The manuscript contains 3249 words and 4 figures.
-(TBT Note: According to the Geology website, we are supposed to indicate whether we are willing to pay for color figures.)
-
-The 2008 Wenchuan earthquake emphasized that the dicussion over tectonics in eastern Tibet has crucial implications for seismic hazard.
-Geodetic slip-deficit rates are a primary way of translating tectonic models into concrete hazard estimates.
-But, estimates are dependent on assumptions of fault geometry and consideration of earthquake cycle effects.
-Recent observations of co- and post-seismic slip on a deep detachment suggest that Longmen Shan regional fault geometry may be more complex than previously thought.
-Using a boundary element model, we analyze the implications of a interseismically locked detachment beneath eastern Tibet and show that geometry can produce a velocity gradient that is steepest 100-200 km northwest of the Longmen Shan rangefront, exactly as observed, with almost no velocity gradient at the rangefront.
-Inferring from GPS velocities measured before the earthquake, the resulting slip-deficit estimate is as high as 10 mm/yr at the steepest parts of the rangefront Beichuan fault.
-These results show a simple analysis of the GPS velocities across the Longmen Shan will produce dramatically different results than an analysis including more detailed fault geometry.
-
-These results are of wider significance in that they demonstrate the impact of fault geometry on the analysis of fault systems, geodetic observation of a detachment participating in the earthquake cycle, and the effects of surface topography on geodesy.
-We are optimistic that these results will contribute to the understanding of the eastern Tibetan Plateau.
-The active tectonics and seismic hazard of eastern Tibet and the Longmen Shan are addressed by other publications in Geology (Clark and Royden 2000, Meade 2007, OTHERS?), so the journal is a natural arena to continue the discussion.
-
-Thank you for considering this paper for publication. We look forward to hearing from you.
-
-T. Ben Thompson, Graduate Student
-
-Department of Earth \& Planetary Sciences, Harvard University
-
-Cambridge, MA 02138
-
-(617)-NNN-NNNN, tthompson@fas.harvard.edu
-
 \section{Abstract}
-The Longmen Shan is the steepest topographic front of the India-Asia collision and was the site of the $\textrm{M}_{\textrm{w}}$ 7.9  Wenchuan earthquake.
+The Longmen Shan is the steepest topographic front assocated with the India-Asia collision and was the site of the $\textrm{M}_{\textrm{w}}$ 7.9  Wenchuan earthquake.
 Shortening estimates across the Longmen Shan provide strain accumulation rates and clarify the eastward extrusion of the Tibetan plateau.
@@
 In the steepest, near-surface portion of the rangefront has minimum slip-deficit rates of ${\sim}$6 mm/yr, with an ${\sim}$25\% chance of slip-deficit rates above 10 mm/yr (Figure \ref{fig:distribution}).
-This geodetically constrained slip deficit rate can provide a mechanism for the rapid loading of the faults which exhibited 5-7 m of near-surface coseismic slip during the Wenchuan earthquake \citep{Xu2009, Shen2009b}.
+This geodetically constrained slip deficit rate can provide a mechanism for the rapid loading of the faults which exhibited 5-7 m of near-surface coseismic slip during the Wenchuan earthquake \citep{Xu2009, Shen2009a}.
 
@@
 The gray region indicates the uncertainty implied by the error in our shortening estimate.
-The discrepancy between the Wenchuan-like recurrence and the paleoseismic estimate may be due to the presence of multiple active faults in the range-front.
-(BJM note: Fixe the Ms in the figure so that they are not italicized and consistent with other in the text)}
+The discrepancy between the Wenchuan-like recurrence and the paleoseismic estimate may be due to the presence of multiple active faults in the range-front.}
     \label{fig:hazard}

### 2014-11-14 paper.tex
Incorporated JS edits.

@@
 \title{Rapid slip-deficit rates at the eastern margin of the Tibetan plateau prior to the 2008 $\textrm{M}_\textrm{w}$ 7.9 Wenchuan earthquake}
-\author{T. Ben Thompson, Brendan J. Meade}
-\date{September 2014}
+\author{T. Ben Thompson, Brendan J. Meade -- where to put JS, AP?}
 \usepackage{graphicx}
@@
 The Longmen Shan range is the steepest margin of the Tibetan Plateau rising 5 km over a distance of only 30 km at the western edge of the Sichuan Basin.
-In 2008, the $\textrm{M}_{\textrm{w}}$ 7.9 Wenchuan earthquake ruptured the Beichan and Penguan rangefront thrusts ${\sim}250$ km along strike, leading to approximately 70,000 fatalities.
+In 2008, the $\textrm{M}_{\textrm{w}}$ 7.9 Wenchuan earthquake ruptured the Beichan and Penguan rangefront thrusts ${\sim}250$ km along strike, leading to approximately 70,000 fatalities \citep{Hubbard2009, Xu2009, Lin2009}.
 The energy release associated with the Wenchuan earthquake requires a complimentary period of strain accumulation.
@@
 This apparent lack of present-day shortening across the steepest topography of the India-Asia collision has been hypothesized to result from lower crustal inflation tectonic model for the eastern Tibetan Plateau \citep{royden97, bird91, Burchfiel2008, Clark2000}.
-But, these models do not provide a mechanism for accumulating the elastic strain released coseismically during the Wenchuan earthquake.
+But, these models do not provide a mechanism for accumulating the elastic strain released coseismically during the Wenchuan earthquake, nor do they explain the significant crustal shortening and active geologic structures along the range front \citep{Hubbard2010, Li2010a, Wang2013, Wang2014}.
 Recent observations of co- and post-seismic slip on a deep detachment to the west of the Beichuan fault\citep{Qi2011, Fielding2013b} motivate a new model of Longmen Shan interseismic shortening that includes a locked detachment.
@@
 Here, we present results from a two-dimensional boundary element model that includes earthquake cycle and topographic effects to explain the regional geodetic velocities as the result of interseismic locking on both the range-front Beichuan fault and a 20 km deep detachment beneath the hinterland.
-This fault system geometry is consistent with the inference of 2-6 m of slip of a 20 km deep detachment that extends 90 km northwest using InSAR measurements and post-Wenchuan GPS observations \citep{Qi2011, Fielding2013b} and interpretation of structural observations \citep{Hubbard2010, Li2010a}.
+This fault system geometry is consistent with structural observations \citep{Hubbard2010, Li2010a} and the inference of 2-6 m of slip of a 20 km deep detachment that extends 90 km northwest using InSAR measurements and post-Wenchuan GPS observations \citep{Qi2011, Fielding2013b}.
 Integrating rangefront and detachment fault system geometry, we develop a model of interseismic strain accumulation that is consistent with both interseismic GPS velocities measured prior to the Wenchuan earthquake and previous observations of co- and post-seismic strain release at depth.
@@
 \section{Geodetic analysis of Longmen Shan shortening and slip deficit rates}
-The geometry of the Beichaun fault and the detachment to the west have been inferred from 1) structural interpretations  \citep{Hubbard2010} (Plesch and Shaw, personal communication), 2) seismic relection observations \citep{Zhang2009, Guo2013}, and 3) geodetic modeling of co- and post-seismic deformation associated with the Wenchuan earthquake \citep{Qi2011, Fielding2013b}.
-The depth and dip of the detachment are similar to the best-fitting surface found by \citet{Qi2011}.
-The fault system geometry consists of two main segments (Figure \ref{fig:regional_map}).
-The range-front Beichuan fault dips steeply at ${\sim}60^{\circ}$ at its surface trace shallowing to sub-horizontal ${\sim}20$ km depth where it merges with the detachment surface.
-The detachment dips at ${\sim}1.5^{\circ}$ from an initial depth of 20 km and extends 110 km to the northwest to a depth of 23 km.
+The geometry of the Beichuan fault and the detachment to the west has been inferred from structural interpretations \citep{Hubbard2010} and seismic reflections observations \citep{Zhang2009, Guo2013}.
+The Beichuan fault is part of an imbricate thrust system, with fault dips steepening to the northwest.
+Structural imbrication is typically produced by stacked listric faults that sole to detachments (TODO: CITE?).
+Furthermore, the large synclinorium parallel to, and west of, the Beichuan fault is consistent with the Beichuan fault shallowing to a detachment based on fault-bend folding theory \citep{Hubbard2010, Suppe1983}.
+In our model, the range-front Beichuan fault dips steeply at ${\sim}60^{\circ}$ at its surface trace shallowing to sub-horizontal ${\sim}20$ km depth where it merges with the detachment surface.
+At seismogenic depths, a 35$^\circ$ dip has been confirmed via focal mechanisms \citep{Hubbard2010}.
+Extrapolating from the regional dip of the Sichuan Basin as it extends beneath the Longmenshan range front \citep{Hubbard2010, Guo2013}, the detachment in our model dips ${\sim}1.5^{\circ}$ from an initial depth of 20km to a depth of 23km at a distance of 110km to the northwest.
+The depth, dip, and extent of the detachment are similar to the best-fitting surface found by \citet{Qi2011} in fitting the co- and post-seismic slip of the Wenchuan earthquake.
 For our interseismic shortening and slip-deficit models we use an idealized two-dimensional representation of these fault surfaces (Figure \ref{fig:regional_map}).
@@
 The velocity gradient across the Longmen Shan rangefront is negligible (Figure \ref{fig:big_stack}a).
-To fit these observed near-field data, a Beichuan-only fault geometry, dipping ${\sim}60^{\circ}$ at the surface and shallowing to ${\sim}10^{\circ}$ at a maximum locking depth of 20 km would require near-zero slip-deficit rates.
-Such a model is inconsistent with slip during the Wenchuan earthquake on the Beichuan and Penguan faults.
+To fit these observed, any planar or listric geometry with interseismic locking limited to the range-front (for example, a Beichuan-only fault geometry) would require near-zero slip-deficit rates.
+Such a model is inconsistent with slip during the Wenchuan earthquake on the Beichuan and Penguan faults and significant geologic estmates of shortening \citep{Hubbard2010, Li2010, Wang2013, Wang2014}, and .
 It is also inconsistent with co- and post-seismic slip on the detachment \citep{Qi2011, Fielding2013b}.
@@
 The least squares uncertainty is increased (from 1.5 mm/yr to 1.9 mm/yr) because the northwestern portion of velocities are poorly constrained once the observation is excluded.
-This northwestern-most observation is adjacent to the Longriba fault, suggesting that it should not be included.
+This northwestern-most observation is adjacent to the Longriba fault, and thus may be impacted by interseismic strain accumulation on this structure.
 However, the Longriba fault is primarily a dextral strike-slip fault \citep{Ren2013}, so it's influence should be predominantly orthogonal to the profile parallel velocities we examine here.
@@
 
+We also quantify the effect of explicitly treating surface topography by comparing models with and without surface topography, the local minimum in velocity (Figure \ref{fig:big_stack}a) northwest of the Beichuan is identified as the main topographically-influenced feature.
+The effects may be more pronounced with a shallower fault dip (e.g., many subduction zones).
+Despite the mediocre effect (${\sim}$15\%), geodetic analyses in steep regions should consider topographic effects to avoid systematic bias.
+This effect may be even more pronounced for coseismic faulting where deformation is localized in regions of steep topography.
+
 \section{Implications for the Longmen Shan earthquake cycle}
-The fundamental effect of allowing a 20 km detachment to accumulate and release elastic strain through the earthquake cycle is the lack of a range-front geodetic signature despite the significant ($5-10$ mm/yr) slip-deficit rate (\ref{fig:big_stack}a).
-Locking on the deep detachment obscures the signal of deformation on the range-front faults.
-The interseismic deformation is much more broadly distributed across more than 300 kilometers.
-According to this model, interseismically, the eastern Tibetan Plateau moves cohesively with the Sichuan Basin, whereas coseismic slip is concentrated at the boundary between these two blocks.
+The fundamental effect of allowing a 20 km deep detachment to accumulate and release elastic strain through the earthquake cycle is the lack of a range-front geodetic signature despite the significant ($5-10$ mm/yr) slip-deficit rate (\ref{fig:big_stack}a).
+Locking on the deep detachment obscures the signal of deformation on range-front faults.
+The interseismic deformation is much more broadly distributed across over 300 kilometers.
+According to this model, interseismically, the easternmost Tibetan Plateau moves cohesively with the Sichuan Basin, whereas coseismic slip is concentrated at the boundary between these two blocks.
 
@@
 
-While we have explained the observed the broad GPS interseismic velocity gradient as a result of a locked fault geometry including a deep detachment, we have not addressed when accumulated slip-deficit on the deep detachment is released.
+While we have explained the observed broad GPS interseismic velocity gradient as a result of a locked fault geometry including a deep detachment, we have not addressed when accumulated slip-deficit on the deep detachment is released.
 Geodetically constrained estimates of co- and post-seismic slip shows up to 5 m of slip on a deep detachment \citep{Qi2011}.
@@
 
-We also quantify the effect of explicitly treating surface topography by comparing models with and without surface topography, the local minimum in velocity (Figure \ref{fig:big_stack}a) northwest of the Beichuan is identified as the main topographically-influenced feature.
-The effects may be more pronounced with a shallower fault dip (e.g., many subduction zones).
-Despite the mediocre effect (${\sim}$15\%), geodetic analyses in steep regions should consider topographic effects to avoid systematic bias.
-This effect may be even more pronounced for coseismic faulting where deformation is localized in regions of steep topography.
-
 The 6-9 mm/yr slip-deficit rates estimated here can be used to provide a new kinematic constraint on earthquake recurrence intervals at the Longmen Shan (Figure \ref{fig:hazard}).
-Our shortening estimate suggests a recurrence interval for Wenchuan-like ruptures of approximately 600 years, with much shorter recurrence intervals for smaller events.
+Our shortening estimate suggests a recurrence interval for Wenchuan-like ruptures between 470 and 800 years (best fit: 590 years), with much shorter recurrence intervals for smaller events.
 This contrasts with the average recurrence of ~2000 years inferred from paleoseismic field investigations for the Beichuan fault \citep{Ran2010}.
 The 6-9 mm/yr slip-deficit rate estimated here would imply $\textrm{M}_{\textrm{w}}$ 9.0 ruptures for that recurrence interval, which would seemingly require an along strike rupture equal to or greater than the ${\sim}300$ km along strike length of the Longmen Shan.
-However, our slip-deficit estimate could include the effects of other imbricated structures.
+However, our slip-deficit estimate could include the effects of other imbricated structures, such as the Range Front blind thrust \citep{Hubbard2010, Wang2014}.
 As a result our estimate is a system-wide estimate while paleoseismic estimates have focused on the analysis of individual thrust faults.

### 2014-11-14 paper.tex
Slight typo edits.

@@
 The velocity gradient across the Longmen Shan rangefront is negligible (Figure \ref{fig:big_stack}a).
-To fit these observed, any planar or listric geometry with interseismic locking limited to the range-front (for example, a Beichuan-only fault geometry) would require near-zero slip-deficit rates.
-Such a model is inconsistent with slip during the Wenchuan earthquake on the Beichuan and Penguan faults and significant geologic estmates of shortening \citep{Hubbard2010, Li2010, Wang2013, Wang2014}, and .
+To fit these observed near-field data, any planar or listric geometry with interseismic locking limited to the range-front (for example, a Beichuan-only fault geometry) would require near-zero slip-deficit rates.
+Such a model is inconsistent with slip during the Wenchuan earthquake on the Beichuan and Penguan faults and with geologic estimates of significant shortening \citep{Hubbard2010, Li2010a, Wang2013, Wang2014}.
 It is also inconsistent with co- and post-seismic slip on the detachment \citep{Qi2011, Fielding2013b}.

### 2014-11-19 paper.tex
Applied Brendan's suggested edits as of Nov 17.

@@
 \title{Rapid slip-deficit rates at the eastern margin of the Tibetan plateau prior to the 2008 $\textrm{M}_\textrm{w}$ 7.9 Wenchuan earthquake}
-\author{T. Ben Thompson, Brendan J. Meade -- where to put JS, AP?}
+\author{T. Ben Thompson, Brendan J. Meade, Andreas Plesch, John H. Shaw}
 \usepackage{graphicx}
@@
 The Beichuan fault is part of an imbricate thrust system, with fault dips steepening to the northwest.
-Structural imbrication is typically produced by stacked listric faults that sole to detachments (TODO: CITE?).
+Structural imbrication is typically produced by stacked listric faults that sole to detachments.
 Furthermore, the large synclinorium parallel to, and west of, the Beichuan fault is consistent with the Beichuan fault shallowing to a detachment based on fault-bend folding theory \citep{Hubbard2010, Suppe1983}.
 In our model, the range-front Beichuan fault dips steeply at ${\sim}60^{\circ}$ at its surface trace shallowing to sub-horizontal ${\sim}20$ km depth where it merges with the detachment surface.
-At seismogenic depths, a 35$^\circ$ dip has been confirmed via focal mechanisms \citep{Hubbard2010}.
-Extrapolating from the regional dip of the Sichuan Basin as it extends beneath the Longmenshan range front \citep{Hubbard2010, Guo2013}, the detachment in our model dips ${\sim}1.5^{\circ}$ from an initial depth of 20km to a depth of 23km at a distance of 110km to the northwest.
+At 15 km depth, a 35$^{\circ}$ dip is consistent with focal mechanism inversions \citep{Hubbard2010}.
+Extrapolating from the regional dip of the Sichuan Basin as it extends beneath the Longmenshan range front \citep{Hubbard2010, Guo2013}, the detachment in our model dips ${\sim}1.5^{\circ}$ from an initial depth of 20 km to a depth of 23 km at a distance of 110 km to the northwest.
 The depth, dip, and extent of the detachment are similar to the best-fitting surface found by \citet{Qi2011} in fitting the co- and post-seismic slip of the Wenchuan earthquake.
@@
 We also quantify the effect of explicitly treating surface topography by comparing models with and without surface topography, the local minimum in velocity (Figure \ref{fig:big_stack}a) northwest of the Beichuan is identified as the main topographically-influenced feature.
-The effects may be more pronounced with a shallower fault dip (e.g., many subduction zones).
-Despite the mediocre effect (${\sim}$15\%), geodetic analyses in steep regions should consider topographic effects to avoid systematic bias.
-This effect may be even more pronounced for coseismic faulting where deformation is localized in regions of steep topography.
+Despite the ${\sim}$15\% effect, geodetic analyses in steep regions should consider topographic effects to avoid systematic bias.
+The effects may be more pronounced with a shallower fault dip or for coseismic displacements where deformation is localized in regions of steep topography.
 

### 2014-12-09 paper.tex
Reordered author list on paper.tex
Getting ready to submit to GRL.

@@
 \title{Rapid slip-deficit rates at the eastern margin of the Tibetan plateau prior to the 2008 $\textrm{M}_\textrm{w}$ 7.9 Wenchuan earthquake}
-\author{T. Ben Thompson, Brendan J. Meade, Andreas Plesch, John H. Shaw}
+\author{T. Ben Thompson, Andreas Plesch, John H. Shaw, Brendan J. Meade}
 \usepackage{graphicx}
@@
 \section{Abstract}
-The Longmen Shan is the steepest topographic front assocated with the India-Asia collision and was the site of the $\textrm{M}_{\textrm{w}}$ 7.9  Wenchuan earthquake.
-Shortening estimates across the Longmen Shan provide strain accumulation rates and clarify the eastward extrusion of the Tibetan plateau.
-Here, to explain the interseismic GPS velocities across the greater Longmen Shan region, we develop a boundary element model including earthquake cycle effects, topography, the westward dipping Beichuan fault, and a ${\sim}20$ km deep, shallowly dipping, detachment.
-The detachment is inferred from observations of slip during and after the Wenchuan earthquake and from structural considerations.
+The Longmen Shan is the steepest topographic front in the India-Asia collision and the site of the $\textrm{M}_{\textrm{w}}$ 7.9  Wenchuan earthquake.
+Longmen Shan shortening estimates provide strain accumulation rates and clarify the extrusion of the Tibetan plateau.
+Here, to explain the interseismic GPS velocities across the greater Longmen Shan region, we develop a boundary element model including earthquake cycle effects, topography, the westward dipping Beichuan fault, and a ${\sim}20$ km deep, shallowly dipping, detachment, inferred from observed afterslip and from structural considerations.
 Previous analyses which neglected the detachment and earthquake cycle effects have found shortening rates near zero.
-In contrast, we find that interseismic GPS data are consistent with a shortening rate of 5.7$\pm$1.5 mm/yr and maximum surface slip-deficit rate of 9.5 $\pm$ 2.5 mm/yr.
-This model provides a unified interpretation of geodetic deformation throughout the earthquake cycle and suggests that the Longmen Shan is an active fold-and-thrust belt with Wenchuan style earthquake recurrence intervals of as short as 600 years.
+In contrast, we find that interseismic GPS data are consistent with surface slip-deficit rate of 9.5 $\pm$ 2.5 mm/yr.
+This model unifies the interpretation of geodetic deformation throughout the earthquake cycle and suggests that the Longmen Shan is an active fold-and-thrust belt with Wenchuan-like recurrence intervals as short as 600 years.
 
@@
 We also quantify the effect of explicitly treating surface topography by comparing models with and without surface topography, the local minimum in velocity (Figure \ref{fig:big_stack}a) northwest of the Beichuan is identified as the main topographically-influenced feature.
-Despite the ${\sim}$15\% effect, geodetic analyses in steep regions should consider topographic effects to avoid systematic bias.
+Given the ${\sim}$15\% effect, geodetic analyses in steep regions should consider topographic effects to avoid systematic bias.
 The effects may be more pronounced with a shallower fault dip or for coseismic displacements where deformation is localized in regions of steep topography.
@@
 In addition to coseismic and short-term postseismic activity on the detachment, it is possible that intermittent aseismic slip could be an important mode of strain release.
-This is difficult to observationally constrain given that intersimsic GPS observations cover, at best, 6\% of a complete Longmen Shan earthquake cycle.
+This is difficult to observationally constrain given that interseismic GPS observations cover, at best, 6\% of a complete Longmen Shan earthquake cycle.
 Seismicity is approximately limited to the upper 20 km of crust in the range-front \citep{Li2010a}.

### 2014-12-10 agutmpl.tex
Submitting to GRL.

--- 2014-12-09 paper.tex
@@
-\documentclass[12pt]{article}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% AGUtmpl.tex: this template file is for articles formatted with LaTeX2e,
+% Modified July 2014
+%
+% This template includes commands and instructions
+% given in the order necessary to produce a final output that will
+% satisfy AGU requirements.
+%
+% PLEASE DO NOT USE YOUR OWN MACROS
+% DO NOT USE \newcommand, \renewcommand, or \def.
+%
+% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% All questions should be e-mailed to latex@agu.org.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Step 1: Set the \documentclass
+%
+% There are two options for article format: two column (default)
+% and draft.
+%
+% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
+% The draft option produces double spaced output.
+%
+% Choose the journal abbreviation for the journal you are
+% submitting to:
 
-\usepackage[utf8]{inputenc}
+% jgrga JOURNAL OF GEOPHYSICAL RESEARCH
+% gbc   GLOBAL BIOCHEMICAL CYCLES
+% grl   GEOPHYSICAL RESEARCH LETTERS
+% pal   PALEOCEANOGRAPHY
+% ras   RADIO SCIENCE
+% rog   REVIEWS OF GEOPHYSICS
+% tec   TECTONICS
+% wrr   WATER RESOURCES RESEARCH
+% gc    GEOCHEMISTRY, GEOPHYSICS, GEOSYSTEMS
+% sw    SPACE WEATHER
+% ms    JAMES
+% ef    EARTH'S FUTURE
+% ea    EARTH AND SPACE SCIENCE
+%
+%
+%
+% (If you are submitting to a journal other than jgrga,
+% substitute the initials of the journal for "jgrga" below.)
 
-\title{Rapid slip-deficit rates at the eastern margin of the Tibetan plateau prior to the 2008 $\textrm{M}_\textrm{w}$ 7.9 Wenchuan earthquake}
-\author{T. Ben Thompson, Andreas Plesch, John H. Shaw, Brendan J. Meade}
-\usepackage{graphicx}
-\usepackage[margin=1.0in]{geometry}
-\usepackage[nomarkers,figuresonly]{endfloat}
-\usepackage{setspace}
-\usepackage{natbib}
-\doublespacing
+\documentclass[draft,jgrga]{agutex}
+% To create numbered lines:
+
+% If you don't already have lineno.sty, you can download it from
+% http://www.ctan.org/tex-archive/macros/latex/contrib/ednotes/
+% (or search the internet for lineno.sty ctan), available at TeX Archive Network (CTAN).
+% Take care that you always use the latest version.
+
+% To activate the commands, uncomment \usepackage{lineno}
+% and \linenumbers*[1]command, below:
+
+% \usepackage{lineno}
+% \linenumbers*[1]
+%  To add line numbers to lines with equations:
+%  \begin{linenomath*}
+%  \begin{equation}
+%  \end{equation}
+%  \end{linenomath*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Figures and Tables
+%
+%
+% DO NOT USE \psfrag or \subfigure commands.
+%
+%
+%  Uncomment the following command to include .eps files
+%  (comment out this line for draft format):
+%  \usepackage[dvips]{graphicx}
+%
+%  Uncomment the following command to allow illustrations to print
+%   when using Draft:
+%  \setkeys{Gin}{draft=false}
+%
+% Substitute one of the following for [dvips] above
+% if you are using a different driver program and want to
+% proof your illustrations on your machine:
+%
+% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
+% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
+% [oztex], [textures]
+%
+% See how to enter figures and tables at the end of the article, after
+% references.
+%
+%% ------------------------------------------------------------------------ %%
+%
+%  ENTER PREAMBLE
+%
+%% ------------------------------------------------------------------------ %%
+
+% Author names in capital letters:
+\authorrunninghead{THOMPSON ET AL.}
+
+% Shorter version of title entered in capital letters:
+\titlerunninghead{LONGMEN SHAN BOUNDARY ELEMENT MODELS}
+
+%Corresponding author mailing address and e-mail address:
+%\authoraddr{Corresponding author: A. B. Smith,
+%Department of Hydrology and Water Resources, University of
+%Arizona, Harshbarger Building 11, Tucson, AZ 85721, USA.
+%(a.b.smith@hwr.arizona.edu)}
 
@@
 
-\maketitle
+%% ------------------------------------------------------------------------ %%
+%
+%  TITLE
+%
+%% ------------------------------------------------------------------------ %%
 
-\section{Abstract}
-The Longmen Shan is the steepest topographic front in the India-Asia collision and the site of the $\textrm{M}_{\textrm{w}}$ 7.9  Wenchuan earthquake.
-Longmen Shan shortening estimates provide strain accumulation rates and clarify the extrusion of the Tibetan plateau.
-Here, to explain the interseismic GPS velocities across the greater Longmen Shan region, we develop a boundary element model including earthquake cycle effects, topography, the westward dipping Beichuan fault, and a ${\sim}20$ km deep, shallowly dipping, detachment, inferred from observed afterslip and from structural considerations.
+\title{Rapid slip-deficit rates at the eastern margin of the Tibetan plateau prior to the 2008 $\textrm{M}_\textrm{w}$ 7.9 Wenchuan earthquake}
+%
+% e.g., \title{Terrestrial ring current:
+% Origin, formation, and decay $\alpha\beta\Gamma\Delta$}
+%
+
+%% ------------------------------------------------------------------------ %%
+%
+%  AUTHORS AND AFFILIATIONS
+%
+%% ------------------------------------------------------------------------ %%
+
+\author{T. Ben Thompson, Andreas Plesch, John H. Shaw, Brendan J. Meade}
+%Use \author{\altaffilmark{}} and \altaffiltext{}
+
+% \altaffilmark will produce footnote;
+% matching \altaffiltext will appear at bottom of page.
+
+% \authors{A. B. Smith,\altaffilmark{1}
+% Eric Brown,\altaffilmark{1,2} Rick Williams,\altaffilmark{3}
+% John B. McDougall\altaffilmark{4}, and S. Visconti\altaffilmark{5}}
+
+%\altaffiltext{1}{Department of Hydrology and Water Resources,
+%University of Arizona, Tucson, Arizona, USA.}
+
+%\altaffiltext{2}{Department of Geography, Ohio State University,
+%Columbus, Ohio, USA.}
+
+%\altaffiltext{3}{Department of Space Sciences, University of
+%Michigan, Ann Arbor, Michigan, USA.}
+
+%\altaffiltext{4}{Division of Hydrologic Sciences, Desert Research
+%Institute, Reno, Nevada, USA.}
+
+%\altaffiltext{5}{Dipartimento di Idraulica, Trasporti ed
+%Infrastrutture Civili, Politecnico di Torino, Turin, Italy.}
+
+%% ------------------------------------------------------------------------ %%
+%
+%  ABSTRACT
+%
+%% ------------------------------------------------------------------------ %%
+
+% >> Do NOT include any \begin...\end commands within
+% >> the body of the abstract.
+
+\begin{abstract}
+The Longmen Shan is the steepest topographic front at the India-Asia collision zone and the site of the $\textrm{M}_{\textrm{w}}$ 7.9 Wenchuan earthquake.
+Here, to explain the interseismic GPS velocities across the greater Longmen Shan region, we develop a boundary element model including earthquake cycle effects, topography, the westward dipping Beichuan fault, and a ${\sim}$20 km deep, shallowly dipping, detachment, inferred from observed afterslip and from structural considerations.
 Previous analyses which neglected the detachment and earthquake cycle effects have found shortening rates near zero.
-In contrast, we find that interseismic GPS data are consistent with surface slip-deficit rate of 9.5 $\pm$ 2.5 mm/yr.
-This model unifies the interpretation of geodetic deformation throughout the earthquake cycle and suggests that the Longmen Shan is an active fold-and-thrust belt with Wenchuan-like recurrence intervals as short as 600 years.
+In contrast, we find that interseismic GPS data are consistent with a shortening rate of 5.7 $\pm$ 1.5 mm/yr and maximum surface slip-deficit rate of 9.5 $\pm$ 2.5 mm/yr.
+This model unifies the interpretation of geodetic deformation throughout the earthquake cycle and suggests that the Longmen Shan is an active fold-and-thrust belt with of Wenchuan-like recurrence intervals as short as 600 years.
+\end{abstract}
+
+%% ------------------------------------------------------------------------ %%
+%
+%  BEGIN ARTICLE
+%
+%% ------------------------------------------------------------------------ %%
+
+% The body of the article must start with a \begin{article} command
+%
+% \end{article} must follow the references section, before the figures
+%  and tables.
+
+\begin{article}
+
+%% ------------------------------------------------------------------------ %%
+%
+%  TEXT
+%
+%% ------------------------------------------------------------------------ %%
 
 \section{Introduction}
+
 The Longmen Shan range is the steepest margin of the Tibetan Plateau rising 5 km over a distance of only 30 km at the western edge of the Sichuan Basin.
@@
 
-\bibliographystyle{Jcks}
-\bibliography{biblio,library}
+%%% End of body of article:
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional Appendix goes here
+%
+% \appendix resets counters and redefines section heads
+% but doesn't print anything.
+% After typing \appendix
+%
+%\section{Here Is Appendix Title}
+% will show
+% Appendix A: Here Is Appendix Title
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Optional Glossary or Notation section, goes here
+%
+%%%%%%%%%%%%%%
+% Glossary is only allowed in Reviews of Geophysics
+% \section*{Glossary}
+% \paragraph{Term}
+% Term Definition here
+%
+%%%%%%%%%%%%%%
+% Notation -- End each entry with a period.
+% \begin{notation}
+% Term & definition.\\
+% Second term & second definition.\\
+% \end{notation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%  ACKNOWLEDGMENTS
+
+\begin{acknowledgments}
+(Text here)
+\end{acknowledgments}
+
+%% ------------------------------------------------------------------------ %%
+%%  REFERENCE LIST AND TEXT CITATIONS
+%
+% Either type in your references using
+% \begin{thebibliography}{}
+% \bibitem{}
+% Text
+% \end{thebibliography}
+%
+% Or,
+%
+% If you use BiBTeX for your references, please use the agufull08.bst file (available at % ftp://ftp.agu.org/journals/latex/journals/Manuscript-Preparation/) to produce your .bbl
+% file and copy the contents into your paper here.
+%
+% Follow these steps:
+% 1. Run LaTeX on your LaTeX file.
+%
+% 2. Make sure the bibliography style appears as \bibliographystyle{agufull08}. Run BiBTeX on your LaTeX
+% file.
+%
+% 3. Open the new .bbl file containing the reference list and
+%   copy all the contents into your LaTeX file here.
+%
+% 4. Comment out the old \bibliographystyle and \bibliography commands.
+%
+% 5. Run LaTeX on your new file before submitting.
+%
+% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.
+
+\begin{thebibliography}{}
+
+%\providecommand{\natexlab}[1]{#1}
+%\expandafter\ifx\csname urlstyle\endcsname\relax
+%  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
+%  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
+%  \urlstyle{rm}\Url}\fi
+%
+%\bibitem[{\textit{Atkinson and Sloan}(1991)}]{AtkinsonSloan}
+%Atkinson, K., and I.~Sloan (1991), The numerical solution of first-kind
+%  logarithmic-kernel integral equations on smooth open arcs, \textit{Math.
+%  Comp.}, \textit{56}(193), 119--139.
+%
+%\bibitem[{\textit{Colton and Kress}(1983)}]{ColtonKress1}
+%Colton, D., and R.~Kress (1983), \textit{Integral Equation Methods in
+%  Scattering Theory}, John Wiley, New York.
+%
+%\bibitem[{\textit{Hsiao et~al.}(1991)\textit{Hsiao, Stephan, and
+%  Wendland}}]{StephanHsiao}
+%Hsiao, G.~C., E.~P. Stephan, and W.~L. Wendland (1991), On the {D}irichlet
+%  problem in elasticity for a domain exterior to an arc, \textit{J. Comput.
+%  Appl. Math.}, \textit{34}(1), 1--19.
+%
+%\bibitem[{\textit{Lu and Ando}(2012)}]{LuAndo}
+%Lu, P., and M.~Ando (2012), Difference of scattering geometrical optics
+%  components and line integrals of currents in modified edge representation,
+%  \textit{Radio Sci.}, \textit{47},  RS3007, \doi{10.1029/2011RS004899}.
+
+\end{thebibliography}
+
+%Reference citation examples:
+
+%...as shown by \textit{Kilby} [2008].
+%...as shown by {\textit  {Lewin}} [1976], {\textit  {Carson}} [1986], {\textit  {Bartholdy and Billi}} [2002], and {\textit  {Rinaldi}} [2003].
+%...has been shown [\textit{Kilby et al.}, 2008].
+%...has been shown [{\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].
+%...has been shown [e.g., {\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].
+
+%...as shown by \citet{jskilby}.
+%...as shown by \citet{lewin76}, \citet{carson86}, \citet{bartoldy02}, and \citet{rinaldi03}.
+%...has been shown \citep{jskilbye}.
+%...has been shown \citep{lewin76,carson86,bartoldy02,rinaldi03}.
+%...has been shown \citep [e.g.,][]{lewin76,carson86,bartoldy02,rinaldi03}.
+%
+% Please use ONLY \citet and \citep for reference citations.
+% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).
+
+%% ------------------------------------------------------------------------ %%
+%
+%  END ARTICLE
+%
+%% ------------------------------------------------------------------------ %%
+\end{article}
+%
+%
+%% Enter Figures and Tables here:
+%
+% DO NOT USE \psfrag or \subfigure commands.
+%
+% Figure captions go below the figure.
+% Table titles go above tables; all other caption information
+%  should be placed in footnotes below the table.
+%
+%----------------
+% EXAMPLE FIGURE
+%
+ %\begin{figure}
+ %\noindent\includegraphics[width=20pc]{samplefigure.eps}
+ %\caption{Caption text here}
+ %\label{figure_label}
+ %\end{figure}
+%
+% ---------------
+% EXAMPLE TABLE
+%
+%\begin{table}
+%\caption{Time of the Transition Between Phase 1 and Phase 2\tablenotemark{a}}
+%\centering
+%\begin{tabular}{l c}
+%\hline
+% Run  & Time (min)  \\
+%\hline
+%  $l1$  & 260   \\
+%  $l2$  & 300   \\
+%  $l3$  & 340   \\
+%  $h1$  & 270   \\
+%  $h2$  & 250   \\
+%  $h3$  & 380   \\
+%  $r1$  & 370   \\
+%  $r2$  & 390   \\
+%\hline
+%\end{tabular}
+%\tablenotetext{a}{Footnote text here.}
+%\end{table}
+
+% See below for how to make sideways figures or tables.
+
 \end{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+More Information and Advice:
+
+%% ------------------------------------------------------------------------ %%
+%
+%  SECTION HEADS
+%
+%% ------------------------------------------------------------------------ %%
+
+% Capitalize the first letter of each word (except for
+% prepositions, conjunctions, and articles that are
+% three or fewer letters).
+
+% AGU follows standard outline style; therefore, there cannot be a section 1 without
+% a section 2, or a section 2.3.1 without a section 2.3.2.
+% Please make sure your section numbers are balanced.
+% ---------------
+% Level 1 head
+%
+% Use the \section{} command to identify level 1 heads;
+% type the appropriate head wording between the curly
+% brackets, as shown below.
+%
+%An example:
+%\section{Level 1 Head: Introduction}
+%
+% ---------------
+% Level 2 head
+%
+% Use the \subsection{} command to identify level 2 heads.
+%An example:
+%\subsection{Level 2 Head}
+%
+% ---------------
+% Level 3 head
+%
+% Use the \subsubsection{} command to identify level 3 heads
+%An example:
+%\subsubsection{Level 3 Head}
+%
+%---------------
+% Level 4 head
+%
+% Use the \subsubsubsection{} command to identify level 3 heads
+% An example:
+%\subsubsubsection{Level 4 Head} An example.
+%
+%% ------------------------------------------------------------------------ %%
+%
+%  IN-TEXT LISTS
+%
+%% ------------------------------------------------------------------------ %%
+%
+% Do not use bulleted lists; enumerated lists are okay.
+% \begin{enumerate}
+% \item
+% \item
+% \item
+% \end{enumerate}
+%
+%% ------------------------------------------------------------------------ %%
+%
+%  EQUATIONS
+%
+%% ------------------------------------------------------------------------ %%
+
+% Single-line equations are centered.
+% Equation arrays will appear left-aligned.
+
+Math coded inside display math mode \[ ...\]
+ will not be numbered, e.g.,:
+ \[ x^2=y^2 + z^2\]
+
+ Math coded inside \begin{equation} and \end{equation} will
+ be automatically numbered, e.g.,:
+ \begin{equation}
+ x^2=y^2 + z^2
+ \end{equation}
+
+% IF YOU HAVE MULTI-LINE EQUATIONS, PLEASE
+% BREAK THE EQUATIONS INTO TWO OR MORE LINES
+% OF SINGLE COLUMN WIDTH (20 pc, 8.3 cm)
+% using double backslashes (\\).
+
+% To create multiline equations, use the
+% \begin{eqnarray} and \end{eqnarray} environment
+% as demonstrated below.
+\begin{eqnarray}
+  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
+        && + (y - y_{0}) \sin \Theta  \nonumber \\
+  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
+        && + (y - y_{0}) \cos \Theta.
+\end{eqnarray}
+
+%If you don't want an equation number, use the star form:
+%\begin{eqnarray*}...\end{eqnarray*}
+
+% Break each line at a sign of operation
+% (+, -, etc.) if possible, with the sign of operation
+% on the new line.
+
+% Indent second and subsequent lines to align with
+% the first character following the equal sign on the
+% first line.
+
+% Use an \hspace{} command to insert horizontal space
+% into your equation if necessary. Place an appropriate
+% unit of measure between the curly braces, e.g.
+% \hspace{1in}; you may have to experiment to achieve
+% the correct amount of space.
+
+%% ------------------------------------------------------------------------ %%
+%
+%  EQUATION NUMBERING: COUNTER
+%
+%% ------------------------------------------------------------------------ %%
+
+% You may change equation numbering by resetting
+% the equation counter or by explicitly numbering
+% an equation.
+
+% To explicitly number an equation, type \eqnum{}
+% (with the desired number between the brackets)
+% after the \begin{equation} or \begin{eqnarray}
+% command.  The \eqnum{} command will affect only
+% the equation it appears with; LaTeX will number
+% any equations appearing later in the manuscript
+% according to the equation counter.
+%
+
+% If you have a multiline equation that needs only
+% one equation number, use a \nonumber command in
+% front of the double backslashes (\\) as shown in
+% the multiline equation above.
+
+%% ------------------------------------------------------------------------ %%
+%
+%  SIDEWAYS FIGURE AND TABLE EXAMPLES
+%
+%% ------------------------------------------------------------------------ %%
+%
+% For tables and figures, add \usepackage{rotating} to the paper and add the rotating.sty file to the folder.
+% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
+%
+% \begin{sidewaysfigure}
+% \includegraphics[width=20pc]{samplefigure.eps}
+% \caption{caption here}
+% \label{label_here}
+% \end{sidewaysfigure}
+%
+%
+%
+% \begin{sidewaystable}
+% \caption{}
+% \begin{tabular}
+% Table layout here.
+% \end{tabular}
+% \end{sidewaystable}
+%
+%

### 2014-12-10 paper.tex
Submitting to GRL.

@@
 \section{Abstract}
-The Longmen Shan is the steepest topographic front in the India-Asia collision and the site of the $\textrm{M}_{\textrm{w}}$ 7.9  Wenchuan earthquake.
-Longmen Shan shortening estimates provide strain accumulation rates and clarify the extrusion of the Tibetan plateau.
-Here, to explain the interseismic GPS velocities across the greater Longmen Shan region, we develop a boundary element model including earthquake cycle effects, topography, the westward dipping Beichuan fault, and a ${\sim}20$ km deep, shallowly dipping, detachment, inferred from observed afterslip and from structural considerations.
+The Longmen Shan is the steepest topographic front at the India-Asia collision zone and the site of the $\textrm{M}_{\textrm{w}}$ 7.9 Wenchuan earthquake.
+Here, to explain the interseismic GPS velocities across the greater Longmen Shan region, we develop a boundary element model including earthquake cycle effects, topography, the westward dipping Beichuan fault, and a ${\sim}$20 km deep, shallowly dipping, detachment, inferred from observed afterslip and from structural considerations.
 Previous analyses which neglected the detachment and earthquake cycle effects have found shortening rates near zero.
-In contrast, we find that interseismic GPS data are consistent with surface slip-deficit rate of 9.5 $\pm$ 2.5 mm/yr.
-This model unifies the interpretation of geodetic deformation throughout the earthquake cycle and suggests that the Longmen Shan is an active fold-and-thrust belt with Wenchuan-like recurrence intervals as short as 600 years.
+In contrast, we find that interseismic GPS data are consistent with a shortening rate of 5.7 $\pm$ 1.5 mm/yr and maximum surface slip-deficit rate of 9.5 $\pm$ 2.5 mm/yr.
+This model unifies the interpretation of geodetic deformation throughout the earthquake cycle and suggests that the Longmen Shan is an active fold-and-thrust belt with of Wenchuan-like recurrence intervals as short as 600 years.
 
